---
title: "Zooming in on specific species - dsLibs v. ssLibs (NPost dataset)"
author: "Keri Burge"
date: "`r Sys.Date()`"
output: html_document
---

```{r, eval=TRUE, echo=FALSE, message=FALSE}
#Load in all necessary packages
library(knitr)
library(data.table)
library(ggplot2)
library(viridis)
library(rstatix)
library(mixOmics)
library(janitor)
library(dplyr)
library(vegan)
library(ggrepel)
library(ggpubr)
library(tidyverse)

```


```{r setup}
knitr::opts_knit$set(root.dir = normalizePath(".."))
```


# Colors
```{r, colors}
double_color <- "#04387a"
single_color <- "#6fbfdc"
  
```

#Metadata
```{r}
metadata <- fread("./05-results/normalize_kraken/dsvss_npost_w_rl_subsampling_metadata.tsv") 

metadata_no_rl <- metadata %>%
  filter(Read_length_subsampling == "none")
```



#Prepare the read length tables
```{r}
files <- list.files(path = "./05-results/normalize_kraken/oral_well_preserved_PCs", pattern = ".oral_well_preserved_PCs.tsv", full.names = T)

taxIDs <- fread("./05-results/normalize_kraken/dsvss_npost_oral_PCs_taxIDs.tsv")

rl_oral_PCs_table <- function(files) {
  oral_gc_rl <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(read_header = 1,
          taxid = 2,
          read_length = 3,
          GC = 4) %>%
    left_join(., taxIDs %>%
              dplyr::select(taxid, taxonomy)) %>%
    select(-read_header, -taxid, -GC) %>%
    group_by(taxonomy, read_length) %>%
    summarise(counts = n()) %>%
    mutate(Sequencing_ID = basename(fn)) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".oral_well_preserved_PCs.tsv", "")) %>%
    inner_join(., metadata_no_rl %>%
              dplyr::select(Sequencing_ID, Strandedness)) 
  })
  
  oral_gc_rl_df <- do.call(rbind, oral_gc_rl) 
  oral_gc_rl_df <- oral_gc_rl_df %>%
    filter(!str_detect(Sequencing_ID, "BAN001")) %>%
    filter(!str_detect(Sequencing_ID, "FUM003")) 
  
  return(oral_gc_rl_df)
}


oral_pcs_rl <- rl_oral_PCs_table(files)

gc_categories <- fread("./05-results/normalize_kraken/dsvss_npost_oral_pc_species_gc_categories.csv")

oral_pcs_rl <- oral_pcs_rl %>%
  full_join(., gc_categories, by = "taxonomy")
```

# Calculate the median read length for each library
```{r}
#function to calculate the median read length of oral reads per libraryID
median_rl_summary_table <- function(sample_list, oral_reads_length_table) {
  
  #Create a blank dataframe
  read_length_summary_table <- data.frame(Individual_ID = sample_list,
                                          Double = numeric(length(sample_list)),
                                          Single = numeric(length(sample_list)),
                                          stringsAsFactors = FALSE)
  
  #Function to calculate median read length based on dataset, sample, and strandedness
  calculate_median_rl <- function(data, sample, strandedness) {
    rl_summary <- data %>%
      dplyr::select(Library_ID, Individual_ID, Strandedness, read_length, counts) %>%
      mutate(read_length = as.integer(as.character(read_length))) %>%
      filter(Individual_ID == sample, Strandedness == strandedness) %>%
      uncount(counts) %>%
      mutate(read_length = as.integer(read_length)) %>%
      summarize(median_rl = median(read_length))
    
    return(rl_summary$median_rl)
  }
  
  
  for (sample in sample_list) {
    rl_value_of_median_ds <- calculate_median_rl(oral_reads_length_table, sample, "Double")
    rl_value_of_median_ss <- calculate_median_rl(oral_reads_length_table, sample, "Single")
    read_length_summary_table[read_length_summary_table$Individual_ID == sample, "Double"] <- rl_value_of_median_ds
    read_length_summary_table[read_length_summary_table$Individual_ID == sample, "Single"] <- rl_value_of_median_ss
    
   }

  return(read_length_summary_table)
  
}

samples <- metadata_no_rl %>%
  dplyr::select(Individual_ID) %>%
  unique() %>%
  pull()

```


#Read length distributions
## Lautropia mirabilis -> high GC content (65.5)
```{r}
l_mirabilis_table <- oral_pcs_rl %>%
  ungroup() %>%
  filter(taxonomy == "Lautropia mirabilis") %>%
  full_join(., metadata_no_rl %>%
              select(Sequencing_ID, Library_ID, Individual_ID)) %>%
  select(-taxonomy)
```

Plot total number of reads per library
```{r}
column_colors <- c(double_color, single_color)

reads_histogram <- l_mirabilis_table %>%
  select(Sequencing_ID, counts) %>%
  group_by(Sequencing_ID) %>%
  mutate(total_counts = sum(counts)) %>%
  ungroup() %>%
  select(-counts) %>%
  unique() %>%
  mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>%
  filter(total_counts > 100) %>%
  inner_join(., metadata_no_rl %>%
              select(Sequencing_ID, Individual_ID, Strandedness)) %>%
  ggplot(., aes(x=Strandedness, y=total_counts, fill=Strandedness)) +
    geom_bar(stat="identity") +
    theme_minimal() +
    labs(title = "", x = "Individual ID & Strandedness", y = "Read Count") +
    guides(fill=guide_legend(title="")) +
    theme(legend.position="bottom") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    guides(fill = guide_legend(reverse=TRUE)) +
    scale_fill_discrete(name = "", type = column_colors) +
    facet_grid(~Individual_ID, scales = "free_x", space = "free") +
    theme(strip.text.x = element_text(angle = 90, hjust = 1, size = 14),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 14),
          strip.text = element_text(size = 12),
          legend.text = element_text(size = 10),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14)) 

reads_histogram

#ggsave("./06_publication/supplemental_figures/npost_species/l_mirabilis/l_mirabilis_read_count_bar_plot.pdf", plot = reads_histogram, device = "pdf", scale = 1, width = 10, height = 6, units = c("in"), dpi = 300)
```


```{r}
median_rl_l_mirabilis <- median_rl_summary_table(samples, l_mirabilis_table) %>%
  pivot_longer(-Individual_ID, values_to = "median_rl", names_to = "Strandedness")

```

FUM003 and BAN001 do not have reads for this species 

CDM019
```{r}
read_length_table <- l_mirabilis_table %>%
  filter(Individual_ID == "CDM019") 

median_rl_info <- median_rl_l_mirabilis %>%
  filter(Individual_ID == "CDM019") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100000) %>%
  #mutate(Length = as.numeric(Length)) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 50)) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/l_mirabilis/l_mirabilis_CDM019_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


CDM020
```{r}
read_length_table <- l_mirabilis_table %>%
  filter(Individual_ID == "CDM020") 

median_rl_info <- median_rl_l_mirabilis %>%
  filter(Individual_ID == "CDM020") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 1500)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/l_mirabilis/l_mirabilis_CDM020_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```

FUM002    
```{r}
read_length_table <- l_mirabilis_table %>%
  filter(Individual_ID == "FUM002") 

median_rl_info <- median_rl_l_mirabilis %>%
  filter(Individual_ID == "FUM002") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 60)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/l_mirabilis/l_mirabilis_FUM002_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GDN001 
```{r}
read_length_table <- l_mirabilis_table %>%
  filter(Individual_ID == "GDN001") 

median_rl_info <- median_rl_l_mirabilis %>%
  filter(Individual_ID == "GDN001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_col(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, width = 1) +
    geom_col(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, width = 1) +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 2)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    #geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/l_mirabilis/l_mirabilis_GDN001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```



GOY005  
```{r}
read_length_table <- l_mirabilis_table %>%
  filter(Individual_ID == "GOY005") 

median_rl_info <- median_rl_l_mirabilis %>%
  filter(Individual_ID == "GOY005") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
     theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 1500)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/l_mirabilis/l_mirabilis_GOY005_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GOY006 
```{r}
read_length_table <- l_mirabilis_table %>%
  filter(Individual_ID == "GOY006") 

median_rl_info <- median_rl_l_mirabilis %>%
  filter(Individual_ID == "GOY006") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 300)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/l_mirabilis/l_mirabilis_GOY006_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


PES001 
```{r}
read_length_table <- l_mirabilis_table %>%
  filter(Individual_ID == "PES001") 

median_rl_info <- median_rl_l_mirabilis %>%
  filter(Individual_ID == "PES001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 300)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/l_mirabilis/l_mirabilis_PES001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


VLD001 
```{r}
read_length_table <- l_mirabilis_table %>%
  filter(Individual_ID == "VLD001") 

median_rl_info <- median_rl_l_mirabilis %>%
  filter(Individual_ID == "VLD001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
     theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 3)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/l_mirabilis/l_mirabilis_VLD001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


## Neisseria elongata -> mid GC content (54)
```{r}
n_elongata_table <- oral_pcs_rl %>%
  ungroup() %>%
  filter(taxonomy == "Neisseria elongata") %>%
  full_join(., metadata_no_rl %>%
              select(Sequencing_ID, Library_ID, Individual_ID)) %>%
  select(-taxonomy)
```


Plot total number of reads per library
```{r}
column_colors <- c(double_color, single_color)

reads_histogram <- n_elongata_table %>%
  select(Sequencing_ID, counts) %>%
  group_by(Sequencing_ID) %>%
  mutate(total_counts = sum(counts)) %>%
  ungroup() %>%
  select(-counts) %>%
  unique() %>%
  mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>%
  filter(total_counts > 100) %>%
  inner_join(., metadata_no_rl %>%
              select(Sequencing_ID, Individual_ID, Strandedness)) %>%
  ggplot(., aes(x=Strandedness, y=total_counts, fill=Strandedness)) +
    geom_bar(stat="identity") +
    theme_minimal() +
    labs(title = "", x = "Individual ID & Strandedness", y = "Read Count") +
    guides(fill=guide_legend(title="")) +
    theme(legend.position="bottom") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    guides(fill = guide_legend(reverse=TRUE)) +
    scale_fill_discrete(name = "", type = column_colors) +
    facet_grid(~Individual_ID, scales = "free_x", space = "free") +
    theme(strip.text.x = element_text(angle = 90, hjust = 1, size = 14),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 14),
          strip.text = element_text(size = 12),
          legend.text = element_text(size = 10),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14)) 

reads_histogram

#ggsave("./06_publication/supplemental_figures/npost_species/n_elongata/n_elongata_read_count_bar_plot.pdf", plot = reads_histogram, device = "pdf", scale = 1, width = 10, height = 6, units = c("in"), dpi = 300)
```


```{r}
median_rl_n_elongata <- median_rl_summary_table(samples, n_elongata_table) %>%
  pivot_longer(-Individual_ID, values_to = "median_rl", names_to = "Strandedness")
```

FUM003 and BAN001 do not have reads for this species 

CDM019 
```{r}
read_length_table <- n_elongata_table %>%
  filter(Individual_ID == "CDM019") 

median_rl_info <- median_rl_n_elongata %>%
  filter(Individual_ID == "CDM019") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 6)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/n_elongata/n_elongata_CDM019_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


CDM020
```{r}
read_length_table <- n_elongata_table %>%
  filter(Individual_ID == "CDM020") 

median_rl_info <- median_rl_n_elongata %>%
  filter(Individual_ID == "CDM020") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 300)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/n_elongata/n_elongata_CDM020_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


FUM002
```{r}
read_length_table <- n_elongata_table %>%
  filter(Individual_ID == "FUM002") 

median_rl_info <- median_rl_n_elongata %>%
  filter(Individual_ID == "FUM002") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 60)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/n_elongata/n_elongata_FUM002_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GDN001
```{r}
read_length_table <- n_elongata_table %>%
  filter(Individual_ID == "GDN001") 

median_rl_info <- median_rl_n_elongata %>%
  filter(Individual_ID == "GDN001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_col(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, width = 1) +
    geom_col(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, width = 1) +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 3)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    #geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/n_elongata/n_elongata_GDN001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GOY005
```{r}
read_length_table <- n_elongata_table %>%
  filter(Individual_ID == "GOY005") 

median_rl_info <- median_rl_n_elongata %>%
  filter(Individual_ID == "GOY005") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 200)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/n_elongata/n_elongata_GOY005_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GOY006
```{r}
read_length_table <- n_elongata_table %>%
  filter(Individual_ID == "GOY006") 

median_rl_info <- median_rl_n_elongata %>%
  filter(Individual_ID == "GOY006") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 60)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/n_elongata/n_elongata_GOY006_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


PES001
```{r}
read_length_table <- n_elongata_table %>%
  filter(Individual_ID == "PES001") 

median_rl_info <- median_rl_n_elongata %>%
  filter(Individual_ID == "PES001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 20)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/n_elongata/n_elongata_PES001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


VLD001
```{r}
read_length_table <- n_elongata_table %>%
  filter(Individual_ID == "VLD001") 

median_rl_info <- median_rl_n_elongata %>%
  filter(Individual_ID == "VLD001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_col(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, width = 1) +
    geom_col(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, width = 1) +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 1)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    #geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/n_elongata/n_elongata_VLD001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


## Aggregatibacter kilianii -> low GC content (43)
```{r}
a_kilianii_table <- oral_pcs_rl %>%
  ungroup() %>%
  filter(taxonomy == "Aggregatibacter kilianii") %>%
  full_join(., metadata_no_rl %>%
              select(Sequencing_ID, Library_ID, Individual_ID)) %>%
  select(-taxonomy)

median_rl_a_kilianii <- median_rl_summary_table(samples, a_kilianii_table) %>%
  pivot_longer(-Individual_ID, values_to = "median_rl", names_to = "Strandedness")
```


Plot total number of reads per library
```{r}
column_colors <- c(double_color, single_color)

reads_histogram <- a_kilianii_table %>%
  select(Sequencing_ID, counts) %>%
  group_by(Sequencing_ID) %>%
  mutate(total_counts = sum(counts)) %>%
  ungroup() %>%
  select(-counts) %>%
  unique() %>%
  mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>%
  filter(total_counts > 100) %>%
  inner_join(., metadata_no_rl %>%
              select(Sequencing_ID, Individual_ID, Strandedness)) %>%
  ggplot(., aes(x=Strandedness, y=total_counts, fill=Strandedness)) +
    geom_bar(stat="identity") +
    theme_minimal() +
    labs(title = "", x = "Individual ID & Strandedness", y = "Read Count") +
    guides(fill=guide_legend(title="")) +
    theme(legend.position="bottom") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    guides(fill = guide_legend(reverse=TRUE)) +
    scale_fill_discrete(name = "", type = column_colors) +
    facet_grid(~Individual_ID, scales = "free_x", space = "free") +
    theme(strip.text.x = element_text(angle = 90, hjust = 1, size = 14),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 14),
          strip.text = element_text(size = 12),
          legend.text = element_text(size = 10),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14)) 

reads_histogram

#ggsave("./06_publication/supplemental_figures/npost_species/a_kilianii/a_kilianii_read_count_bar_plot.pdf", plot = reads_histogram, device = "pdf", scale = 1, width = 10, height = 6, units = c("in"), dpi = 300)
```

BAN001, FUM003, GDN001 dsLib, and VLD001 dsLib do not have this species 

CDM019 
```{r}
read_length_table <- a_kilianii_table %>%
  filter(Individual_ID == "CDM019") 

median_rl_info <- median_rl_a_kilianii %>%
  filter(Individual_ID == "CDM019") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 3)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_kilianii/a_kilianii_CDM019_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


CDM020
```{r}
read_length_table <- a_kilianii_table %>%
  filter(Individual_ID == "CDM020") 

median_rl_info <- median_rl_a_kilianii %>%
  filter(Individual_ID == "CDM020") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 40)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_kilianii/a_kilianii_CDM020_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```

FUM002
```{r}
read_length_table <- a_kilianii_table %>%
  filter(Individual_ID == "FUM002") 

median_rl_info <- median_rl_a_kilianii %>%
  filter(Individual_ID == "FUM002") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 15)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_kilianii/a_kilianii_FUM002_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GDN001 - only ssLib
```{r}
read_length_table <- a_kilianii_table %>%
  filter(Individual_ID == "GDN001") 

median_rl_info <- median_rl_a_kilianii %>%
  filter(Individual_ID == "GDN001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    #geom_col(data = . %>% filter(Strandedness == "Double"),
    #          aes(x = read_length, y = counts),
    #          fill = "#04387a", alpha = 1, width = 1) +
    geom_col(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, width = 1) +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 1)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    #geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_kilianii/a_kilianii_GDN001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GOY005
```{r}
read_length_table <- a_kilianii_table %>%
  filter(Individual_ID == "GOY005") 

median_rl_info <- median_rl_a_kilianii %>%
  filter(Individual_ID == "GOY005") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 800)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_kilianii/a_kilianii_GOY005_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GOY006
```{r}
read_length_table <- a_kilianii_table %>%
  filter(Individual_ID == "GOY006") 

median_rl_info <- median_rl_a_kilianii %>%
  filter(Individual_ID == "GOY006") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 50)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_kilianii/a_kilianii_GOY006_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


PES001
```{r}
read_length_table <- a_kilianii_table %>%
  filter(Individual_ID == "PES001") 

median_rl_info <- median_rl_a_kilianii %>%
  filter(Individual_ID == "PES001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 25)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_kilianii/a_kilianii_PES001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


VLD001 - only ssLibs
```{r}
read_length_table <- a_kilianii_table %>%
  filter(Individual_ID == "VLD001") 

median_rl_info <- median_rl_a_kilianii %>%
  filter(Individual_ID == "VLD001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    #geom_bar(data = . %>% filter(Strandedness == "Double"),
    #          aes(x = read_length, y = counts),
    #          fill = "#04387a", alpha = 1, stat = "identity") +
    geom_col(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, width = 1) +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 1)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    #geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_kilianii/a_kilianii_VLD001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```

## Actinomyces dentalis -> high GC content (73)
```{r}
a_dentalis_table <- oral_pcs_rl %>%
  ungroup() %>%
  filter(taxonomy == "Actinomyces dentalis") %>%
  full_join(., metadata_no_rl %>%
              select(Sequencing_ID, Library_ID, Individual_ID)) %>%
  select(-taxonomy)
```

Plot total number of reads per library
```{r}
column_colors <- c(double_color, single_color)

reads_histogram <- a_dentalis_table %>%
  select(Sequencing_ID, counts) %>%
  group_by(Sequencing_ID) %>%
  mutate(total_counts = sum(counts)) %>%
  ungroup() %>%
  select(-counts) %>%
  unique() %>%
  mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>%
  filter(total_counts > 100) %>%
  inner_join(., metadata_no_rl %>%
              select(Sequencing_ID, Individual_ID, Strandedness)) %>%
  ggplot(., aes(x=Strandedness, y=total_counts, fill=Strandedness)) +
    geom_bar(stat="identity") +
    theme_minimal() +
    labs(title = "", x = "Individual ID & Strandedness", y = "Read Count") +
    guides(fill=guide_legend(title="")) +
    theme(legend.position="bottom") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    guides(fill = guide_legend(reverse=TRUE)) +
    scale_fill_discrete(name = "", type = column_colors) +
    facet_grid(~Individual_ID, scales = "free_x", space = "free") +
    theme(strip.text.x = element_text(angle = 90, hjust = 1, size = 14),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 14),
          strip.text = element_text(size = 12),
          legend.text = element_text(size = 10),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14)) 

reads_histogram

#ggsave("./06_publication/supplemental_figures/npost_species/a_dentalis/a_dentalis_read_count_bar_plot.pdf", plot = reads_histogram, device = "pdf", scale = 1, width = 10, height = 6, units = c("in"), dpi = 300)
```


```{r}
median_rl_a_dentalis <- median_rl_summary_table(samples, a_dentalis_table) %>%
  pivot_longer(-Individual_ID, values_to = "median_rl", names_to = "Strandedness")
```

FUM003 and BAN001 do not have reads for this species 

CDM019
```{r}
read_length_table <- a_dentalis_table %>%
  filter(Individual_ID == "CDM019") 

median_rl_info <- median_rl_a_dentalis %>%
  filter(Individual_ID == "CDM019") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100000) %>%
  #mutate(Length = as.numeric(Length)) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 50)) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_dentalis/a_dentalis_CDM019_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


CDM020
```{r}
read_length_table <- a_dentalis_table %>%
  filter(Individual_ID == "CDM020") 

median_rl_info <- median_rl_a_dentalis %>%
  filter(Individual_ID == "CDM020") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 1500)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_dentalis/a_dentalis_CDM020_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```

FUM002    
```{r}
read_length_table <- a_dentalis_table %>%
  filter(Individual_ID == "FUM002") 

median_rl_info <- median_rl_a_dentalis %>%
  filter(Individual_ID == "FUM002") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 60)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_dentalis/a_dentalis_FUM002_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GDN001 
```{r}
read_length_table <- a_dentalis_table %>%
  filter(Individual_ID == "GDN001") 

median_rl_info <- median_rl_a_dentalis %>%
  filter(Individual_ID == "GDN001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 60)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_dentalis/a_dentalis_GDN001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```



GOY005  
```{r}
read_length_table <- a_dentalis_table %>%
  filter(Individual_ID == "GOY005") 

median_rl_info <- median_rl_a_dentalis %>%
  filter(Individual_ID == "GOY005") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
     theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 100)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_dentalis/a_dentalis_GOY005_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GOY006 
```{r}
read_length_table <- a_dentalis_table %>%
  filter(Individual_ID == "GOY006") 

median_rl_info <- median_rl_a_dentalis %>%
  filter(Individual_ID == "GOY006") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 300)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_dentalis/a_dentalis_GOY006_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


PES001 
```{r}
read_length_table <- a_dentalis_table %>%
  filter(Individual_ID == "PES001") 

median_rl_info <- median_rl_a_dentalis %>%
  filter(Individual_ID == "PES001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 300)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_dentalis/a_dentalis_PES001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


VLD001 
```{r}
read_length_table <- a_dentalis_table %>%
  filter(Individual_ID == "VLD001") 

median_rl_info <- median_rl_a_dentalis %>%
  filter(Individual_ID == "VLD001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
     theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 1000)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/a_dentalis/a_dentalis_VLD001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


## Flexilinea sp001717545 -> mid GC content (55)
```{r}
flexilinea_table <- oral_pcs_rl %>%
  ungroup() %>%
  filter(taxonomy == "Flexilinea sp001717545") %>%
  full_join(., metadata_no_rl %>%
              select(Sequencing_ID, Library_ID, Individual_ID)) %>%
  select(-taxonomy)
```

Plot total number of reads per library
```{r}
column_colors <- c(double_color, single_color)

reads_histogram <- flexilinea_table %>%
  select(Sequencing_ID, counts) %>%
  group_by(Sequencing_ID) %>%
  mutate(total_counts = sum(counts)) %>%
  ungroup() %>%
  select(-counts) %>%
  unique() %>%
  mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>%
  filter(total_counts > 100) %>%
  inner_join(., metadata_no_rl %>%
              select(Sequencing_ID, Individual_ID, Strandedness)) %>%
  ggplot(., aes(x=Strandedness, y=total_counts, fill=Strandedness)) +
    geom_bar(stat="identity") +
    theme_minimal() +
    labs(title = "", x = "Individual ID & Strandedness", y = "Read Count") +
    guides(fill=guide_legend(title="")) +
    theme(legend.position="bottom") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    guides(fill = guide_legend(reverse=TRUE)) +
    scale_fill_discrete(name = "", type = column_colors) +
    facet_grid(~Individual_ID, scales = "free_x", space = "free") +
    theme(strip.text.x = element_text(angle = 90, hjust = 1, size = 14),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 14),
          strip.text = element_text(size = 12),
          legend.text = element_text(size = 10),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14)) 

reads_histogram

#ggsave("./06_publication/supplemental_figures/npost_species/flexilinea/flexilinea_read_count_bar_plot.pdf", plot = reads_histogram, device = "pdf", scale = 1, width = 10, height = 6, units = c("in"), dpi = 300)
```


```{r}
median_rl_flexilinea <- median_rl_summary_table(samples, flexilinea_table) %>%
  pivot_longer(-Individual_ID, values_to = "median_rl", names_to = "Strandedness")
```

FUM003 and BAN001 do not have reads for this species 

CDM019
```{r}
read_length_table <- flexilinea_table %>%
  filter(Individual_ID == "CDM019") 

median_rl_info <- median_rl_flexilinea %>%
  filter(Individual_ID == "CDM019") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100000) %>%
  #mutate(Length = as.numeric(Length)) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    #expand_limits(y = c(0, 50)) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/flexilinea/flexilinea_CDM019_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


CDM020
```{r}
read_length_table <- flexilinea_table %>%
  filter(Individual_ID == "CDM020") 

median_rl_info <- median_rl_flexilinea %>%
  filter(Individual_ID == "CDM020") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 1500)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/flexilinea/flexilinea_CDM020_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```

FUM002    
```{r}
read_length_table <- flexilinea_table %>%
  filter(Individual_ID == "FUM002") 

median_rl_info <- median_rl_flexilinea %>%
  filter(Individual_ID == "FUM002") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    #expand_limits(y = c(0, 60)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/flexilinea/flexilinea_FUM002_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GDN001 
```{r}
read_length_table <- flexilinea_table %>%
  filter(Individual_ID == "GDN001") 

median_rl_info <- median_rl_flexilinea %>%
  filter(Individual_ID == "GDN001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    #expand_limits(y = c(0, 60)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/flexilinea/flexilinea_GDN001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GOY005  
```{r}
read_length_table <- flexilinea_table %>%
  filter(Individual_ID == "GOY005") 

median_rl_info <- median_rl_flexilinea %>%
  filter(Individual_ID == "GOY005") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
     theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 50)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/flexilinea/flexilinea_GOY005_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GOY006 
```{r}
read_length_table <- flexilinea_table %>%
  filter(Individual_ID == "GOY006") 

median_rl_info <- median_rl_flexilinea %>%
  filter(Individual_ID == "GOY006") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    #expand_limits(y = c(0, 300)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/flexilinea/flexilinea_GOY006_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


PES001 
```{r}
read_length_table <- flexilinea_table %>%
  filter(Individual_ID == "PES001") 

median_rl_info <- median_rl_flexilinea %>%
  filter(Individual_ID == "PES001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 300)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/flexilinea/flexilinea_PES001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


VLD001 
```{r}
read_length_table <- flexilinea_table %>%
  filter(Individual_ID == "VLD001") 

median_rl_info <- median_rl_flexilinea %>%
  filter(Individual_ID == "VLD001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
     theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    #expand_limits(y = c(0, 1000)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/flexilinea/flexilinea_VLD001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```

## Methanobrevibacter A oralis -> low GC content (27.5)
```{r}
m_oralis_table <- oral_pcs_rl %>%
  ungroup() %>%
  filter(taxonomy == "Methanobrevibacter A oralis") %>%
  full_join(., metadata_no_rl %>%
              select(Sequencing_ID, Library_ID, Individual_ID)) %>%
  select(-taxonomy)
```

Plot total number of reads per library
```{r}
column_colors <- c(double_color, single_color)

reads_histogram <- m_oralis_table %>%
  select(Sequencing_ID, counts) %>%
  group_by(Sequencing_ID) %>%
  mutate(total_counts = sum(counts)) %>%
  ungroup() %>%
  select(-counts) %>%
  unique() %>%
  mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>%
  filter(total_counts > 100) %>%
  inner_join(., metadata_no_rl %>%
              select(Sequencing_ID, Individual_ID, Strandedness)) %>%
  ggplot(., aes(x=Strandedness, y=total_counts, fill=Strandedness)) +
    geom_bar(stat="identity") +
    theme_minimal() +
    labs(title = "", x = "Individual ID & Strandedness", y = "Read Count") +
    guides(fill=guide_legend(title="")) +
    theme(legend.position="bottom") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    guides(fill = guide_legend(reverse=TRUE)) +
    scale_fill_discrete(name = "", type = column_colors) +
    facet_grid(~Individual_ID, scales = "free_x", space = "free") +
    theme(strip.text.x = element_text(angle = 90, hjust = 1, size = 14),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 14),
          strip.text = element_text(size = 12),
          legend.text = element_text(size = 10),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14)) 

reads_histogram

#ggsave("./06_publication/supplemental_figures/npost_species/m_oralis/m_oralis_read_count_bar_plot.pdf", plot = reads_histogram, device = "pdf", scale = 1, width = 10, height = 6, units = c("in"), dpi = 300)
```


```{r}
median_rl_m_oralis <- median_rl_summary_table(samples, m_oralis_table) %>%
  pivot_longer(-Individual_ID, values_to = "median_rl", names_to = "Strandedness")
```

FUM003 and BAN001 do not have reads for this species 

CDM019
```{r}
read_length_table <- m_oralis_table %>%
  filter(Individual_ID == "CDM019") 

median_rl_info <- median_rl_m_oralis %>%
  filter(Individual_ID == "CDM019") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100000) %>%
  #mutate(Length = as.numeric(Length)) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    #expand_limits(y = c(0, 50)) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/m_oralis/m_oralis_CDM019_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


CDM020
```{r}
read_length_table <- m_oralis_table %>%
  filter(Individual_ID == "CDM020") 

median_rl_info <- median_rl_m_oralis %>%
  filter(Individual_ID == "CDM020") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    #expand_limits(y = c(0, 1500)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/m_oralis/m_oralis_CDM020_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```

FUM002    
```{r}
read_length_table <- m_oralis_table %>%
  filter(Individual_ID == "FUM002") 

median_rl_info <- median_rl_m_oralis %>%
  filter(Individual_ID == "FUM002") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 60)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/m_oralis/m_oralis_FUM002_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GDN001 
```{r}
read_length_table <- m_oralis_table %>%
  filter(Individual_ID == "GDN001") 

median_rl_info <- median_rl_m_oralis %>%
  filter(Individual_ID == "GDN001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    #expand_limits(y = c(0, 60)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/m_oralis/m_oralis_GDN001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GOY005  
```{r}
read_length_table <- m_oralis_table %>%
  filter(Individual_ID == "GOY005") 

median_rl_info <- median_rl_m_oralis %>%
  filter(Individual_ID == "GOY005") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
     theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    #expand_limits(y = c(0, 100)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/m_oralis/m_oralis_GOY005_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


GOY006 
```{r}
read_length_table <- m_oralis_table %>%
  filter(Individual_ID == "GOY006") 

median_rl_info <- median_rl_m_oralis %>%
  filter(Individual_ID == "GOY006") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    #expand_limits(y = c(0, 300)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/m_oralis/m_oralis_GOY006_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


PES001 
```{r}
read_length_table <- m_oralis_table %>%
  filter(Individual_ID == "PES001") 

median_rl_info <- median_rl_m_oralis %>%
  filter(Individual_ID == "PES001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    #expand_limits(y = c(0, 300)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/m_oralis/m_oralis_PES001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


VLD001 
```{r}
read_length_table <- m_oralis_table %>%
  filter(Individual_ID == "VLD001") 

median_rl_info <- median_rl_m_oralis %>%
  filter(Individual_ID == "VLD001") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  #mutate(counts = counts / 100) %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    #labs(x = "Read Length (bp)", y = expression(`Read Count` ~ (x10^2))) + 
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
     theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    #expand_limits(y = c(0, 1000)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/supplemental_figures/npost_species/m_oralis/m_oralis_VLD001_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```

#Damage plots
```{r}
#colors for well-preserved samples

sample_colors <- c("#88ccef", "#44aa9a", "#147733", "#342288", "#ddcc77", "#999934", "#cc6777", "#882355", "#aa4599", "#dddddd")
```

```{r}
#well-preserved samples

metadata_wp <- metadata_no_rl %>%
  filter(kraken2_preservation == "Good")
  
wp_samples <- metadata_wp %>%
  select(Sequencing_ID) %>%
  pull()

```

##Individual damage plots function
```{r}
double_color <- "#04387a"
single_color <- "#6fbfdc"

strandedness_colors <- c(double_color, single_color)

individual_damage_plots <- function(combined_5_prime_table, combined_3_prime_table, sample_list, folder_path, species) {
  #Find the max y-axis value
  find_max_y <- function(combined_5_prime_table, combined_3_prime_table, sample) {
    max_y_value <- combined_5_prime_table %>%
      full_join(., combined_3_prime_table) %>%
      filter(Individual_ID == sample) %>%
      summarize(max_frequency = max(frequency, na.rm = TRUE)) %>%
      select(max_frequency) %>%
      unique() %>%
      pull(max_frequency)
  }
  
  #5' prime plot function
  create_5_prime_plot <- function(combined_5_prime_table, sample) {
    combined_5_prime_plot <- combined_5_prime_table %>%
      filter(Individual_ID == sample) %>%
      ggplot(aes(x=position, y=frequency, color = Strandedness)) +
        geom_line() +
        scale_color_manual(values = strandedness_colors) +
        labs(x = "Position", y = "Frequency") +
        theme_minimal() +
        theme(legend.position = "right",
              axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
              axis.text.y = element_text(size = 22), 
              axis.title = element_text(size = 22),  
              legend.text = element_text(size = 16),  
              legend.title = element_text(size = 20),
              plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) +
      expand_limits(y = c(0, max_y_value))
  }
  
   #3' prime plot function
  create_3_prime_plot <- function(combined_3_prime_table, sample) {
    combined_3_prime_plot <- combined_3_prime_table %>%
      filter(Individual_ID == sample) %>%
      mutate(position = (position*-1)) %>%
      ggplot(aes(x=position, y=frequency, color = Strandedness)) +
      geom_line() +
      scale_y_continuous(
        name = NULL,                  
        sec.axis = dup_axis(name = "Frequency")) +
        labs(x = "Position") +
        scale_color_manual(values = strandedness_colors) +
        theme_minimal() +
      theme(
        legend.position = "right",
        axis.text.x = element_text(angle = 90, hjust = 1, size = 20),
        axis.text.y.left  = element_blank(),   # hide left labels
        axis.ticks.y.left = element_blank(),
        axis.text.y.right = element_text(size = 22),
        axis.title.y.right = element_text(size = 22),
        axis.title = element_text(size = 22),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 20),
        plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) +
      expand_limits(y = c(0, max_y_value))
  }
  
  #Combine the 5' and 3' plots
  create_combined_damage_plot <- function(combined_5_prime_plot, create_3_prime_plot, plot_title) {
  damage_plot <- ggarrange(
    combined_5_prime_plot, combined_3_prime_plot,
    labels = c("5'", "3'"), 
    ncol = 2, nrow = 1,  common.legend = TRUE, legend = "right")
  }
  
  #Iterate over each sample in sample_list to create a plot for each Individual_ID
  for (sample in sample_list) {
    
    #Find the max y-axis value
    max_y_value <- find_max_y(combined_5_prime_table, combined_3_prime_table, sample)
    
    #Create a 5' and 3' plot for each Individual
    plot_5_prime <- create_5_prime_plot(combined_5_prime_table, sample)
    plot_3_prime <- create_3_prime_plot(combined_3_prime_table, sample)
    
   #Combine the 5' and 3' plots
    damage_plot_combined <- ggarrange(
      plot_5_prime, plot_3_prime,
      labels = c("5'", "3'"), 
      ncol = 2, nrow = 1,  common.legend = TRUE, legend = "right", align = "hv")
  
    #Annotate the plot to add the Individual_ID 
    damage_plot_combined_annotated <- annotate_figure(damage_plot_combined, top = text_grob(paste0(sample), face = "bold", size = 22))
    
    #Save the combined plot to a single file for each sample
    ggsave(
      filename = file.path(folder_path, paste0(sample, "_", species, "_damage_plot.pdf")), 
      plot = damage_plot_combined_annotated, 
      device = "pdf", 
      scale = 1, 
      width = 10, 
      height = 4, 
      units = "in", 
      dpi = 300
    )
  }
  
}
```

## Lautropia mirabilis
###dsLibs
```{r}
files_5_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/lautropia_mirabilis", pattern = "5pCtoT_freq.txt", recursive = TRUE, full.names = T)


smiley_plot_table_5_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/lautropia_mirabilis/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_l_mirabilis_q25l30_rmdup/5pCtoT_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Double")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_5_prime_dsLibs <- smiley_plot_table_5_prime(files_5_prime)

```


```{r}
files_3_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/lautropia_mirabilis", pattern = "3pGtoA_freq.txt", recursive = TRUE, full.names = T)

smiley_plot_table_3_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/lautropia_mirabilis/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_l_mirabilis_q25l30_rmdup/3pGtoA_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Double")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_3_prime_dsLibs <- smiley_plot_table_3_prime(files_3_prime)

```


```{r}
dsLib_5_prime <- table_5_prime_dsLibs %>%
  filter(Strandedness == "Double") %>%
  ggplot(aes(x=position, y=frequency, color = Individual_ID)) +
  geom_line() +
  scale_color_manual(values = sample_colors) +
  labs(x = "Position", y = "Frequency") +
  theme_minimal() +
  theme(legend.position = "right",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 16),  
          legend.title = element_text(size = 20),
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) +
  expand_limits(y = c(0, 0.4))

dsLib_5_prime 
```

```{r}
dsLib_3_prime <- table_3_prime_dsLibs %>%
  mutate(position = (position*-1)) %>%
  filter(Strandedness == "Double") %>%
  ggplot(aes(x=position, y=frequency, color = Individual_ID)) +
  geom_line() +
  scale_y_continuous(
    name = NULL,                  
    sec.axis = dup_axis(name = "Frequency")) +
  labs(x = "Position") +
  scale_color_manual(values = sample_colors) +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 90, hjust = 1, size = 20),
    axis.text.y.left  = element_blank(),   # hide left labels
    axis.ticks.y.left = element_blank(),
    axis.text.y.right = element_text(size = 22),
    axis.title.y.right = element_text(size = 22),
     axis.title = element_text(size = 22),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 20),
    plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt"))


dsLib_3_prime 
```


```{r}
summary_plots_dsLibs <- ggarrange(
  dsLib_5_prime, dsLib_3_prime,
  labels = c("", ""), 
  ncol = 2, nrow = 1,  common.legend = TRUE, legend = "right"
)

summary_plots_dsLibs

#ggsave("./06_publication/supplemental_figures/npost_species/l_mirabilis/l_mirabilis_dsLib_damage.pdf", plot = summary_plots_dsLibs, device = "pdf", scale = 1, width = 10, height = 4, units = c("in"), dpi = 300)
```


###ssLibs
```{r}
files_5_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/lautropia_mirabilis", pattern = "5pCtoT_freq.txt", recursive = TRUE, full.names = T)


smiley_plot_table_5_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/lautropia_mirabilis/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_l_mirabilis_q25l30_rmdup/5pCtoT_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
    filter(Strandedness == "Single")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_5_prime_ssLibs <- smiley_plot_table_5_prime(files_5_prime)

```


```{r}
files_3_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/lautropia_mirabilis", pattern = "3pGtoA_freq.txt", recursive = TRUE, full.names = T)

smiley_plot_table_3_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/lautropia_mirabilis/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_l_mirabilis_q25l30_rmdup/3pGtoA_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) 
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_3_prime_ssLibs <- smiley_plot_table_3_prime(files_3_prime)

```


```{r}
ssLib_5_prime <- table_5_prime_ssLibs %>%
  ggplot(aes(x=position, y=frequency, color = Individual_ID)) +
  geom_line() +
  scale_color_manual(values = sample_colors) +
  labs(x = "Position", y = "Frequency") +
  theme_minimal() +
  theme(legend.position = "right",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 16),  
          legend.title = element_text(size = 20),
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) +
  expand_limits(y = c(0, 0.4))

ssLib_5_prime 
```

```{r}
ssLib_3_prime <- table_3_prime_ssLibs %>%
  mutate(position = (position*-1)) %>%
  ggplot(aes(x=position, y=frequency, color = Individual_ID)) +
  geom_line() +
  scale_y_continuous(
    name = NULL,                  
    sec.axis = dup_axis(name = "Frequency")) +
  labs(x = "Position") +
  scale_color_manual(values = sample_colors) +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 90, hjust = 1, size = 20),
    axis.text.y.left  = element_blank(),   # hide left labels
    axis.ticks.y.left = element_blank(),
    axis.text.y.right = element_text(size = 22),
    axis.title.y.right = element_text(size = 22),
     axis.title = element_text(size = 22),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 20),
    plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) +
  expand_limits(y = c(0, 0.4))


ssLib_3_prime 
```


```{r}
summary_plots_ssLibs <- ggarrange(
  ssLib_5_prime, ssLib_3_prime,
  labels = c("", ""), 
  ncol = 2, nrow = 1,  common.legend = TRUE, legend = "right"
)

summary_plots_ssLibs

#ggsave("./06_publication/supplemental_figures/npost_species/l_mirabilis/l_mirabilis_ssLib_damage.pdf", plot = summary_plots_ssLibs, device = "pdf", scale = 1, width = 10, height = 4, units = c("in"), dpi = 300)
```

###Individual plots
```{r}
combined_5_prime <- table_5_prime_ssLibs %>%
  full_join(., table_5_prime_dsLibs) 

combined_3_prime <- table_3_prime_ssLibs %>%
  full_join(., table_3_prime_dsLibs)
```

```{r}
samples <- c("CDM019", "CDM020", "FUM002", "GOY005", "GOY006", "PES001")

individual_damage_plots(combined_5_prime, combined_3_prime, samples, "./06_publication/supplemental_figures/npost_species/l_mirabilis/damage_plots", "l_mirabilis")
```

## Neisseria elongata
###dsLibs
```{r}
files_5_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/neisseria_elongata", pattern = "5pCtoT_freq.txt", recursive = TRUE, full.names = T)


smiley_plot_table_5_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/neisseria_elongata/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_n_elongata_q25l30_rmdup/5pCtoT_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Double")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_5_prime_dsLibs <- smiley_plot_table_5_prime(files_5_prime)

```


```{r}
files_3_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/neisseria_elongata", pattern = "3pGtoA_freq.txt", recursive = TRUE, full.names = T)

smiley_plot_table_3_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/neisseria_elongata/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_n_elongata_q25l30_rmdup/3pGtoA_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Double")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_3_prime_dsLibs <- smiley_plot_table_3_prime(files_3_prime)

```


```{r}
dsLib_5_prime <- table_5_prime %>%
  filter(Strandedness == "Double") %>%
  ggplot(aes(x=position, y=frequency, color = Individual_ID)) +
  geom_line() +
  scale_color_manual(values = sample_colors) +
  labs(x = "Position", y = "Frequency") +
  theme_minimal() +
  theme(legend.position = "right",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 16),  
          legend.title = element_text(size = 20),
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) +
  expand_limits(y = c(0, 0.4))

dsLib_5_prime 
```


```{r}
dsLib_3_prime <- table_3_prime %>%
  mutate(position = (position*-1)) %>%
  filter(Strandedness == "Double") %>%
  ggplot(aes(x=position, y=frequency, color = Individual_ID)) +
  geom_line() +
  scale_y_continuous(
    name = NULL,                  
    sec.axis = dup_axis(name = "Frequency")) +
  labs(x = "Position") +
  scale_color_manual(values = sample_colors) +
  theme_minimal() +
  theme(
    lGCF_000429225.1egend.position = "right",
    axis.text.x = element_text(angle = 90, hjust = 1, size = 20),
    axis.text.y.left  = element_blank(),   # hide left labels
    axis.ticks.y.left = element_blank(),
    axis.text.y.right = element_text(size = 22),
    axis.title.y.right = element_text(size = 22),
     axis.title = element_text(size = 22),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 20),
    plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt"))


dsLib_3_prime 
```


```{r}
summary_plots_dsLibs <- ggarrange(
  dsLib_5_prime, dsLib_3_prime,
  labels = c("", ""), 
  ncol = 2, nrow = 1,  common.legend = TRUE, legend = "right"
)

summary_plots_dsLibs

#ggsave("./06_publication/supplemental_figures/npost_species/n_elongata/n_elongata_dsLib_damage.pdf", plot = summary_plots_dsLibs, device = "pdf", scale = 1, width = 10, height = 4, units = c("in"), dpi = 300)
```


###ssLibs
```{r}
files_5_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/neisseria_elongata", pattern = "5pCtoT_freq.txt", recursive = TRUE, full.names = T)


smiley_plot_table_5_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/neisseria_elongata/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_n_elongata_q25l30_rmdup/5pCtoT_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Single")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_5_prime_ssLibs <- smiley_plot_table_5_prime(files_5_prime)

```


```{r}
files_3_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/neisseria_elongata", pattern = "3pGtoA_freq.txt", recursive = TRUE, full.names = T)

smiley_plot_table_3_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/neisseria_elongata/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_n_elongata_q25l30_rmdup/3pGtoA_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Single")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_3_prime_ssLibs <- smiley_plot_table_3_prime(files_3_prime)

```


```{r}
ssLib_5_prime <- table_5_prime_ssLibs %>%
  ggplot(aes(x=position, y=frequency, color = Individual_ID)) +
  geom_line() +
  scale_color_manual(values = sample_colors) +
  labs(x = "Position", y = "Frequency") +
  theme_minimal() +
  theme(legend.position = "right",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 16),  
          legend.title = element_text(size = 20),
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) +
  expand_limits(y = c(0, 0.4))

ssLib_5_prime 
```


```{r}
ssLib_3_prime <- table_3_prime_ssLibs %>%
  mutate(position = (position*-1)) %>%
  ggplot(aes(x=position, y=frequency, color = Individual_ID)) +
  geom_line() +
  scale_y_continuous(
    name = NULL,                  
    sec.axis = dup_axis(name = "Frequency")) +
  labs(x = "Position") +
  scale_color_manual(values = sample_colors) +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 90, hjust = 1, size = 20),
    axis.text.y.left  = element_blank(),   # hide left labels
    axis.ticks.y.left = element_blank(),
    axis.text.y.right = element_text(size = 22),
    axis.title.y.right = element_text(size = 22),
     axis.title = element_text(size = 22),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 20),
    plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) +
  expand_limits(y = c(0, 0.4))


ssLib_3_prime 
```


```{r}
summary_plots_ssLibs <- ggarrange(
  ssLib_5_prime, ssLib_3_prime,
  labels = c("", ""), 
  ncol = 2, nrow = 1,  common.legend = TRUE, legend = "right"
)

summary_plots_ssLibs

#ggsave("./06_publication/supplemental_figures/npost_species/n_elongata/n_elongata_ssLib_damage.pdf", plot = summary_plots_ssLibs, device = "pdf", scale = 1, width = 10, height = 4, units = c("in"), dpi = 300)
```

###Individual plots
```{r}
combined_5_prime <- table_5_prime_ssLibs %>%
  full_join(., table_5_prime_dsLibs) 

combined_3_prime <- table_3_prime_ssLibs %>%
  full_join(., table_3_prime_dsLibs)
```

```{r}
samples <- c("CDM019", "CDM020", "FUM002", "GOY005", "GOY006", "PES001")

individual_damage_plots(combined_5_prime, combined_3_prime, samples, "./06_publication/supplemental_figures/npost_species/n_elongata/damage_plots", "n_elongata")
```


## Aggregatibacter kilianni
###dsLibs
```{r}
files_5_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/aggregatibacter_kilianii", pattern = "5pCtoT_freq.txt", recursive = TRUE, full.names = T)


smiley_plot_table_5_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/aggregatibacter_kilianii/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_a_kilianii_q25l30_rmdup/5pCtoT_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Double")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_5_prime_dsLibs <- smiley_plot_table_5_prime(files_5_prime)

```


```{r}
files_3_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/aggregatibacter_kilianii", pattern = "3pGtoA_freq.txt", recursive = TRUE, full.names = T)

smiley_plot_table_3_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/aggregatibacter_kilianii/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_a_kilianii_q25l30_rmdup/3pGtoA_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Double")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_3_prime_dsLibs <- smiley_plot_table_3_prime(files_3_prime)

```


```{r}
dsLib_5_prime <- table_5_prime %>%
  filter(Strandedness == "Double") %>%
  ggplot(aes(x=position, y=frequency, color = Individual_ID)) +
  geom_line() +
  scale_color_manual(values = sample_colors) +
  labs(x = "Position", y = "Frequency") +
  theme_minimal() +
  theme(legend.position = "right",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 16),  
          legend.title = element_text(size = 20),
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) +
  expand_limits(y = c(0, 1))

dsLib_5_prime 
```

```{r}
dsLib_3_prime <- table_3_prime %>%
  mutate(position = (position*-1)) %>%
  filter(Strandedness == "Double") %>%
  ggplot(aes(x=position, y=frequency, color = Individual_ID)) +
  geom_line() +
  scale_y_continuous(
    name = NULL,                  
    sec.axis = dup_axis(name = "Frequency")) +
  labs(x = "Position") +
  scale_color_manual(values = sample_colors) +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 90, hjust = 1, size = 20),
    axis.text.y.left  = element_blank(),   # hide left labels
    axis.ticks.y.left = element_blank(),
    axis.text.y.right = element_text(size = 22),
    axis.title.y.right = element_text(size = 22),
     axis.title = element_text(size = 22),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 20),
    plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) +
    expand_limits(y = c(0, 0.5))


dsLib_3_prime 
```


```{r}
summary_plots_dsLibs <- ggarrange(
  dsLib_5_prime, dsLib_3_prime,
  labels = c("", ""), 
  ncol = 2, nrow = 1,  common.legend = TRUE, legend = "right"
)

summary_plots_dsLibs

#ggsave("./06_publication/supplemental_figures/npost_species/a_kilianii/a_kilianii_dsLib_damage.pdf", plot = summary_plots_dsLibs, device = "pdf", scale = 1, width = 10, height = 4, units = c("in"), dpi = 300)
```


###ssLibs
```{r}
files_5_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/aggregatibacter_kilianii", pattern = "5pCtoT_freq.txt", recursive = TRUE, full.names = T)


smiley_plot_table_5_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/aggregatibacter_kilianii/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_a_kilianii_q25l30_rmdup/5pCtoT_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Single")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_5_prime_ssLibs <- smiley_plot_table_5_prime(files_5_prime)

```


```{r}
files_3_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/aggregatibacter_kilianii", pattern = "3pGtoA_freq.txt", recursive = TRUE, full.names = T)

smiley_plot_table_3_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/aggregatibacter_kilianii/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_a_kilianii_q25l30_rmdup/3pGtoA_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Single")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_3_prime_ssLibs <- smiley_plot_table_3_prime(files_3_prime)

```


```{r}
ssLib_5_prime <- table_5_prime_ssLibs %>%
  ggplot(aes(x=position, y=frequency, color = Individual_ID)) +
  geom_line() +
  scale_color_manual(values = sample_colors) +
  labs(x = "Position", y = "Frequency") +
  theme_minimal() +
  theme(legend.position = "right",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 16),  
          legend.title = element_text(size = 20),
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) +
  expand_limits(y = c(0, 0.4))

ssLib_5_prime 
```

```{r}
ssLib_3_prime <- table_3_prime_ssLibs %>%
  mutate(position = (position*-1)) %>%
  ggplot(aes(x=position, y=frequency, color = Individual_ID)) +
  geom_line() +
  scale_y_continuous(
    name = NULL,                  
    sec.axis = dup_axis(name = "Frequency")) +
  labs(x = "Position") +
  scale_color_manual(values = sample_colors) +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 90, hjust = 1, size = 20),
    axis.text.y.left  = element_blank(),   # hide left labels
    axis.ticks.y.left = element_blank(),
    axis.text.y.right = element_text(size = 22),
    axis.title.y.right = element_text(size = 22),
     axis.title = element_text(size = 22),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 20),
    plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) +
  expand_limits(y = c(0, 0.5))


ssLib_3_prime 
```


```{r}
summary_plots_ssLibs <- ggarrange(
  ssLib_5_prime, ssLib_3_prime,
  labels = c("", ""), 
  ncol = 2, nrow = 1,  common.legend = TRUE, legend = "right"
)

summary_plots_ssLibs

#ggsave("./06_publication/supplemental_figures/npost_species/a_kilianii/a_kilianii_ssLib_damage.pdf", plot = summary_plots_ssLibs, device = "pdf", scale = 1, width = 10, height = 4, units = c("in"), dpi = 300)
```


###Individual plots
```{r}
combined_5_prime <- table_5_prime_ssLibs %>%
  full_join(., table_5_prime_dsLibs) 

combined_3_prime <- table_3_prime_ssLibs %>%
  full_join(., table_3_prime_dsLibs)
```

```{r}
samples <- c("CDM020", "FUM002", "GOY005", "GOY006", "PES001")

individual_damage_plots(combined_5_prime, combined_3_prime, samples, "./06_publication/supplemental_figures/npost_species/a_kilianii/damage_plots", "a_kiliani")
```


## Actinomyces dentalis
###dsLibs
```{r}
files_5_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/actinomyces_dentalis", pattern = "5pCtoT_freq.txt", recursive = TRUE, full.names = T)


smiley_plot_table_5_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/actinomyces_dentalis/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_a_dentalis_q25l30_rmdup/5pCtoT_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Double")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_5_prime_dsLibs <- smiley_plot_table_5_prime(files_5_prime)

```


```{r}
files_3_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/actinomyces_dentalis", pattern = "3pGtoA_freq.txt", recursive = TRUE, full.names = T)

smiley_plot_table_3_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/actinomyces_dentalis/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_a_dentalis_q25l30_rmdup/3pGtoA_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Double")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_3_prime_dsLibs <- smiley_plot_table_3_prime(files_3_prime)

```


###ssLibs
```{r}
files_5_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/actinomyces_dentalis", pattern = "5pCtoT_freq.txt", recursive = TRUE, full.names = T)


smiley_plot_table_5_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/actinomyces_dentalis/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_a_dentalis_q25l30_rmdup/5pCtoT_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Single")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_5_prime_ssLibs <- smiley_plot_table_5_prime(files_5_prime)

```


```{r}
files_3_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/actinomyces_dentalis", pattern = "3pGtoA_freq.txt", recursive = TRUE, full.names = T)

smiley_plot_table_3_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/actinomyces_dentalis/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_a_dentalis_q25l30_rmdup/3pGtoA_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Single")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_3_prime_ssLibs <- smiley_plot_table_3_prime(files_3_prime)

```


###Individual plots
```{r}
combined_5_prime <- table_5_prime_ssLibs %>%
  full_join(., table_5_prime_dsLibs) 

combined_3_prime <- table_3_prime_ssLibs %>%
  full_join(., table_3_prime_dsLibs)
```


```{r}
samples <- c("CDM019", "CDM020", "FUM002", "GDN001", "GOY005", "GOY006", "PES001", "VLD001")

individual_damage_plots(combined_5_prime, combined_3_prime, samples, "./06_publication/supplemental_figures/npost_species/a_dentalis/damage_plots", "a_dentalis")
```


## Flexilinea sp001717545
###dsLibs
```{r}
files_5_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/flexilinea_sp001717545", pattern = "5pCtoT_freq.txt", recursive = TRUE, full.names = T)


smiley_plot_table_5_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/flexilinea_sp001717545/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_f_sp001717545_q25l30_rmdup/5pCtoT_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Double")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_5_prime_dsLibs <- smiley_plot_table_5_prime(files_5_prime)

```


```{r}
files_3_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/flexilinea_sp001717545", pattern = "3pGtoA_freq.txt", recursive = TRUE, full.names = T)

smiley_plot_table_3_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/flexilinea_sp001717545/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_f_sp001717545_q25l30_rmdup/3pGtoA_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Double")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_3_prime_dsLibs <- smiley_plot_table_3_prime(files_3_prime)

```


###ssLibs
```{r}
files_5_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/flexilinea_sp001717545", pattern = "5pCtoT_freq.txt", recursive = TRUE, full.names = T)


smiley_plot_table_5_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/flexilinea_sp001717545/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_f_sp001717545_q25l30_rmdup/5pCtoT_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Single")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_5_prime_ssLibs <- smiley_plot_table_5_prime(files_5_prime)

```


```{r}
files_3_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/flexilinea_sp001717545", pattern = "3pGtoA_freq.txt", recursive = TRUE, full.names = T)

smiley_plot_table_3_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/flexilinea_sp001717545/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_f_sp001717545_q25l30_rmdup/3pGtoA_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Single")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_3_prime_ssLibs <- smiley_plot_table_3_prime(files_3_prime)

```


###Individual plots
```{r}
combined_5_prime <- table_5_prime_ssLibs %>%
  full_join(., table_5_prime_dsLibs) 

combined_3_prime <- table_3_prime_ssLibs %>%
  full_join(., table_3_prime_dsLibs)
```


```{r}
samples <- c("CDM019", "CDM020", "FUM002", "GDN001", "GOY005", "GOY006", "PES001", "VLD001")

individual_damage_plots(combined_5_prime, combined_3_prime, samples, "./06_publication/supplemental_figures/npost_species/flexilinea/damage_plots", "flexilinea")
```



## Methanobrevibacter_A oralis
###dsLibs
```{r}
files_5_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/methanobrevibacter_oralis", pattern = "5pCtoT_freq.txt", recursive = TRUE, full.names = T)


smiley_plot_table_5_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/methanobrevibacter_oralis/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_m_oralis_q25l30_rmdup/5pCtoT_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Double")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_5_prime_dsLibs <- smiley_plot_table_5_prime(files_5_prime)

```


```{r}
files_3_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/methanobrevibacter_oralis", pattern = "3pGtoA_freq.txt", recursive = TRUE, full.names = T)

smiley_plot_table_3_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/methanobrevibacter_oralis/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_m_oralis_q25l30_rmdup/3pGtoA_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Double")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_3_prime_dsLibs <- smiley_plot_table_3_prime(files_3_prime)

```


###ssLibs
```{r}
files_5_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/methanobrevibacter_oralis", pattern = "5pCtoT_freq.txt", recursive = TRUE, full.names = T)


smiley_plot_table_5_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/methanobrevibacter_oralis/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_m_oralis_q25l30_rmdup/5pCtoT_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Single")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_5_prime_ssLibs <- smiley_plot_table_5_prime(files_5_prime)

```


```{r}
files_3_prime <- list.files(path = "./05-results/normalize_kraken/mapDamge2/methanobrevibacter_oralis", pattern = "3pGtoA_freq.txt", recursive = TRUE, full.names = T)

smiley_plot_table_3_prime <- function(files) {
  table <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(position = 1,
          frequency = 2) %>%
    mutate(Sequencing_ID = fn) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, "./05-results/normalize_kraken/mapDamge2/methanobrevibacter_oralis/results_", "")) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".mapped_m_oralis_q25l30_rmdup/3pGtoA_freq.txt", "")) %>%  
    inner_join(., metadata_wp %>%
              dplyr::select(Sequencing_ID, Strandedness, Individual_ID)) %>%
      filter(Strandedness == "Single")
  })
  
  table_all <- do.call(rbind, table) 
  
  return(table_all)
}

table_3_prime_ssLibs <- smiley_plot_table_3_prime(files_3_prime)

```


###Individual plots
```{r}
combined_5_prime <- table_5_prime_ssLibs %>%
  full_join(., table_5_prime_dsLibs) 

combined_3_prime <- table_3_prime_ssLibs %>%
  full_join(., table_3_prime_dsLibs)
```


```{r}
samples <- c("CDM019", "CDM020", "FUM002", "GDN001", "GOY005", "GOY006", "PES001", "VLD001")

individual_damage_plots(combined_5_prime, combined_3_prime, samples, "./06_publication/supplemental_figures/npost_species/m_oralis/damage_plots", "m_oralis")
```



