---
title: "Figure 2 - Sequencing depth, dataset complexity, contamination profiles, and read retention, across library preparation protocols"
author: "Keri Burge"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output:
    github_document:
        fig_width: 8
        fig_height: 6
        toc: true
---

```{r, eval=TRUE, echo=FALSE, message=FALSE}

library(knitr)
library(dplyr)
library(ggplot2)
library(rstatix)
library(viridis)
library(data.table)
library(ggpattern) 
library(janitor)
library(ggrepel)
library(tidyverse)
library(Nonpareil)

```


```{r setup}
knitr::opts_knit$set(root.dir = normalizePath(".."))
```


#A: Initial read count
```{r}
metadata <- fread("./05-results/no_normalization/dsvss_no_normalization_metadata.tsv")

double_color <- "#04387a"
single_color <- "#6fbfdc"
```


```{r}
median_initial_read_count <- metadata %>%
  select(Sequencing_ID, Strandedness, pre_eager_tot_read_pairs) %>%
  group_by(Strandedness) %>%
  summarise(
    median = median(pre_eager_tot_read_pairs, na.rm = TRUE),
    sd_overall = sd(pre_eager_tot_read_pairs, na.rm = TRUE)
  )


median_initial_gc <- metadata %>%
  select(Sequencing_ID, Strandedness, pre_eager_avg_gc)

median_initial_rl <- metadata %>%
  select(Sequencing_ID, Strandedness, pre_eager_avg_rl)
```


```{r}
pre_v_post_colors <- c(double_color, single_color)

reads_histogram <- metadata %>%
  dplyr::select(Library_ID, Individual_ID, Strandedness, pre_eager_tot_read_pairs) %>%
  mutate(initial_reads = pre_eager_tot_read_pairs) %>%
  mutate(initial_reads = initial_reads / 1000000) %>%
  ggplot(., aes(x=Strandedness, y=initial_reads, fill=Strandedness)) +
    geom_bar(stat="identity") +
    theme_minimal() +
    labs(title = "", x = "Individual ID & Strandedness", y = expression(Reads~(x10^6))) +
    guides(fill=guide_legend(title="")) +
    theme(legend.position="bottom") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    guides(fill = guide_legend(reverse=TRUE)) +
    scale_fill_discrete(name = "", type = pre_v_post_colors) +
    facet_grid(~Individual_ID, scales = "free_x", space = "free") +
    theme(strip.text.x = element_text(angle = 90, hjust = 1, size = 14),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 14),
          strip.text = element_text(size = 12),
          legend.text = element_text(size = 10),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14)) 


reads_histogram

#ggsave("./06-publication/main_figures/Figure2A_initial_read_count_plot.pdf", plot = reads_histogram, device = "pdf", scale = 1, width = 10, height = 6, units = c("in"), dpi = 300)
```

#B: Nopareil ()
Plot nonpareil curves for all individual IDs - Supplemental figures 1-2
```{r}
#use ./05-results/normalize_eager/npre_nonpareil_npo_file_paths.tsv for NPre data
#use ./05-results/no_normalization/NN_nonpareil_npo_file_paths.tsv for NN data
samples <- fread('./05-results/normalize_eager/npre_nonpareil_npo_file_paths.tsv',
                      sep = '\t', header = TRUE) %>%
  mutate(File = str_replace(File, "\\../../", "./"))
  

individual_ids <- c("BAN001", "CDM019", "CDM020", "FUM001", "FUM002", "FUM003",
                    "GDN001", "GOY005", "GOY006", "GUE001", "PES001", "SPM002", 
                    "VLD001", "ZAF001")

for (individual_id in individual_ids) {
  #Define output file name
  file_name <- file.path("./06-publication/supplemental_figures/", paste0(individual_id, "_nonpareil_curve_NPre.pdf"))
  
  #Start PNG device
 pdf(file = file_name, width = 5, height = 4)

  #Filter and plot
  sample_subset <- samples %>%
    filter(Individual_ID == individual_id) %>%
    dplyr::select(-Name) %>%
    rename(Name = Strandedness)

  Nonpareil.set(sample_subset$File,
                col = sample_subset$Col,
                labels = sample_subset$Name,
                plot.opts = list(main = "", plot.observed = FALSE))

  title(paste0(individual_id, " - 10M"))
  
  #Save file 
  dev.off()
}
```

Create a summary plot showing the predicted sequencing effort to reach 95% coverage 
```{r}
samples <- fread(
  "./05-results/normalize_eager/npre_nonpareil_npo_file_paths.tsv",
  sep = "\t",
  header = TRUE
) %>%
  mutate(File = str_replace(File, "\\../../", "./"))

individual_ids <- c("BAN001", "CDM019", "CDM020", "FUM001", "FUM002", "FUM003", "GDN001", "GOY005", "GOY006", "GUE001", "PES001", "SPM002", "VLD001", "ZAF001")

samples_subset <- samples %>%
  filter(Individual_ID %in% individual_ids) %>%
  mutate(Library = Strandedness,
         Plot_label = paste(Individual_ID, Library, sep = " - "))
```


```{r}
nps <- Nonpareil.set(
  files  = samples_subset$File,
  col    = samples_subset$Col,
  labels = samples_subset$Plot_label,
  plot   = FALSE,
  star   = 95)

np_summary <- as.data.frame(summary(nps)) %>%
  mutate(Effort95_Gbp = LRstar/1e9) %>%
  rownames_to_column() %>%
  separate(rowname, into = c("Individual_ID", "Strandedness"), sep = " - ") %>%
  mutate(Coverage = C*100)

```


```{r}
#Stats using rstatix package
wilcox_results <- np_summary %>%
  wilcox_test(Effort95_Gbp ~ Strandedness, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- np_summary %>%
  wilcox_effsize(formula = Effort95_Gbp ~ Strandedness, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat))

annotations <- wilcox_results %>%
  summarize(annotation = first(annotation),
            y_position = 95, .groups = 'drop')
```


```{r}
#Plot
nonpareil_95_cov <- np_summary %>%
  mutate(Strandedness = str_replace(Strandedness, "Double", "dsLib"),
         Strandedness = str_replace(Strandedness, "Single", "ssLib")) %>%
  ggplot(., aes(x = Strandedness, y = Effort95_Gbp)) + 
  geom_boxplot() +
  scale_y_log10() +
  geom_jitter(width = 0.06, size = 3, aes(color = Coverage)) +
  scale_color_viridis(option = "inferno", begin = 0.1) +
  theme_classic() +
  scale_shape_manual(values = c(16, 18)) +
  labs(x = "Library Type",
       y = "Predicted sequencing effort for 95% coverage (Gbp)",
       color = "Initial coverage (%)") +
  expand_limits(y = 100) +
  guides(alpha = FALSE) +
  geom_text(data = annotations, aes(x = 1.5, y = y_position, label = annotation), vjust = -0.5, size = 5) +
  theme(axis.text.x = element_text( size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
  geom_line(aes(group = Individual_ID), linetype = "dashed", color = "#555555", linewidth = 0.2) +
  geom_text_repel(aes(label = Individual_ID), size = 2)

nonpareil_95_cov

#ggsave("./06-publication/main_figures/Figure2B_NPre_nonpareil_95x_seq_effort_summary.pdf", plot = nonpareil_95_cov, device = "pdf", scale = 1, width = 8, height = 5, units = c("in"), dpi = 300)
```


#C: Sourcetracker normalizing 10M into nf-core/eager
```{r}
source_colors <- c(
  "Modern Calculus" = "#bfaed4",  
  "Plaque"          = "#3f007d",  
  "Gut"             = "#c05b16",  
  "Skin"            = "#fdc187",  
  "Environment"     = "#7fc97f",  
  "Unknown"         = "#d9d9d9"   
)

source_patterns <- c(
  "Unknown"         = "none",
  "Environment"     = "stripe",
  "Skin"            = "stripe",
  "Gut"             = "stripe",
  "Plaque"          = "none",
  "Modern Calculus" = "none"
)

source_angles <- c(
  "Unknown"         = 0,
  "Environment"     = 45,
  "Skin"            = 135,
  "Gut"             = 0,
  "Plaque"          = 0,
  "Modern Calculus" = 0
)

#Read in metadata 
metadata <- fread("./05-results/normalize_eager/dsvss_npre_metadata.tsv") %>%
  filter(!str_detect(Sequencing_ID, "bp")) 

#Prepare SourceTracker Results
source_results <- fread("./05-results/normalize_eager/dsvss_npre_sourcetracker_sink_predictions.txt") %>%
  rename(Sequencing_ID = 1) %>%
  full_join(., metadata)

sources_plot_no_subsampling_individual_facet <- source_results %>%
  rename(Environment = environment, Gut = gut, "Modern Calculus" = modern_calc, Plaque = plaque, Skin = skin) %>%
  gather("Sources", "Percent",2:7) %>%
  mutate(Percent = Percent * 100) %>%
  mutate(Sources = fct_relevel(Sources, "Unknown", "Environment", "Skin","Gut","Plaque","Modern Calculus")) %>%
  ggplot(., aes(x = Strandedness, y = Percent, fill = Sources, pattern = Sources)) +
          geom_bar_pattern(
            aes(pattern_angle = Sources),
            linewidth = .02,
            stat = "identity",
            color = "black",
            #pattern_alpha = 0.45,
            pattern_fill = NA,
            pattern_density = 0.03,
            pattern_spacing = 0.07,
            pattern_size = 0.15,
            pattern_key_scale_factor = 0.2) +
          theme_minimal(base_size = 14) +
          scale_fill_manual(values = source_colors) +
          scale_pattern_manual(values = source_patterns) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1),
                axis.text = element_text(size = 8)) +
          geom_hline(yintercept=50, linetype="dotted") +
          geom_hline(yintercept=90, linetype="dotted", color = "gray50") +
          #scale_y_continuous(breaks=scales::pretty_breaks(n=6)) +
          ylab("% attributed to source") +
          xlab("Individual ID and Strandedness") +
          facet_wrap(~Individual_ID, ncol = 7) +
          scale_pattern_angle_manual(values = source_angles)

sources_plot_no_subsampling_individual_facet

#ggsave("./06-publication/main_figures/Figure2C_NPre_no_rl_kraken2_sourcetracker_IndividualID_facet.pdf", plot = sources_plot_no_subsampling_individual_facet, device = "pdf", scale = 1, width = 12, height = 6, units = c("in"), dpi = 300)
```

Run a paired wilcoxon signed-rank test to compare the oral component (plaque, modern calculus) between library types
```{r}
oral_sources <- source_results %>%
  mutate(oral_source = plaque + modern_calc) %>%
  select(Sequencing_ID, Strandedness, oral_source)

wilcox_results <- oral_sources %>%
  wilcox_test(oral_source ~ Strandedness, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- oral_sources %>%
  wilcox_effsize(formula = oral_source ~ Strandedness, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat))
```


#D: Loss of reads after processing via nf-core/eager
```{r}
#Library info
metadata <- fread("./05-results/normalize_eager/dsvss_npre_w_rl_subsampling_metadata.tsv") %>%
  filter(!str_detect(Sequencing_ID, "bp")) %>%
  mutate(Sequencing_ID = str_replace(Sequencing_ID, "_10M", ""))

#Read counts throughout eager
read_counts <- fread("./05-results/normalize_eager/dsvss_npre_eager_read_counts.tsv")

#Combine tables
metadata <- metadata %>%
  full_join(., read_counts, by = c("Sequencing_ID", "Strandedness"))

eager <- metadata %>%
  pivot_longer(Adapter_removal_merged:Cutadapt_merged, names_to = "step", values_to = "reads") %>%
  mutate(step = (factor(step, levels = c("Adapter_removal_merged", "Remove_less_30bp_merged", "Artifact_removal_merged", "Cutadapt_merged")))) %>%
  mutate(step = str_replace_all(step, "Adapter_removal_merged", "Reads After Adapter Removal and Merging"),
         step = str_replace_all(step, "Remove_less_30bp_merged", "Reads After Removing Reads <30bp"),
         step = str_replace_all(step, "Artifact_removal_merged", "Reads After Removing Sequencing and Lab Artifacts"),
         step = str_replace_all(step, "Cutadapt_merged", "Reads After Cutadapt")) %>%
  rename(Library = Strandedness) %>%
  mutate(Library = str_replace_all(Library, "Single", "ssLib"),
         Library = str_replace_all(Library, "Double", "dsLib")) %>%
  mutate(reads = str_replace_all(reads, ",", "")) %>%
  mutate(reads = as.numeric(reads))

eager$step <- factor(eager$step, levels = c("Reads After Adapter Removal and Merging", "Reads After Removing Reads <30bp", "Reads After Removing Sequencing and Lab Artifacts", "Reads After Cutadapt")) 
```


```{r}
#Percent of original reads retained after nf-core/eager
eager_percent <- eager %>%
  pivot_wider(names_from = "step", values_from = "reads") %>%
  rename(percent = `%_retained`) 

eager_percent_average <- eager_percent %>%
  group_by(Library) %>%
  mutate(av_percent_retained = mean(percent)) %>%
  select(Library, av_percent_retained) %>%
  unique()

#Stats using rstatix package
wilcox_results <- eager_percent %>%
  wilcox_test(percent ~ Library, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- eager_percent %>%
  wilcox_effsize(formula = percent ~ Library, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat))

annotations <- wilcox_results %>%
  summarize(annotation = first(annotation),
            y_position = 95, .groups = 'drop')


kraken2_filtered_oral_species <- fread("./05-results/normalize_eager/dsvss_npre_no_rl_kraken2_gtdb207_oral_species_decontam_filtered.tsv") %>%
  adorn_totals("row") %>%
  filter(Species == "Total") %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "number_oral_reads") %>%
  mutate(number_oral_reads = as.numeric(number_oral_reads)) %>%
  dplyr::select(-Species)

number_species <- kraken2_filtered_oral_species %>%
  mutate(Sequencing_ID = str_replace_all(Sequencing_ID, "_10M", "")) %>%
  full_join(., metadata) %>%
  rename(Sample = Individual_ID, Library = Strandedness) %>%
  mutate(Library = str_replace(Library, "Double", "dsLib"),
         Library = str_replace(Library, "Single", "ssLib")) %>%
  mutate(percent_oral_reads = (number_oral_reads/Initial_reads)*100) 

#Plot
eager_percent_plots <- eager_percent %>%
  left_join(., number_species) %>%
  ggplot(., aes(x = Library, y = percent)) + 
  geom_boxplot() +
  geom_jitter(width = 0.075, size = 3, aes(color = percent_oral_reads)) +
  scale_color_viridis(option = "viridis", begin = 0.1) +
  theme_classic() +
  scale_shape_manual(values = c(16, 18)) +
  labs(x = "Library Type",
       y = "Percent of Original Reads Retained (%)",
       color = "Percent of Reads Classified as Oral") +
  expand_limits(y = 100) +
  guides(alpha = FALSE) +
  geom_text(data = annotations, aes(x = 1.5, y = y_position, label = annotation), vjust = -0.5, size = 5) +
  theme(axis.text.x = element_text( size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
  geom_line(aes(group = Individual_ID), linetype = "dashed", color = "#555555", linewidth = 0.2) +
  geom_text_repel(aes(label = Sample), size = 2)

eager_percent_plots

#ggsave("./06-publication/main_figures/Figure2D_NPre_percent_reads_retained_after_processing.pdf", plot = eager_percent_plots, device = "pdf", scale = 1, width = 8, height = 5, units = c("in"), dpi = 300)
```

