---
title: "Figure 4"
author: "Keri Burge"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output:
    github_document:
        fig_width: 8
        fig_height: 6
        toc: true
---

```{r, eval=TRUE, echo=FALSE, message=FALSE}
#Load in all necessary packages
library(knitr)
library(data.table)
library(ggplot2)
library(viridis)
library(ggpubr)
library(rstatix)
library(mixOmics)
library(ggrepel)
library(vegan)
library(janitor)
library(dplyr)
library(cuperdec)
library(colorspace)
library(tidyverse)
```


```{r setup}
knitr::opts_knit$set(root.dir = normalizePath(".."))
```


# Colors
```{r, colors}
double_color <- "#04387a"
single_color <- "#6fbfdc"
  
```


#Metadata
```{r}
metadata <- fread("./05-results/normalize_kraken/dsvss_npost_w_rl_subsampling_metadata.tsv") 

metadata_no_rl <-metadata %>%
  filter(Read_length_subsampling == "none")
```


#Oral Species List 
List of oral species in the format of the GTDB r207 database
```{r}
#GTDB Database
gtdb_sources <- fread("./00-documentation/dsvss_gtdb_r207_species_sources.tsv", header = TRUE) 

oral_gtdb <- gtdb_sources %>%
  mutate(Isolation_Source = str_replace_all(Isolation_Source, "oral", "TRUE"),
         Isolation_Source = str_replace_all(Isolation_Source, "other", "FALSE")) %>%
  rename(Taxon = 1) %>%
  mutate(Isolation_Source = as.logical(Isolation_Source)) %>%
  mutate(Taxon = str_replace(Taxon, " (?:[A-Z]+)$", "")) %>%
  unique() 

```


#Kraken2 species table
```{r}
kraken2_total_species <- fread("./05-results/normalize_kraken/dsvss_npost_kraken2_gtdb207_total_species_decontam.tsv") %>%
  mutate(Species = str_replace(Species, " (?:[A-Z]+)$", "")) %>%
  group_by(Species) %>%
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) %>%
  mutate(across(-Species, ~ if_else(.x < 10, 0, .x))) #only count a species as present if it has at least 10 reads

#filter for no read length subsampling
no_rl_samples <- metadata_no_rl %>%
  dplyr::select(Sequencing_ID) %>%
  pull()

kraken2_total_species_no_rl <- kraken2_total_species %>%
  column_to_rownames("Species") %>%
  dplyr::select(all_of(no_rl_samples)) %>%
  rownames_to_column("Species")

#filter for only oral species
oral_species <- oral_gtdb %>%
  filter(str_detect(Isolation_Source, "TRUE")) %>%
  rename(Species = 1) %>%
  dplyr::select(-Isolation_Source)

##with read length subsampling
oral_species_table <- kraken2_total_species %>%
  inner_join(., oral_species, by = "Species") 

##no read length subsampling
oral_species_table_no_rl <- kraken2_total_species_no_rl %>%
  inner_join(., oral_species, by = "Species") 


```


#A: Median read length of classified reads
```{r}
#color by the percentage of reads classified by kraken2 -> out of initial 5M, how many are classified?

#read in file with each the number of reads classified by kraken2 by read length (per library)
rl_classified <- fread("./05-results/normalize_kraken/dsvss_npost_classified_read_length.tsv") %>%
  mutate(Study = str_replace(Study, ".classified_gc_rl.tsv.gz", "")) %>%
  separate(Study, into = "Sequencing_ID", sep = ".tract", extra = "drop") %>%
  full_join(., metadata_no_rl) %>%
  dplyr::rename(counts_classified = counts_length)

rl_levels <- c(30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 73, 74, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99, 100, 101, 102, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 113, 114, 115, 116, 117, 118, 119, 120, 121, 122, 123, 124, 125, 126, 127, 128, 129, 130, 131, 132, 133, 134, 135, 136, 137, 138, 139, 140, 141, 142, 143, 144, 145, 146, 147, 148, 149, 150)

rl_summary_classified <- rl_classified %>%
    mutate(Length = factor(Length, level=rl_levels)) 

#function to calculate the median read length of oral reads per libraryID
median_rl_summary_table <- function(metadata, sample_list, oral_reads_length_table) {
  
  #Create a blank dataframe
  read_length_summary_table <- data.frame(Individual_ID = sample_list,
                                          Double = numeric(length(sample_list)),
                                          Single = numeric(length(sample_list)),
                                          stringsAsFactors = FALSE)
  
  #Function to calculate median read length based on dataset, sample, and strandedness
  calculate_median_rl <- function(data, sample, strandedness) {
    rl_summary <- data %>%
      filter(!str_detect(Library_ID, "bp")) %>%
      dplyr::select(Library_ID, Individual_ID, Strandedness, Length, counts_classified) %>%
      mutate(Length = as.integer(as.character(Length))) %>%
      filter(Individual_ID == sample, Strandedness == strandedness) %>%
      uncount(counts_classified) %>%
      mutate(Length = as.integer(Length)) %>%
      summarize(median_rl = median(Length))
    
    return(rl_summary$median_rl)
  }
  
  
  for (sample in sample_list) {
    #adapter removal
    rl_value_of_median_ds <- calculate_median_rl(oral_reads_length_table, sample, "Double")
    rl_value_of_median_ss <- calculate_median_rl(oral_reads_length_table, sample, "Single")
    read_length_summary_table[read_length_summary_table$Individual_ID == sample, "Double"] <- rl_value_of_median_ds
    read_length_summary_table[read_length_summary_table$Individual_ID == sample, "Single"] <- rl_value_of_median_ss
    
   }

  return(read_length_summary_table)
  
}

samples <- metadata_no_rl %>%
  dplyr::select(Individual_ID) %>%
  unique() %>%
  pull()

median_rl <- median_rl_summary_table(metadata_no_rl, samples, rl_summary_classified) %>%
  pivot_longer(-Individual_ID, values_to = "median_rl", names_to = "Strandedness")

#calculate the percent of reads classified (out of 5M reads going into kraken2)
kraken5M_classified_species_table <- kraken2_total_species_no_rl %>%
  column_to_rownames("Species") %>%
  dplyr::select(all_of(no_rl_samples)) %>%
  rownames_to_column("Species") %>%
  adorn_totals() %>%
  filter(Species == "Total") %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "classified_reads") %>%
  dplyr::select(-Species)

metadata_no_rl <- metadata_no_rl %>%
  full_join(., kraken5M_classified_species_table) %>%
  full_join(., median_rl) %>%
  mutate(percent_classified = (classified_reads/read_count)*100) %>%
  mutate(Library = Strandedness) %>%
  mutate(Library = str_replace_all(Library, "Double", "dsLib"),
         Library = str_replace_all(Library, "Single", "ssLib"))

```


```{r}
wilcox_results <- metadata_no_rl %>%
  wilcox_test(median_rl ~ Library, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- metadata_no_rl %>%
  wilcox_effsize(formula = median_rl ~ Library, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat))

annotations <- wilcox_results %>%
  summarize(annotation = first(annotation),
            y_position = 77, .groups = 'drop')

#Median across dataset
overal_median_rl <- metadata_no_rl %>%
  dplyr::select(Sequencing_ID, Strandedness, median_rl) %>%
  group_by(Strandedness) %>%
  summarize(
    median = median(median_rl, na.rm=TRUE),
    sd = sd(median_rl,na.rm=TRUE) 
  )


#Plot
read_length_median_percent_plot <- metadata_no_rl %>%
  ggplot(., aes(x = Library, y = median_rl)) + 
  geom_boxplot() +
  geom_jitter(width = 0.075, size = 3, aes(color = percent_classified)) +
  scale_color_viridis(
    option = "inferno",
    begin = 0.1,
    breaks = c(0, 10, 20, 30),
    limits = c(0, 30)) +
  theme_classic() +
  scale_shape_manual(values = c(16, 18)) +
  labs(x = "Library Type",
       y = "Median Read Length of Classified Reads",
       color = "Percent of Reads Classified by kraken2") +
  expand_limits(y = 80) +
  guides(alpha = "none") +
  geom_text(data = annotations, aes(x = 1.5, y = y_position, label = annotation), vjust = -0.5, size = 5) +
  theme(axis.text.x = element_text( size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
  geom_line(aes(group = Individual_ID), linetype = "dashed", color = "#555555", linewidth = 0.2) 
+
  geom_text_repel(aes(label = Individual_ID), size = 2)

read_length_median_percent_plot
```

Stats for % of reads (out of 5M) classified by Kraken2
```{r}
av_percent_classified <- metadata_no_rl %>%
  dplyr::select(Sequencing_ID, Strandedness, percent_classified) %>%
  group_by(Strandedness) %>%
  summarize(
    av_percent_class = mean(percent_classified, na.rm=TRUE),
    sd = sd(percent_classified,na.rm=TRUE) 
  )


wilcox_results_av_percent_class <- metadata_no_rl %>%
  wilcox_test(percent_classified ~ Library, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels_av_percent_class <- metadata_no_rl %>%
  wilcox_effsize(formula = percent_classified ~ Library, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))
```


#B: Median %GC of classified reads
```{r}
gc_levels <- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 73, 74, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99, 100, 101, 102, 103, 104)

#read in file with each the number of reads classified by kraken2 by %GC (per library)
gc_classified <- fread("./05-results/normalize_kraken/dsvss_npost_classified_gc.tsv") %>%
  mutate(Study = str_replace(Study, ".classified_gc_rl.tsv.gz", "")) %>%
  separate(Study, into = "Sequencing_ID", sep = ".tract", extra = "drop") %>%
  full_join(., metadata_no_rl) %>%
  dplyr::rename(counts_classified = counts_gc) %>%
  separate(gc_bin, into = c("drop", "gc"), sep = ",", remove = FALSE) %>%
  mutate(gc = str_remove(gc, "\\]")) %>%
  dplyr::select(-gc_bin, -drop) %>%
  mutate(gc = as.factor(gc))

gc_summary_classified <- gc_classified %>%
    mutate(gc = factor(gc, level=gc_levels)) 

#function to calculate the median read length of oral reads per libraryID
median_gc_summary_table <- function(metadata, sample_list, oral_reads_gc_table) {
  
  #Create a blank dataframe
  gc_summary_table <- data.frame(Individual_ID = sample_list,
                                          Double = numeric(length(sample_list)),
                                          Single = numeric(length(sample_list)),
                                          stringsAsFactors = FALSE)
  
  #Function to calculate median read length based on dataset, sample, and strandedness
  calculate_median_gc <- function(data, sample, strandedness) {
    gc_summary <- data %>%
      filter(!str_detect(Library_ID, "bp")) %>%
      dplyr::select(Library_ID, Individual_ID, Strandedness, gc, counts_classified) %>%
      filter(Individual_ID == sample, Strandedness == strandedness) %>%
      mutate(gc = as.numeric(as.character(gc))) %>%
      uncount(counts_classified) %>%
      summarize(median = median(gc))
    
    return(gc_summary$median)
  }
  
  
  for (sample in sample_list) {
    #adapter removal
    gc_value_of_median_ds <- calculate_median_gc(oral_reads_gc_table, sample, "Double")
    gc_value_of_median_ss <- calculate_median_gc(oral_reads_gc_table, sample, "Single")
    gc_summary_table[gc_summary_table$Individual_ID == sample, "Double"] <- gc_value_of_median_ds
    gc_summary_table[gc_summary_table$Individual_ID == sample, "Single"] <- gc_value_of_median_ss
    
   }

  return(gc_summary_table)
  
}

samples <- metadata_no_rl %>%
  dplyr::select(Individual_ID) %>%
  unique() %>%
  pull()

median_gc <- median_gc_summary_table(metadata_no_rl, samples, gc_summary_classified) %>%
  pivot_longer(-Individual_ID, values_to = "median_gc", names_to = "Strandedness")

metadata_no_rl <- metadata_no_rl %>%
  full_join(., median_gc)

```


```{r}
wilcox_results <- metadata_no_rl %>%
  wilcox_test(median_gc ~ Library, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- metadata_no_rl %>%
  wilcox_effsize(formula = median_gc ~ Library, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat))

annotations <- wilcox_results %>%
  summarize(annotation = first(annotation),
            y_position = 65, .groups = 'drop')

#Median across dataset
overal_median_gc <- metadata_no_rl %>%
  dplyr::select(Sequencing_ID, Strandedness, median_gc) %>%
  group_by(Strandedness) %>%
  summarise(
    median = median(median_gc, na.rm = TRUE),
    sd = sd(median_gc, na.rm = TRUE)
    )


#Plot
gc_median_percent_plot <- metadata_no_rl %>%
  ggplot(., aes(x = Library, y = median_gc)) + 
  geom_boxplot() +
  geom_jitter(width = 0.075, size = 3, aes(color = percent_classified)) +
  scale_color_viridis(
    option = "inferno",
    begin = 0.1,
    breaks = c(0, 10, 20, 30),
    limits = c(0, 30)) +
    theme_classic() +
  scale_shape_manual(values = c(16, 18)) +
  labs(x = "Library Type",
       y = "Median %GC of Classified Reads",
       color = "Percent of Reads Classified by kraken2") +
  expand_limits(y = 70) +
  guides(alpha = "none") +
  geom_text(data = annotations, aes(x = 1.5, y = y_position, label = annotation), vjust = -0.5, size = 5) +
  theme(axis.text.x = element_text( size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
  geom_line(aes(group = Individual_ID), linetype = "dashed", color = "#555555", linewidth = 0.2) 
+
  geom_text_repel(aes(label = Individual_ID), size = 2)

gc_median_percent_plot
```


```{r}
summary_plots <- ggarrange(
  read_length_median_percent_plot, gc_median_percent_plot,
  labels = c("A", "B"), 
  ncol = 2, nrow = 1,  common.legend = TRUE, legend = "bottom"
)

summary_plots

#ggsave("./06-publication/main_figures/Figure4AB_npost_classified_reads_gc_rl_boxplot.pdf", plot = summary_plots, device = "pdf", scale = 1, width = 10, height = 5, units = c("in"), dpi = 300)
```


#C: GOY005 read length distribution of classified reads
```{r}
rl_classified_GOY005 <- rl_classified %>%
  filter(Individual_ID == "GOY005") 

median_rl_info <- metadata_no_rl %>%
  filter(Individual_ID == "GOY005") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot_GOY005 <- rl_classified_GOY005 %>%
  mutate(total_counts = counts_classified / 10000) %>%
  mutate(Length = as.numeric(Length)) %>%
  ggplot(aes(x = Length, y = total_counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = Length, y = total_counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = Length, y = total_counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    labs(x = "Read Length (bp)", y = expression(`Classified Reads` ~ (x10^4))) + 
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100", "110", "120", "130", "140")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "bottom",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 5)) +
    #scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2), 
                       #labels = c("0", "0.5", "1", "1.5", "2")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot_GOY005

#ggsave("./06_publication/main_figures/Figure4C_npost_GOY005_classified_read_length_plot.pdf", plot = rl_plot_GOY005, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```

## All samples
```{r}
#function that creates a plot per individual ID and then arranges them all together using ggpubr (ggarange)
make_rl_plot <- function(individual_id, rl_classified, metadata_no_rl) {

  rl_df <- rl_classified %>%
    filter(Individual_ID == individual_id) %>%
    mutate(total_counts = counts_classified / 10000) %>%
    mutate(Length = as.numeric(Length))

  median_rl_info <- metadata_no_rl %>%
    filter(Individual_ID == individual_id) %>%
    select(Strandedness, median_rl)

  rl_plot <- ggplot(rl_df, aes(x = Length, y = total_counts,
                         fill = Strandedness, linetype = Strandedness)) +
    geom_bar(
      data = rl_df %>% filter(Strandedness == "Double"),
      aes(x = Length, y = total_counts),
      fill = "#04387a", alpha = 1, stat = "identity"
    ) +
    geom_bar(
      data = rl_df %>% filter(Strandedness == "Single"),
      aes(x = Length, y = total_counts),
      fill = "#6fbfdc", alpha = 0.6, stat = "identity"
    ) +
    theme_minimal() +
    labs(
      title = individual_id,
      x = "Read Length (bp)",
      y = expression(`Classified Reads` ~ (x10^4))
    ) +
    guides(
      fill = guide_legend(title = "Strandedness"),
      linetype = guide_legend(title = "Strandedness")
    ) +
    #scale_x_discrete(
    #  breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140),
    #  labels = as.character(c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140))
    #) +
    scale_x_continuous(
      limits = c(30, 150),
      breaks = seq(30, 140, by = 15),
      expand = c(0, 0)) +
    theme(
      legend.position = "bottom",
      axis.text.x = element_text(angle = 90, hjust = 1, size = 12),
      axis.text.y = element_text(size = 12),
      axis.title = element_text(size = 14),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 14),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")
    ) +
    #expand_limits(y = c(0, y_max)) +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red")

  rl_plot
}

#Run the function for all individual IDs and then arrange
all_samples <- rl_classified %>%
  distinct(Individual_ID) %>%
  pull(Individual_ID)

plot_list <- map(
  all_samples,
  ~ make_rl_plot(.x, rl_classified = rl_classified, metadata_no_rl = metadata_no_rl)
)

classified_rl_plots <- ggarrange(plotlist = plot_list, nrow = 2, ncol = 5, common.legend = TRUE, legend = "bottom")

classified_rl_plots

#ggsave("./06_publication/supplemental_figures/npost_classified_reads_rl_plots_all.pdf", plot = classified_rl_plots, device = "pdf", scale = 1, width = 15, height = 6, units = c("in"), dpi = 300)


```


#D: GOY005 GC content distribution of classified reads
```{r}
gc_classified_GOY005 <- gc_classified %>%
  filter(Individual_ID == "GOY005") 

median_gc_info <- metadata_no_rl %>%
  filter(Individual_ID == "GOY005") %>%
  dplyr::select(Strandedness, median_gc) 

gc_plot_GOY005 <- gc_classified_GOY005 %>%
  #mutate(gc = factor(gc, level=gc_levels)) %>%
  #mutate(gc = as.numeric(gc)) 
  mutate(gc = as.numeric(as.character(gc))) %>%
  mutate(total_counts = (counts_classified / 10000)) %>%
  ggplot(aes(x = gc, y = total_counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = gc, y = total_counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = gc, y = total_counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    labs(x = "%GC", y = expression(`Classified Reads` ~ (x10^4))) + 
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    #scale_x_discrete(breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100), 
    #                 labels = c("10", "20", "30", "40", "50", "60", "70", "80", "90", "100")) +
    scale_x_continuous(
      limits = c(0, 101),
      breaks = seq(10, 90, by = 10),
      expand = c(0, 0)) + 
    theme(legend.position = "bottom",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 15), 
          axis.text.y = element_text(size = 15), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 4)) +
    #scale_y_continuous(breaks = c(0, 2, 4, 6, 8), 
    #                   labels = c("0", "2", "4", "6", "8")) +
    geom_vline(data = median_gc_info, aes(xintercept = as.integer(median_gc), linetype = Strandedness), size = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
gc_plot_GOY005

#ggsave("./06_publication/main_figures/Figure4D_npost_GOY005_classified_gc_plot.pdf", plot = gc_plot_GOY005, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```

## All samples
```{r}
#function that creates a plot per individual ID and then arranges them all together using ggpubr (ggarange)
make_gc_plot <- function(individual_id, gc_classified, metadata_no_rl) {

  gc_df <- gc_classified %>%
    filter(Individual_ID == individual_id) %>%
    mutate(total_counts = counts_classified / 10000) %>%
    #mutate(gc = factor(gc, level=gc_levels)) 
    #mutate(gc = as.numeric(gc))
    mutate(gc = as.numeric(as.character(gc))) 

  median_gc_info <- metadata_no_rl %>%
    filter(Individual_ID == individual_id) %>%
    select(Strandedness, median_gc)

  gc_plot <- ggplot(gc_df, aes(x = gc, y = total_counts,
                         fill = Strandedness, linetype = Strandedness)) +
    geom_bar(
      data = gc_df %>% filter(Strandedness == "Double"),
      aes(x = gc, y = total_counts),
      fill = "#04387a", alpha = 1, stat = "identity"
    ) +
    geom_bar(
      data = gc_df %>% filter(Strandedness == "Single"),
      aes(x = gc, y = total_counts),
      fill = "#6fbfdc", alpha = 0.6, stat = "identity"
    ) +
    theme_minimal() +
    labs(
      title = individual_id,
      x = "%GC",
      y = expression(`Classified Reads` ~ (x10^4))
    ) +
    guides(
      fill = guide_legend(title = "Strandedness"),
      linetype = guide_legend(title = "Strandedness")
    ) +
    #scale_x_discrete(
    #  breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140),
    #  labels = as.character(c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140))
    #) +
    scale_x_continuous(
      limits = c(0, 101),
      breaks = seq(10, 90, by = 10),
      expand = c(0, 0)) +
    theme(
      legend.position = "bottom",
      axis.text.x = element_text(angle = 90, hjust = 1, size = 12),
      axis.text.y = element_text(size = 12),
      axis.title = element_text(size = 14),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 14),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")
    ) +
    #expand_limits(y = c(0, y_max)) +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) +
    geom_vline(data = median_gc_info, aes(xintercept = median_gc, linetype = Strandedness), linewidth = 0.5, color = "red")

  gc_plot
}

#Run the function for all individual IDs and then arrange
all_samples <- gc_classified %>%
  distinct(Individual_ID) %>%
  pull(Individual_ID)

plot_list <- map(
  all_samples,
  ~ make_gc_plot(.x, gc_classified = gc_classified, metadata_no_rl = metadata_no_rl)
)

classified_gc_plots <- ggarrange(plotlist = plot_list, nrow = 2, ncol = 5, common.legend = TRUE, legend = "bottom")

classified_gc_plots

#ggsave("./06_publication/supplemental_figures/npost_classified_reads_gc_plots_all.pdf", plot = classified_gc_plots, device = "pdf", scale = 1, width = 15, height = 6, units = c("in"), dpi = 300)


```


#E: Count of classified reads total observed species
```{r}
#Kraken2 Total Species Table

#Difference in total number of species
kraken2_species_cuperdec_no_rl <- load_taxa_table(kraken2_total_species_no_rl)

#Calculate number of species reported
kraken2_total_species_table_no_rl <- kraken2_species_cuperdec_no_rl %>%
  group_by(Sample) %>%
  summarize(count_distinct_kraken2 = n_distinct(Taxon)) %>% #count how many unique taxa are reported in each library
  rename(Sequencing_ID = 1) %>%
  full_join(., metadata_no_rl) %>%
  drop_na()
```

```{r}
#Calculate the number of classified reads to add to the metadata table
classified_reads_no_rl <- kraken2_total_species_no_rl %>%
  adorn_totals() %>%
  filter(Species == "Total") %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "classified_reads") %>%
  dplyr::select(-Species)

classified_reads_table_no_rl <- classified_reads_no_rl %>%
  full_join(metadata_no_rl %>%
              dplyr::select(Sequencing_ID, Strandedness, read_count, kraken2_preservation)) %>%
  rename(total_reads = read_count,
         preservation = kraken2_preservation)

oral_reads_no_rl <- oral_species_table_no_rl %>%
  adorn_totals() %>%
  filter(Species == "Total") %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "oral_reads") %>%
  dplyr::select(-Species)

classified_reads_table_no_rl <- classified_reads_table_no_rl %>%
  full_join(., oral_reads_no_rl) %>%
  dplyr::select(Sequencing_ID, Strandedness, preservation, total_reads, classified_reads, oral_reads)

norm_kraken_percentages <- classified_reads_table_no_rl %>%
  mutate(percent_oral_of_total = (oral_reads/total_reads)*100) %>%
  mutate(percent_oral_of_class = (oral_reads/classified_reads)*100)
```

```{r}
kraken2_oral_species_table_no_rl <- load_taxa_table(oral_species_table_no_rl)

kraken2_species_summary_table_no_rl <- kraken2_oral_species_table_no_rl %>%
  group_by(Sample) %>%
  summarize(count_distinct_kraken2_filtered = n_distinct(Taxon)) %>%
  dplyr::rename(Sequencing_ID = 1) %>%
  full_join(., metadata_no_rl) %>%
  drop_na() %>%
  full_join(., kraken2_total_species_table_no_rl) %>%
  dplyr::rename(Total = count_distinct_kraken2,
         Oral = count_distinct_kraken2_filtered) %>%
  relocate(Total, .after = Oral) %>%
  pivot_longer(Oral:Total, names_to = "pre_v_post", values_to = "count")
```


Raincloud plot for total observed species (https://www.cedricscherer.com/2021/06/06/visualizing-distributions-with-raincloud-plots-and-how-to-create-them-with-ggplot2/)
```{r}
kraken2_species_summary_table_no_rl <- kraken2_species_summary_table_no_rl %>%
  mutate(pre_v_post = str_replace(pre_v_post, "Total", "Total Species"),
         pre_v_post = str_replace(pre_v_post, "Oral", "Oral Species")) 

#Statistics - Total v. Oral (ds % ss)
stats_kraken_alpha_diversity_wilcox_strandedness_no_rl <- kraken2_species_summary_table_no_rl %>%
  group_by(Strandedness) %>%
  wilcox_test(count ~ (pre_v_post), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

stats_kraken_alpha_diversity_effectsize_strandedness_no_rl <- kraken2_species_summary_table_no_rl %>%
  group_by(Strandedness) %>%
  wilcox_effsize(formula = count ~ pre_v_post, paired = TRUE) 

#Stats for plot
wilcox_results <- kraken2_species_summary_table_no_rl %>%
  group_by(forcats::fct_rev(pre_v_post)) %>%
  wilcox_test(count ~ Strandedness, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- kraken2_species_summary_table_no_rl %>%
  group_by(forcats::fct_rev(pre_v_post)) %>%
  wilcox_effsize(formula = count ~ Strandedness, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat)) %>%
  dplyr::rename(pre_v_post = 1)

annotations <- wilcox_results %>%
  group_by(pre_v_post) %>%
  summarize(annotation = dplyr::first(annotation),.groups = 'drop') 



#PLOT
color <- c(double_color, single_color)

kraken2_raincloud_plot_no_rl <- kraken2_species_summary_table_no_rl %>%
  ggplot(., aes(x=Strandedness, y=count)) +
  ggdist::stat_halfeye(
     aes(color = Strandedness,
        fill = after_scale(lighten(color, .2))),
    adjust = 0.5, 
    width = 0.6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
    ) + 
  ggdist::stat_dots(
    side = "left", 
    dotsize = 1.5, 
    justification = 1.175, 
    binwidth =10,
    alpha = .3,
    range_scale = .4,
    aes(color = Strandedness, color = after_scale(darken(color, .1, space = "HLS")))
  ) +
  geom_boxplot(
    aes(color = Strandedness,
        color = after_scale(darken(color, .1, space = "HLS")),
        fill = after_scale(desaturate(lighten(color, .8), .4))),
    width = .15, 
    outlier.shape = NA
    ) +
  facet_grid(~forcats::fct_rev(pre_v_post)) +
  theme_minimal() +
  theme(
    legend.position = "none", 
    plot.title = element_text(hjust = 0.5)
    ) +
  labs(
    title = "", 
    y = "Observed Species", 
    x = ""
    ) +
  coord_cartesian(xlim = c(1.2, NA), clip = "off") +
  scale_color_manual(values = color, guide = "none") +
  scale_fill_manual(values = color, guide = "none") +
  theme(axis.text.x = element_text(size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
            strip.text = element_text(size = 14)) +
  expand_limits(y = c(0, 450)) +
  geom_text(data = annotations, aes(x = 1.5, y = 435, label = annotation), vjust = -0.5)

kraken2_raincloud_plot_no_rl

#ggsave("./06_publication/main_figures/Figure4E_npost_no_rl_total_observed_species_raincloud_plot.pdf", plot = kraken2_raincloud_plot_no_rl, device = "pdf", scale = 1, width = 8, height = 5, units = c("in"), dpi = 300)
```

Average number of total species classified
```{r}
av_total_class <- kraken2_species_summary_table_no_rl %>%
  pivot_wider(names_from = "pre_v_post", values_from = "count") %>%
  dplyr::select(Sequencing_ID, Strandedness, `Total Species`) %>%
  rename(total = `Total Species`) %>%
  group_by(Strandedness) %>%
  summarize(
    average_totsl_class = mean(total, na.rm=TRUE),
    sd = sd(total, na.rm=TRUE)
  )

```


#F: Classified reads Shannon diversity
Shannon Diversity plot
```{r}
#Calculate the Shannon diversity index
#Total Species
kraken2_total_species_formatted_no_rl <- kraken2_total_species_no_rl %>%
  pivot_longer(2:ncol(.), names_to = "Sample", values_to = "Counts") %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  column_to_rownames("Sample")

shannon_diversity_total_no_rl <- diversity(kraken2_total_species_formatted_no_rl, index="shannon") %>%
  as.data.frame() %>%
  dplyr::rename(shannon_index_total = 1) %>%
  rownames_to_column("Sequencing_ID")

#Oral Species
#Need to reformat species table so species are the columns and samples are the rows
kraken2_oral_species_formatted_no_rl <- oral_species_table_no_rl %>%
  pivot_longer(2:ncol(.), names_to = "Sample", values_to = "Counts") %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  column_to_rownames("Sample")

shannon_diversity_oral_no_rl <- diversity(kraken2_oral_species_formatted_no_rl, index="shannon") %>%
  as.data.frame() %>%
  dplyr::rename(shannon_index_oral = 1) %>%
  rownames_to_column("Sequencing_ID")

#Combine tables
metadata_w_shannon_no_rl <- metadata_no_rl %>%
  full_join(., shannon_diversity_total_no_rl, by = "Sequencing_ID") %>%
  full_join(., shannon_diversity_oral_no_rl, by = "Sequencing_ID") %>%
  pivot_longer(shannon_index_total:shannon_index_oral, names_to = "total_v_oral", values_to = "shannon_index") 

#Statistics - Total v. Oral (ds + ss)
stats_kraken_alpha_diversity_wilcox_strandedness_no_rl <- metadata_w_shannon_no_rl %>%
  group_by(Strandedness) %>%
  wilcox_test(shannon_index ~ (total_v_oral), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

stats_kraken_alpha_diversity_effectsize_strandedness_no_rl <- metadata_w_shannon_no_rl %>%
  group_by(Strandedness) %>%
  wilcox_effsize(formula = shannon_index ~ total_v_oral, paired = TRUE) 

#Stats for plot
wilcox_results <- metadata_w_shannon_no_rl %>%
  group_by(total_v_oral) %>%
  wilcox_test(shannon_index ~ Strandedness, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- metadata_w_shannon_no_rl %>%
  group_by(total_v_oral) %>%
  wilcox_effsize(formula = shannon_index ~ Strandedness, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat))

annotations <- wilcox_results %>%
  group_by(total_v_oral) %>%
  summarize(annotation = dplyr::first(annotation),
            y_position = 5, .groups = 'drop') %>%
  mutate(total_v_oral = str_replace(total_v_oral, "shannon_index_oral", "Oral Species"),
         total_v_oral = str_replace(total_v_oral, "shannon_index_total", "Total Species"))

#Plot
kraken2_raincloud_plot_shannon_no_rl <- metadata_w_shannon_no_rl %>%
  mutate(total_v_oral = str_replace_all(total_v_oral, "shannon_index_oral", "Oral Species"),
         total_v_oral = str_replace_all(total_v_oral, "shannon_index_total", "Total Species")) %>%
  ggplot(., aes(x=Strandedness, y=shannon_index)) +
  ggdist::stat_halfeye(
     aes(color = Strandedness,
        fill = after_scale(lighten(color, .2))),
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
    ) + 
  ggdist::stat_dots(
    side = "left", 
    dotsize = 0.8, 
    justification = 1.175, 
    binwidth = 0.2,
    alpha = .3,
    range_scale = .4,
    aes(color = Strandedness, color = after_scale(darken(color, .1, space = "HLS")))
  ) +
  geom_boxplot(
    aes(color = Strandedness,
        color = after_scale(darken(color, .1, space = "HLS")),
        fill = after_scale(desaturate(lighten(color, .8), .4))),
    width = .15, 
    outlier.shape = NA
    ) +
  facet_grid(~forcats::fct_rev(total_v_oral)) +
  theme_minimal() +
  theme(
    legend.position = "none", 
    plot.title = element_text(hjust = 0.5)
    ) +
  labs(
    title = "", 
    y = "Shannon Index", 
    x = ""
    ) +
  coord_cartesian(xlim = c(1.2, NA), clip = "off") +
  scale_color_manual(values = color, guide = "none") +
  scale_fill_manual(values = color, guide = "none") +
  theme(axis.text.x = element_text(size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
            strip.text = element_text(size = 14)) +
  #stat_compare_means(method = "wilcox.test", paired = TRUE, label = "p.signif", label.x = 1.5, label.y = 4.6, size = 4, hide.ns = TRUE) +
  expand_limits(y = c(0, 5)) +
  geom_text(data = annotations, aes(x = 1.5, y = 4.7, label = annotation), vjust = -0.5)

kraken2_raincloud_plot_shannon_no_rl

#ggsave("./06_publication/main_figures/Figure4F_npost_no_rl_shannon_diversity_raincloud_plot.pdf", plot = kraken2_raincloud_plot_shannon_no_rl, device = "pdf", scale = 1, width = 8, height = 5, units = c("in"), dpi = 300)
```

Average shannon index of  total species classified
```{r}
av_shannon_class <- metadata_w_shannon_no_rl %>%  pivot_wider(names_from = "total_v_oral", values_from = "shannon_index")%>%
  dplyr::select(Sequencing_ID, Strandedness, shannon_index_total)%>%
  group_by(Strandedness) %>%
  summarize(
    av_shannon_total = mean(shannon_index_total, na.rm=TRUE),
    sd = sd(shannon_index_total, na.rm=TRUE)
  )

```


Combine the Observed species and Shannon diversity plots
```{r}
alpha_diversity_combined_no_rl <- ggarrange(
  kraken2_raincloud_plot_no_rl, kraken2_raincloud_plot_shannon_no_rl, 
  labels = c("A", "B"), 
  ncol = 1, nrow = 2, 
  common.legend = TRUE, 
  legend = "bottom"
)

alpha_diversity_combined_no_rl

#ggsave("./06_publication/main_figures/Figure4EF_npost_alpha_diversity_plots_combined.pdf", plot = alpha_diversity_combined_no_rl, device = "pdf", scale = 1, width = 8, height = 8, units = c("in"), dpi = 300)
```


#G: PCA of classified reads 
Create a PCA of total number of species
```{r}
kraken2_total_species_no_rl_edited <- kraken2_total_species_no_rl %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "Counts") %>%
  mutate(Counts = Counts + 1) %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  filter(!str_detect(Sequencing_ID, "Isolation_Source")) %>%
  column_to_rownames("Sequencing_ID")
  
result.pca.kraken2_no_rl <- pca(kraken2_total_species_no_rl_edited, logratio = "CLR") 

exp_var <- paste0(round(result.pca.kraken2_no_rl$prop_expl_var$X * 100, 2), "%") # explained variance for x- and y-labels

plotIndiv(result.pca.kraken2_no_rl)

pca_kraken2_no_rl <- result.pca.kraken2_no_rl$variates %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  dplyr::rename(Sequencing_ID = 1,
         PC1 = 2,
         PC2 = 3) %>%
  inner_join(., metadata_no_rl, by = "Sequencing_ID")

permanova_dist_matrix <- dist(pca_kraken2_no_rl[, c("PC1", "PC2")], method = "euclidean")
adonis_result <- adonis2(permanova_dist_matrix ~ Strandedness, data = pca_kraken2_no_rl) #Strandedness, Individual_ID, kraken2_preservation
r_squared_value <- adonis_result$R2[1] 
p_value <- adonis_result$`Pr(>F)`[1]
p_value_formatted <- ifelse(p_value < 0.001, "< 0.001",
                            as.character(signif(p_value, digits = 2)))
r_squared_value_formatted <- ifelse(r_squared_value < 0.001, "< 0.001",
                            as.character(signif(r_squared_value, digits = 2)))


#keep color for strandedness, and change shape for preservation 
color <- c(double_color, single_color)


kraken2_pca_strandedness_no_rl <- pca_kraken2_no_rl %>%
  mutate(kraken2_preservation = factor(kraken2_preservation, levels = c("Good", "Bad"))) %>%
  ggplot(., aes(x = PC1, y = PC2, color = Strandedness)) +
    geom_point(aes(shape = kraken2_preservation), size = 4) +
    theme_minimal() +
    labs(title = "", y = paste("PC2", " - ", exp_var[2]), x = paste("PC1", " - ", exp_var[1])) +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_color_discrete(name = "Strandedness", labels = c("Double", "Single")) +
    scale_shape_discrete(name = "Preservation", labels = c("Good", "Bad")) +
    geom_line(aes(group = Individual_ID), color = "#555555", linetype = "dashed") +
  scale_colour_manual(name = "Strandedness", values = color) +
  theme(axis.text.x = element_text(size = 12),  
        axis.text.y = element_text(size = 12),
        strip.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        legend.position = "right") +
        geom_text_repel(data = subset(pca_kraken2_no_rl, Strandedness == "Double"), 
                        aes(label = Individual_ID), 
                        size = 3, color = "black", segment.color = NA,
                        nudge_y = -0.8) +
        coord_equal()
+
    annotate("text", x = 25, y = 75,
           label = paste("R2 =", r_squared_value_formatted,
                         "\np-value =", p_value_formatted))
 
kraken2_pca_strandedness_no_rl

#ggsave("./06_publication/main_figures/Figure4G_npost_no_rl_pca_total_species_all_samples.pdf", plot = kraken2_pca_strandedness_no_rl, device = "pdf", scale = 1, width = 6, height = 4, units = c("in"), dpi = 300)


```
