---
title: "Figure 5 - Oral species–specific drivers of PCA structure and library-dependent recovery of taxa"
author: "Keri Burge"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output:
    github_document:
        fig_width: 8
        fig_height: 6
        toc: true
---


```{r, eval=TRUE, echo=FALSE, message=FALSE}
#Load in all necessary packages
library(knitr)
library(data.table)
library(ggplot2)
library(viridis)
library(rstatix)
library(mixOmics)
library(janitor)
library(dplyr)
library(vegan)
library(ggrepel)
library(ggpubr)
library(tidyverse)

```


```{r setup}
knitr::opts_knit$set(root.dir = normalizePath(".."))
```


# Colors
```{r, colors}
double_color <- "#04387a"
single_color <- "#6fbfdc"
  
```


#Well-preserved oral PCA -> What species are driving both PC1 and PC2 in the negative and positive direction?
```{r}
kraken2_oral_species_no_rl <- fread("./05-results/normalize_kraken/dsvss_npost_kraken2_gtdb207_oral_species_decontam_filtered.tsv", header = TRUE) 

data <- kraken2_oral_species_no_rl #select the oral set to use

metadata <- fread("./05-results/normalize_kraken/dsvss_npost_w_rl_subsampling_metadata.tsv") 

metadata_no_rl <- metadata %>%
  filter(Read_length_subsampling == "none")
```


```{r}
#Filter out the poorly preserved samples (BAN001 and FUM003)
kraken2_oral_species_no_rl_filtered <- data %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "Counts") %>%
  filter(!str_detect(Sequencing_ID, "BAN001")) %>%
  filter(!str_detect(Sequencing_ID, "FUM003")) %>%
  mutate(Counts = Counts + 1) %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  filter(!str_detect(Sequencing_ID, "Isolation_Source")) %>%
  column_to_rownames("Sequencing_ID")
  
result.pca.kraken2_no_rl_filtered <- pca(kraken2_oral_species_no_rl_filtered, logratio = "CLR") 

exp_var <- paste0(round(result.pca.kraken2_no_rl_filtered$prop_expl_var$X * 100, 2), "%") # explained variance for x- and y-labels

pca_kraken2_no_rl_filtered <- result.pca.kraken2_no_rl_filtered$variates %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  dplyr::rename(Sequencing_ID = 1,
         PC1 = 2,
         PC2 = 3) %>%
  inner_join(., metadata_no_rl, by = "Sequencing_ID")


```

Perform a PERMANOVA (Permutational Multivariate Analysis of Variance) test using the adonis2 function from the vegan package
```{r}
#Create a metadata file with sample IDs, grouping info, and PCA scores
permanova_data_no_rl_filtered <- pca_kraken2_no_rl_filtered %>%
  dplyr::select(Sequencing_ID, Individual_ID, PC1, PC2, Strandedness, kraken2_preservation) %>%
  column_to_rownames("Sequencing_ID")

#Create a distance matrix of just PCA scores
permanova_dist_matrix_no_rl_filtered <- dist(permanova_data_no_rl_filtered[, c("PC1", "PC2")], method = "euclidean")

#Run adonis2
adonis_result <- adonis2(permanova_dist_matrix_no_rl_filtered ~ Strandedness, data = permanova_data_no_rl_filtered) #Strandedness, Individual_ID

r_squared_value <- adonis_result$R2[1] 
  
p_value <- adonis_result$`Pr(>F)`[1]

p_value_formatted <- ifelse(p_value < 0.001, "< 0.001",
                            as.character(signif(p_value, digits = 2)))

r_squared_value_formatted <- ifelse(r_squared_value < 0.001, "< 0.001",
                            as.character(signif(r_squared_value, digits = 2)))
```


```{r}
#keep color for strandedness, and change shape for preservation 
color <- c(double_color, single_color)

kraken2_pca_strandedness_no_rl_filtered <- pca_kraken2_no_rl_filtered %>%
  mutate(kraken2_preservation = factor(kraken2_preservation, levels = c("Good", "Bad"))) %>%
  ggplot(., aes(x = PC1, y = PC2, color = Strandedness)) +
    geom_point(aes(shape = kraken2_preservation), size = 4) +
    theme_minimal() +
    labs(title = "", y = paste("PC2", " - ", exp_var[2]), x = paste("PC1", " - ", exp_var[1])) +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_color_discrete(name = "Strandedness", labels = c("Double", "Single")) +
    scale_shape_discrete(name = "Preservation") +
    geom_line(aes(group = Individual_ID), color = "#555555ff", linetype = "dashed") +
    #scale_colour_manual(name = "Strandedness", values = c("Double" = after_scale(lighten("#4059AD", .2)), "Single" = after_scale(lighten("#5fc3a4", .2)))) +
    #scale_color_manual(name = "Strandedness", values = c("Double" = "#098BD9", "Single" = "#7DCCFF")) +
    scale_color_manual(name = "Strandedness", values = color) +
    theme(axis.text.x = element_text(size = 12),  
          axis.text.y = element_text(size = 12),
          strip.text = element_text(size = 12),
          legend.text = element_text(size = 12),
          legend.title = element_text(size = 12),
          axis.title.x = element_text(size = 12),
          axis.title.y = element_text(size = 12),
          legend.position = "bottom") +
    geom_text_repel(data = subset(pca_kraken2_no_rl_filtered, Strandedness == "Double"), 
                        aes(label = Individual_ID), 
                        size = 3, color = "black", segment.color = NA) +
      coord_equal() +
  annotate("text", x = -14, y = 15,
           label = paste("R2 =", r_squared_value_formatted,
                         "\np-value =", p_value_formatted)) 
 
kraken2_pca_strandedness_no_rl_filtered

```

#Pull the reads and read lengths of just those species of interest (those driving the PCs)
```{r}
kraken2_oral_species_no_rl_filtered <- data %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "Counts") %>%
  filter(!str_detect(Sequencing_ID, "BAN001")) %>%
  filter(!str_detect(Sequencing_ID, "FUM003")) %>%
  mutate(Counts = Counts + 1) %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  filter(!str_detect(Sequencing_ID, "Isolation_Source")) %>%
  column_to_rownames("Sequencing_ID")

#Create PCA using mixomics package
result.pca.kraken2_no_rl_filtered <- pca(kraken2_oral_species_no_rl_filtered, ncomp = 2, logratio = "CLR") 

#Metadata table (will be added w/in the function)
full_metadata <- fread("./05-results/normalize_kraken/dsvss_npost_w_rl_subsampling_metadata.tsv") %>%
  filter(Read_length_subsampling == "none") %>%
  dplyr::select(Sequencing_ID, Strandedness, kraken2_preservation, Individual_ID)

#Define colors and shapes for plotting
Strandedness_colors <- c("Double" = double_color, "Single" = single_color)
kraken2_preservation_shapes <- c("Good" = 16, "Bad" = 17)

#Biplot Function
plot_pca_bi <- function(df, pc1, pc2, metadata_group1, metadata_group2, columntitle) {
    metadata_group_colors <- get(paste(metadata_group1, "_colors", sep = ""))
    metadata_group_shapes <- get(paste(metadata_group2, "_shapes", sep = ""))
   
    arrow_pc <- enquo(columntitle)
    
    exp_var <- paste0(round(df$prop_expl_var$X * 100, 2), "%") # explained variance for x- and y-labels
    
    # select only the PCs from the PCA and add metadata
    df_X <- df$variates$X %>%
              as.data.frame() %>%
              rownames_to_column("Sequencing_ID") %>%
              inner_join(full_metadata, by = "Sequencing_ID")
    
    metadata_group1 = df_X[[metadata_group1]]
    metadata_group2 = df_X[[metadata_group2]]
    
    corr_lam <- df$sdev[c("PC1", "PC2")] * sqrt(nrow(df_X))

    df_X <- df_X %>%
      mutate(PC1 = PC1 / corr_lam[1],
             PC2 = PC2 / corr_lam[2])

    # select the correct PC column and explained variance for PC1
    if (pc1 == 'PC1') {
        Pc1 <- df_X$PC1
        exp_var_pc1 <- exp_var[1]
        xaxis <- c("PC1")
    } else if (pc1 == 'PC2') {
        Pc1 <- df_X$PC2
        exp_var_pc1 <- exp_var[2]
        xaxis <- c("PC2")
    } 
    
    # select the correct PC column and explained variance for PC2
    if (pc2 == 'PC1') {
        Pc2 <- df_X$PC1
        exp_var_pc2 <- exp_var[1]
        yaxis <- c("PC1")
 } else if (pc2 == 'PC2') {
       Pc2 <- df_X$PC2
       exp_var_pc2 <- exp_var[2]
       yaxis <- c("PC2")
    }
    
    # Identify the 10 pathways that have highest positive and negative loadings in the selected PC
    pws_10 <- df$loadings$X %>%
      as.data.frame(.) %>%
      rownames_to_column(var = "Taxon") %>%
      top_n(10, !!arrow_pc) %>%
      arrange(desc(!!arrow_pc)) %>%   #sort by size of loading to get strongest first
      mutate(label = row_number())

    neg_10 <- df$loadings$X %>%
      as.data.frame(.) %>%
      rownames_to_column(var = "Taxon") %>%
      top_n(-10, !!arrow_pc) %>%
      arrange(!!arrow_pc) %>%  #sort to get the smallest (most negative/strongest) first
      mutate(label = row_number())

    pca_plot_bi <- df_X %>%
      mutate(metadata_group2 = factor(metadata_group2, levels = c("Good", "Bad"))) %>%
      ggplot(., aes(x = Pc1, y = Pc2, color = metadata_group1)) +
      geom_point(size = 4, aes(shape = metadata_group2) )+
      geom_segment(data = pws_10,
                   aes(xend = get(paste(pc1)), yend = get(paste(pc2))),
                   x = 0, y = 0, colour = "black",
                   size = 0.3,
                   arrow = arrow(length = unit(0.03, "npc"))) +
      #geom_label_repel(data = pws_10,
      #             aes(x = get(paste(pc1)), y = get(paste(pc2)), label = label), #using label instead of taxon
      #             size = 2, 
      #             colour = "black",
      #             label.padding = 0.1, 
      #             box.padding = 0.1, #
      #             point.padding = 0.1, 
      #             force = 10,
      #             max.overlaps = Inf) + 
      geom_segment(data = neg_10,
                   aes(xend = get(paste(pc1)), yend = get(paste(pc2))),
                   x = 0, y = 0, colour = "grey",
                   size = 0.3,
                   arrow = arrow(length = unit(0.03, "npc"))) +
      #geom_label_repel(data = neg_10,
      #             aes(x = get(paste(pc1)), y = get(paste(pc2)), label = label), #using label instead of taxon
      #             size = 2, 
      #             colour = "red",
      #             label.padding = 0.1, 
      #             box.padding = 0.1, 
      #             point.padding = 0.1, 
      #             force = 10, 
      #             max.overlaps = Inf) + 
      labs(x = paste(xaxis, " - ", exp_var_pc1),
           y = paste(yaxis, " - ", exp_var_pc2),
           color = "Strandedness",
           shape = "Preservation") +
      scale_color_manual(values = metadata_group_colors) +
      scale_shape_manual(values = metadata_group_shapes) +
      theme_minimal() + theme(text = element_text(size = 16)) +
      theme(text = element_text(size=16)) +
      theme(legend.position = "bottom") +
      geom_line(aes(group = Individual_ID), color = "#CCCCCC", linetype = "dashed") +
      geom_text_repel(aes(label = Individual_ID), color = "black", segment.color = NA) +
      coord_equal()
      

    return(pca_plot_bi) #pca_plot_bi, pws_10, neg_10 - CHANGE THIS VARIABLE FOR PLOT OR TABLE (pws_10, neg_10)
}

pca_kraken2_biplot_PC1_filtered <- plot_pca_bi(result.pca.kraken2_no_rl_filtered, "PC1", "PC2", "Strandedness", "kraken2_preservation", PC1)
pca_kraken2_biplot_PC2_filtered <- plot_pca_bi(result.pca.kraken2_no_rl_filtered, "PC1", "PC2", "Strandedness", "kraken2_preservation", PC2)

pca_kraken2_biplot_PC1_filtered
pca_kraken2_biplot_PC2_filtered

oral_pca_biplots_combined_no_rl_filtered <- ggarrange(
  pca_kraken2_biplot_PC1_filtered, pca_kraken2_biplot_PC2_filtered, 
  labels = c("A", "B"), 
  ncol = 1, nrow = 2, 
  common.legend = TRUE, 
  legend = "bottom")

oral_pca_biplots_combined_no_rl_filtered

#ggsave("./06-publication/main_figures/Figure5AB_npost_oral_pca_biplot_well_preserved", plot = oral_pca_biplots_combined_no_rl_filtered, device = "pdf", scale = 1, width = 8, height = 12, units = c("in"), dpi = 300)
```


Species of interest
These are the top ten species driving each PC
```{r}
pos_PC1 <- c("Mesocricetibacter sp905373425", "Aggregatibacter kilianii", "Aggregatibacter aphrophilus", "Neisseria elongata", "Lautropia mirabilis", "Neisseria sicca", "Fusobacterium hwasookii", "JABCPE02 sp013333875", "JABCPE02 sp013333255", "Aggregatibacter segnis")

pos_PC2 <- c("Actinomyces sp900323545", "Corynebacterium durum", "Methanobrevibacter A oralis", "Actinomyces howellii", "Actinomyces sp000195595", "Actinomyces oricola", "Actinomyces viscosus", "Actinomyces slackii", "Mogibacterium diversum", "Actinomyces oris")

neg_PC1 <- c("Olsenella uli", "Slackia exigua", "Olsenella F profusa", "Flexilinea sp001717545", "Olsenella F sp001189515", "Actinomyces dentalis", "Actinomyces sp001278845", "Nanosynbacter sp905373795", "Saccharimonas sp018127705", "Nanosynbacter sp905372305")

neg_PC2 <- c("Phocaeicola abscessus", "Bacteroides heparinolyticus", "Desulfovibrio sp003860215", "Centipeda periodontii", "Ottowia sp001262075", "Centipeda sp001717585", "Centipeda sp905372865", "Pyramidobacter piscolens", "Prevotella maculosa", "Leptotrichia buccalis")
```


Isolate the taxIDs of the species of interest so you can pull just those reads along with each read's length (in separate bash script)
```{r}
species_of_interest <- c(pos_PC1, pos_PC2, neg_PC1, neg_PC2)

species_taxids <- fread("./05-results/normalize_kraken/kraken2_names.dmp",
               sep = "\t"
                 ) %>%
  dplyr::select(V1, V3) %>%
  rename(parentID = V1,
         taxonomy = V3) %>%
  mutate(taxonomy = str_replace_all(taxonomy, "_", " ")) 


strain_taxids <- fread("./05-results/normalize_kraken/kraken2_nodes.dmp",
               sep = "\t"
                 ) %>%
  dplyr::select(V1, V3) %>%
  rename(taxid = V1,
         parentID = V3) %>%
  inner_join(., species_taxids, by = "parentID") %>%
  mutate(taxonomy = str_replace(taxonomy, " (?:[A-Z]+)$", "")) %>%
  filter(taxonomy %in% species_of_interest)

#fwrite(strain_taxids, "./05-results/normalize_kraken/dsvss_npost_oral_PCs_taxIDs.tsv", sep = "\t")

```


Find the average read length for the species dominating the PCs in both dsLibs and ssLibs
```{r}
taxIDs <- fread("./05-results/normalize_kraken/dsvss_npost_oral_PCs_taxIDs.tsv")

```


```{r}
files <- list.files(path = "./05-results/normalize_kraken/oral_well_preserved_PCs", pattern = ".oral_well_preserved_PCs.tsv", full.names = T)

gc_rl_oral_PCs_table <- function(files) {
  oral_gc_rl <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(read_header = 1,
          taxid = 2,
          read_length = 3,
          GC = 4) %>%
    left_join(., taxIDs %>%
              dplyr::select(taxid, taxonomy)) %>%
    mutate(Sequencing_ID = basename(fn)) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".oral_well_preserved_PCs.tsv", "")) %>%
    inner_join(., metadata_no_rl %>%
              dplyr::select(Sequencing_ID, Strandedness)) 
  })
  
  oral_gc_rl_df <- do.call(rbind, oral_gc_rl) 
  oral_gc_rl_df <- oral_gc_rl_df %>%
    filter(!str_detect(Sequencing_ID, "BAN001")) %>%
    filter(!str_detect(Sequencing_ID, "FUM003")) 
  
  return(oral_gc_rl_df)
}


oral_pcs_gc_rl <- gc_rl_oral_PCs_table(files)
```


```{r}
#group by taxonomy and strandedness to calculate average read length and %GC for each library type
##Read length
read_length_avs <- oral_pcs_gc_rl %>%
  group_by(taxonomy, Sequencing_ID) %>%
  mutate(av_rl_sample = mean(read_length)) %>%
  ungroup() %>%
  dplyr::select(taxonomy, Sequencing_ID, Strandedness, av_rl_sample) %>%
  unique() %>%
  group_by(taxonomy, Strandedness) %>%
  mutate(av_rl = mean(av_rl_sample)) %>% #average just based on those who have the species present 
  dplyr::select(taxonomy, Strandedness, av_rl) %>%
  unique() %>%
  filter(taxonomy %in% pos_PC1) %>% #fill in PC of interest
  mutate(taxonomy = factor(taxonomy, levels = pos_PC1)) %>% #fill in PC of interest
  arrange(taxonomy) %>%
  mutate(av_rl = round(av_rl, 1))

##GC
GC_avs <- oral_pcs_gc_rl %>%
  group_by(taxonomy, Sequencing_ID) %>%
  mutate(av_gc_sample = mean(GC)) %>% 
  ungroup() %>%
  dplyr::select(taxonomy, Sequencing_ID, Strandedness, av_gc_sample) %>%
  unique() %>%
  group_by(taxonomy, Strandedness) %>%
  mutate(av_gc = mean(av_gc_sample)) %>%
  dplyr::select(taxonomy, Strandedness, av_gc) %>%
  unique() %>%
  filter(taxonomy %in% pos_PC1) %>% #fill in PC of interest
  mutate(taxonomy = factor(taxonomy, levels = pos_PC1)) %>% #fill in PC of interest
  arrange(taxonomy) %>%
  mutate(av_gc = round(av_gc, 1))

#Average number of reads
av_read_count <- oral_pcs_gc_rl %>%
  group_by(taxonomy, Sequencing_ID) %>%
  mutate(read_count = n()) %>%
  ungroup() %>%
  dplyr::select(taxonomy, Sequencing_ID, Strandedness, read_count) %>%
  unique() %>%
  group_by(taxonomy, Strandedness) %>%
  mutate(av_read_count = mean(read_count)) %>%
  dplyr::select(taxonomy, Strandedness, av_read_count) %>%
  unique() %>%
  filter(taxonomy %in% pos_PC1) %>% #fill in PC of interest
  mutate(taxonomy = factor(taxonomy, levels = pos_PC1)) %>% #fill in PC of interest
  arrange(taxonomy) %>%
  mutate(av_read_count = round(av_read_count, 0))
  
```


#Focus on read length distributions of specific species
##Prepare the read length tables
```{r}
files <- list.files(path = "./05-results/normalize_kraken/oral_well_preserved_PCs", pattern = ".oral_well_preserved_PCs.tsv", full.names = T)

taxIDs <- fread("./05-results/normalize_kraken/dsvss_npost_oral_PCs_taxIDs.tsv")

rl_oral_PCs_table <- function(files) {
  oral_gc_rl <- lapply(files, function(fn) {
    fread(fn) %>%
    rename(read_header = 1,
          taxid = 2,
          read_length = 3,
          GC = 4) %>%
    left_join(., taxIDs %>%
              dplyr::select(taxid, taxonomy)) %>%
    select(-read_header, -taxid, -GC) %>%
    group_by(taxonomy, read_length) %>%
    summarise(counts = n()) %>%
    mutate(Sequencing_ID = basename(fn)) %>%
    mutate(Sequencing_ID = str_replace(Sequencing_ID, ".oral_well_preserved_PCs.tsv", "")) %>%
    inner_join(., metadata_no_rl %>%
              dplyr::select(Sequencing_ID, Strandedness)) 
  })
  
  oral_gc_rl_df <- do.call(rbind, oral_gc_rl) 
  oral_gc_rl_df <- oral_gc_rl_df %>%
    filter(!str_detect(Sequencing_ID, "BAN001")) %>%
    filter(!str_detect(Sequencing_ID, "FUM003")) 
  
  return(oral_gc_rl_df)
}


oral_pcs_rl <- rl_oral_PCs_table(files)

gc_categories <- fread("./05-results/normalize_kraken/dsvss_npost_oral_pc_species_gc_categories.csv")

oral_pcs_rl <- oral_pcs_rl %>%
  full_join(., gc_categories, by = "taxonomy")
```

## Calculate the median read length for each library
```{r}
#function to calculate the median read length of oral reads per libraryID
median_rl_summary_table <- function(sample_list, oral_reads_length_table) {
  
  #Create a blank dataframe
  read_length_summary_table <- data.frame(Individual_ID = sample_list,
                                          Double = numeric(length(sample_list)),
                                          Single = numeric(length(sample_list)),
                                          stringsAsFactors = FALSE)
  
  #Function to calculate median read length based on dataset, sample, and strandedness
  calculate_median_rl <- function(data, sample, strandedness) {
    rl_summary <- data %>%
      dplyr::select(Library_ID, Individual_ID, Strandedness, read_length, counts) %>%
      mutate(read_length = as.integer(as.character(read_length))) %>%
      filter(Individual_ID == sample, Strandedness == strandedness) %>%
      uncount(counts) %>%
      mutate(read_length = as.integer(read_length)) %>%
      summarize(median_rl = median(read_length))
    
    return(rl_summary$median_rl)
  }
  
  
  for (sample in sample_list) {
    rl_value_of_median_ds <- calculate_median_rl(oral_reads_length_table, sample, "Double")
    rl_value_of_median_ss <- calculate_median_rl(oral_reads_length_table, sample, "Single")
    read_length_summary_table[read_length_summary_table$Individual_ID == sample, "Double"] <- rl_value_of_median_ds
    read_length_summary_table[read_length_summary_table$Individual_ID == sample, "Single"] <- rl_value_of_median_ss
    
   }

  return(read_length_summary_table)
  
}

samples <- metadata_no_rl %>%
  dplyr::select(Individual_ID) %>%
  unique() %>%
  pull()

```

##Low GC content - Aggregatibacter kilianii
```{r}
a_kilianii_table <- oral_pcs_rl %>%
  ungroup() %>%
  filter(taxonomy == "Aggregatibacter kilianii") %>%
  full_join(., metadata_no_rl %>%
              select(Sequencing_ID, Library_ID, Individual_ID)) %>%
  select(-taxonomy)

median_rl_a_kilianii <- median_rl_summary_table(samples, a_kilianii_table) %>%
  pivot_longer(-Individual_ID, values_to = "median_rl", names_to = "Strandedness")

```

```{r}
read_length_table <- a_kilianii_table %>%
  filter(Individual_ID == "GOY005") 

median_rl_info <- median_rl_a_kilianii %>%
  filter(Individual_ID == "GOY005") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 800)) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/main_figures/Figure5C_npost_a_kilianii_GOY005_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```


##Intermediate GC content -  Neisseria elongata
```{r}
n_elongata_table <- oral_pcs_rl %>%
  ungroup() %>%
  filter(taxonomy == "Neisseria elongata") %>%
  full_join(., metadata_no_rl %>%
              select(Sequencing_ID, Library_ID, Individual_ID)) %>%
  select(-taxonomy)

median_rl_n_elongata <- median_rl_summary_table(samples, n_elongata_table) %>%
  pivot_longer(-Individual_ID, values_to = "median_rl", names_to = "Strandedness")
```


```{r}
read_length_table <- n_elongata_table %>%
  filter(Individual_ID == "GOY005") 

median_rl_info <- median_rl_n_elongata %>%
  filter(Individual_ID == "GOY005") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 200)) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/main_figures/Figure5D_npost_n_elongata_GOY005_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```

##High GC content - Lautropia mirabilis
```{r}
l_mirabilis_table <- oral_pcs_rl %>%
  ungroup() %>%
  filter(taxonomy == "Lautropia mirabilis") %>%
  full_join(., metadata_no_rl %>%
              select(Sequencing_ID, Library_ID, Individual_ID)) %>%
  select(-taxonomy)

median_rl_l_mirabilis <- median_rl_summary_table(samples, l_mirabilis_table) %>%
  pivot_longer(-Individual_ID, values_to = "median_rl", names_to = "Strandedness")
```


```{r}
read_length_table <- l_mirabilis_table %>%
  filter(Individual_ID == "GOY005") 

median_rl_info <- median_rl_l_mirabilis %>%
  filter(Individual_ID == "GOY005") %>%
  dplyr::select(Strandedness, median_rl) 

rl_plot <- read_length_table %>%
  ggplot(aes(x = read_length, y = counts, fill = Strandedness, linetype = Strandedness)) +
    geom_bar(data = . %>% filter(Strandedness == "Double"),
              aes(x = read_length, y = counts),
              fill = "#04387a", alpha = 1, stat = "identity") +
    geom_bar(data = . %>% filter(Strandedness == "Single"),
              aes(x = read_length, y = counts),
              fill = "#6fbfdc", alpha = 0.6, stat = "identity") +
    theme_minimal() +
    labs(x = "Read Length (bp)", y = "Read Count") +
    guides(fill = guide_legend(title = "Strandedness"),
           linetype = guide_legend(title = "Strandedness")) +
    scale_x_continuous(
        limits = c(30, 150),
        breaks = seq(30, 140, by = 10),
        expand = c(0, 0)) +    
     theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20), 
          axis.text.y = element_text(size = 22), 
          axis.title = element_text(size = 22),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")) + 
    expand_limits(y = c(0, 1500)) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red") +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) 
  
rl_plot

#ggsave("./06_publication/main_figures/Figure5E_npost_l_mirabilis_GOY005_rl.pdf", plot = rl_plot, device = "pdf", scale = 1, width = 5, height = 4, units = "in", dpi = 300)
```

