---
title: "Preservation, alpha diversity, and beta diversity analyses for the NPre dataset"
author: "Keri Burge"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output:
    github_document:
        fig_width: 8
        fig_height: 6
        toc: true
---


```{r, eval=FALSE, echo=FALSE, message=FALSE}
install.packages("cuperdec")
```


```{r, eval=TRUE, echo=FALSE, message=FALSE}
#Load in all necessary packages
library(knitr)
library(colorspace)
library(data.table)
library(cuperdec)
library(ggplot2)
library(viridis)
library(ggpubr)
library(rstatix)
library(purrr)
library(patchwork)
library(ggdist)
library(gghalves)
library(ggtext)
library(gplots)
library(mixOmics)
library(ggrepel)
library(vegan)
library(janitor)
library(dplyr)
library(tidyverse)
```


```{r setup}
knitr::opts_knit$set(root.dir = normalizePath(".."))
```



#Information for saving
```{r}
folder <- "supplemental_figures"
file_modifier <- "NPre"

#Ensure the supplemental figures folder exists
#dir.create(file.path("./06-publication", folder), recursive = TRUE, showWarnings = FALSE)
```


#Colors
```{r}
double_color <- "#04387a"
single_color <- "#6fbfdc"
  
```

#Metadata table 
Load in metadata 
```{r}
metadata <- fread("./05-results/normalize_eager/dsvss_npre_w_rl_subsampling_metadata.tsv") 

metadata_no_rl <-metadata %>%
  filter(Read_length_subsampling == "none")

metadata_rl <- metadata %>%
  filter(str_detect(Read_length_subsampling, "bp")) %>%
  mutate(Strandedness2 = Strandedness) %>%
  relocate(Strandedness2, .before = Read_length_subsampling) %>%
  unite(Strandedness_rl, Strandedness2:Read_length_subsampling, sep = "_")

metadata_rl_over50bp <- metadata %>%
  filter(str_detect(Read_length_subsampling, "over50bp"))

metadata_rl_under50bp <- metadata %>%
  filter(str_detect(Read_length_subsampling, "under50bp"))
```


# Oral Species List 
List of oral species in the format of the GTDB r207 database
```{r}
#GTDB Database
gtdb_sources <- fread("./00-documentation/dsvss_gtdb_r207_species_sources.tsv", header = TRUE) 

oral_gtdb <- gtdb_sources %>%
  mutate(Isolation_Source = str_replace_all(Isolation_Source, "oral", "TRUE"),
         Isolation_Source = str_replace_all(Isolation_Source, "other", "FALSE")) %>%
  rename(Taxon = 1) %>%
  mutate(Isolation_Source = as.logical(Isolation_Source)) %>%
  mutate(Taxon = str_replace(Taxon, " (?:[A-Z]+)$", "")) %>%
  unique() 

```


# Read in the kraken2 species table (this table was prepared in the tax_sourcetracker_prep file; decontam was run)

```{r}
kraken2_total_species <- fread("./05-results/normalize_eager/dsvss_npre_w_read_length_kraken2_gtdb207_total_species_decontam.tsv") %>%
  mutate(Species = str_replace(Species, " (?:[A-Z]+)$", "")) %>%
  group_by(Species) %>%
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) %>%
  mutate(across(-Species, ~ if_else(.x < 10, 0, .x))) #only count a species as present if it has at least 10 reads

#fwrite(kraken2_total_species, "./05-results/normalize_eager/dsvss_npre_w_read_length_kraken2_gtdb207_total_species_decontam_filtered.tsv", sep = "\t")

#filter for no read length subsampling
no_rl_samples <- metadata_no_rl %>%
  dplyr::select(Sequencing_ID) %>%
  pull()

kraken2_total_species_no_rl <- kraken2_total_species %>%
  column_to_rownames("Species") %>%
  dplyr::select(all_of(no_rl_samples)) %>%
  rownames_to_column("Species")

#filter for only read length subsampling
## all read length subsampling
rl_samples <- metadata_rl %>%
  dplyr::select(Sequencing_ID) %>%
  pull()

kraken2_total_species_rl <- kraken2_total_species %>%
  column_to_rownames("Species") %>%
  dplyr::select(all_of(rl_samples)) %>%
  rownames_to_column("Species")

## over 50 bp
over50bp_samples <- metadata_rl %>%
  dplyr::select(Sequencing_ID) %>%
  filter(str_detect(Sequencing_ID, "over")) %>%
  pull()

kraken2_total_species_rl_over50bp <- kraken2_total_species %>%
  column_to_rownames("Species") %>%
  dplyr::select(all_of(over50bp_samples)) %>%
  rownames_to_column("Species")

# under 50 bp
under50bp_samples <- metadata_rl %>%
  dplyr::select(Sequencing_ID) %>%
  filter(str_detect(Sequencing_ID, "under")) %>%
  pull()

kraken2_total_species_rl_under50bp <- kraken2_total_species %>%
  column_to_rownames("Species") %>%
  dplyr::select(all_of(under50bp_samples)) %>%
  rownames_to_column("Species")

oral_db <- oral_gtdb
```


# Cuperdec Curves
Cumulative percent decay (cuperdec) curves generated. Adaptive burn-in filter with a threshold of 50%. ‘Good’ (blue lines) indicates the sample passes the threshold and is “well preserved” with a sufficiently high proportion of oral taxa. ‘Bad’ (red lines) indicates that the sample does not pass the threshold and is “poorly preserved”, with an insufficient proportion of oral taxa.

Indicate the oral database 
```{r}
#Determine what oral database will be used
oral_db <- oral_gtdb

```

## No read length subsampling
```{r}
#Format metadata using load_map()  
sample_metadata_no_rl <- load_map(metadata_no_rl,
                     sample_col = "Sequencing_ID",
                     source_col = "Subsampled_Individual_ID")

#Format Kraken2 Species table using load_taxa_table() - converts to long form with proper column headings
kraken2_species_cuperdec_no_rl <- load_taxa_table(kraken2_total_species_no_rl)


#Calculate the curves
curves_no_rl <- calculate_curve(kraken2_species_cuperdec_no_rl, database = oral_db)

#Apply an adaptive burnin (only starting to considering whether a curve exceeds minimum percentage target source threshold (50%), from the point at which the difference of the percentage target source of one rank to the next, does not exceed the standard-deviation of all differences)
burnin_result_no_rl <- adaptive_burnin_filter(curves_no_rl, percent_threshold = 50) 

#Add additional metadata to the burnin result for a pass/fail heatmap
burnin_result_no_rl <- burnin_result_no_rl %>%
  rename(Sequencing_ID = 1) %>%
  full_join(., metadata_no_rl)

#Heat map showing whether or not samples passed the 50% threshold with an adaptive burn-in 
#Reverse factor levels so samples are alphabetical
burnin_result_no_rl$Subsampled_Individual_ID <- factor(burnin_result_no_rl$Subsampled_Individual_ID, levels = unique(burnin_result_no_rl$Subsampled_Individual_ID))
reversed_levels_no_rl <- rev(levels(burnin_result_no_rl$Subsampled_Individual_ID))
burnin_result_no_rl$Subsampled_Individual_ID <- factor(burnin_result_no_rl$Subsampled_Individual_ID, levels = reversed_levels_no_rl)

#Plot the heat map 
cuperdec_heatmap_no_rl <- burnin_result_no_rl %>%
  rename(Preservation = Passed) %>%
  mutate(Preservation = str_replace(Preservation, "TRUE", "Good"),
         Preservation = str_replace(Preservation, "FALSE", "Bad"),
         Preservation = factor(Preservation, levels = c("Good", "Bad"))) %>%
  ggplot(., aes(x = Strandedness, y = Subsampled_Individual_ID, fill = Preservation)) +
  geom_tile() +
  theme_minimal() +
  geom_tile(color = "white",
            lwd = 1.5,
            linetype = 1) +
  labs(title = "", x = "Strandedness", y = "Individual ID") +
  scale_fill_discrete(name = "Preservation", type = c("#666666", "#CCCCCC")) +
  theme(legend.position="bottom", plot.title = element_text(hjust = 0.5)) 
  
cuperdec_heatmap_no_rl

```


```{r}
#Add example cuperdec data so a plaque and environmental curve panel can be added to the panel figure of each library type
#Cuperdec examples
data(cuperdec_taxatable_ex)
data(cuperdec_database_ex)
data(cuperdec_metadata_ex)
taxatable <- load_taxa_table(cuperdec_taxatable_ex)
curves_ex <- calculate_curve(taxatable, database = oral_db)

#Select 5 subplaque samples and 5 supplaque samples
supPlaque_samples <- cuperdec_metadata_ex %>%
  filter(str_detect(Env, "supPlaque")) %>%
  slice(1:5) %>%
  dplyr::select('#SampleID') %>%
  pull()
  
subPlaque_samples <- cuperdec_metadata_ex %>%
  filter(str_detect(Env, "subPlaque")) %>%
  slice(1:5) %>%
  dplyr::select('#SampleID') %>%
  pull()
  
plaque_samples <- unlist(c(supPlaque_samples,subPlaque_samples ))

#Select 10 environmental samples
environmental_control_samples <- cuperdec_metadata_ex %>%
  filter(str_detect(Env, "EnvironmentalControl")) %>%
  slice(1:10) %>%
  dplyr::select('#SampleID') %>%
  pull()

#Combine all samples
samples <- unlist(c(plaque_samples, environmental_control_samples))

#Calculate curves
curves_ex_plaque <- curves_ex %>%
  filter(Sample %in% samples)

#Join with the ancient samples
curves_full_no_rl <- curves_no_rl %>%
  full_join(., curves_ex_plaque)

#Create metadata table rows 
samples_df <- as.data.frame(samples) %>%
  rename(Sample = 1) %>%
  mutate(Sample_Source = ifelse(str_detect(Sample, "SRR"), "Plaque", "Environment"))

#Add the plaque and environmental data
sample_metadata_w_sources_no_rl <- bind_rows(sample_metadata_no_rl, samples_df)

#Arrange in alphabetical order of library types followed by plaque and environment
group_order <- c("BAN001_10M", "CDM019_10M", "CDM020_10M", "FUM001_10M", "FUM002_10M", "FUM003_10M", "GDN001_10M", "GOY005_10M", "GOY006_10M", "GUE001_10M", "PES001_10M", "SPM002_10M", "VLD001_10M", "ZAF001_10M", "Plaque", "Environment")

#Calculate curved with adaptive burnin (50%)
burnin_result_no_rl <- adaptive_burnin_filter(curves_full_no_rl, percent_threshold = 50)

#Set order of panes
sample_metadata_w_sources_no_rl$Sample_Source <- factor(sample_metadata_w_sources_no_rl$Sample_Source, levels = group_order)

```


```{r}
#Lets adjust the plot function to change axes labels and make the transparancey of lines based on strandedness
plot_grouped_burnin <- function(curves, metadata_cuperdec, burnin_result, metadata_full, restrict_x, facet_cols = NULL) {
  table_meta <- dplyr::left_join(curves,
    metadata_cuperdec,
    by = c("Sample")
  ) %>%
    dplyr::left_join(burnin_result, by = c("Sample"))
  
  metadata_full <- metadata_full %>%
    rename(Sample = sample_column) %>%
    dplyr::select(Sample, Strandedness)
  
  alpha_strandedness <- c("Double" = 1, "Single" = 0.5)
  colors_preservation <- c("Good" = "#03bfc4", "Bad" = "#f8766d")
  
  table_meta <- table_meta %>%
    full_join(., metadata_full) %>%
    mutate(Strandedness = ifelse(is.na(Strandedness), "Double", Strandedness)) %>%
    rename(Preservation = Passed) %>%
    mutate(Preservation = str_replace(Preservation, "TRUE", "Good"),
           Preservation = str_replace(Preservation, "FALSE", "Bad"),
           Preservation = factor(Preservation, levels = c("Good", "Bad")))
  
  ggplot(table_meta, aes(x = Rank, y = Fraction_Target, group = Sample, color = Preservation, alpha = Strandedness)) +
    geom_line() +
    ylim(0, 100) +
    xlab("Species Abundance Rank") +
    ylab("Percentage Oral") +
    facet_wrap(~Sample_Source, ncol = facet_cols) +
    theme_minimal() +
    scale_color_manual(values = colors_preservation, name = "Preservation") +
    xlim(0, restrict_x) +
    scale_alpha_manual(values = alpha_strandedness, name = "Strandedness")
    
}

#Identify the column in the metadata_full table that contains the sample names
sample_column <- "Sequencing_ID"

#Plot 
cuperdec_w_sources_no_rl <- plot_grouped_burnin(curves_full_no_rl, sample_metadata_w_sources_no_rl, burnin_result_no_rl, metadata_no_rl, restrict_x = 250)

cuperdec_w_sources_no_rl

```


```{r}
#Combine plots
cuperdec_combined_no_rl <- ggarrange(
  cuperdec_w_sources_no_rl, cuperdec_heatmap_no_rl, 
  labels = c("A", "B"), 
  ncol = 2, nrow = 1, 
  legend = "bottom"
)

#ggsave(filename = file.path("./06-publication", paste0(folder), paste0("SI_Fig4A_cuperdec_combined_plots_", file_modifier, "_no_rl.pdf")), plot = cuperdec_combined_no_rl, device = "pdf", scale = 1, width = 16, height = 6, units = c("in"), dpi = 300)
```


## Over 50 bp
```{r}
#Format metadata using load_map()  
sample_metadata_rl_over50bp <- load_map(metadata_rl_over50bp,
                     sample_col = "Sequencing_ID",
                     source_col = "Subsampled_Individual_ID")

#Format Kraken2 Species table using load_taxa_table() - converts to long form with proper column headings
kraken2_species_cuperdec_rl_over50bp <- load_taxa_table(kraken2_total_species_rl_over50bp)


#Calculate the curves
curves_rl_over50bp <- calculate_curve(kraken2_species_cuperdec_rl_over50bp, database = oral_db)

#Apply an adaptive burnin (only starting to considering whether a curve exceeds minimum percentage target source threshold (50%), from the point at which the difference of the percentage target source of one rank to the next, does not exceed the standard-deviation of all differences)
burnin_result_rl_over50bp <- adaptive_burnin_filter(curves_rl_over50bp, percent_threshold = 50) 

#Add additional metadata to the burnin result for a pass/fail heatmap
burnin_result_rl_over50bp <- burnin_result_rl_over50bp %>%
  rename(Sequencing_ID = 1) %>%
  full_join(., metadata_rl_over50bp)

#Heat map showing whether or not samples passed the 50% threshold with an adaptive burn-in 
#Reverse factor levels so samples are alphabetical
burnin_result_rl_over50bp$Subsampled_Individual_ID <- factor(burnin_result_rl_over50bp$Subsampled_Individual_ID, levels = unique(burnin_result_rl_over50bp$Subsampled_Individual_ID))

reversed_levels_rl_over50bp <- rev(levels(burnin_result_rl_over50bp$Subsampled_Individual_ID))

burnin_result_rl_over50bp$Subsampled_Individual_ID <- factor(burnin_result_rl_over50bp$Subsampled_Individual_ID, levels = reversed_levels_rl_over50bp)

#Plot the heat map 
cuperdec_heatmap_rl_over50bp <- burnin_result_rl_over50bp %>%
  rename(Preservation = Passed) %>%
  mutate(Preservation = str_replace(Preservation, "TRUE", "Good"),
         Preservation = str_replace(Preservation, "FALSE", "Bad"),
         Preservation = factor(Preservation, levels = c("Good", "Bad"))) %>%
  ggplot(., aes(x = Strandedness, y = Subsampled_Individual_ID, fill = Preservation)) +
  geom_tile() +
  theme_minimal() +
  geom_tile(color = "white",
            lwd = 1.5,
            linetype = 1) +
  labs(title = "", x = "Strandedness", y = "Individual ID") +
  scale_fill_discrete(name = "Preservation", type = c("#666666", "#CCCCCC")) +
  theme(legend.position="bottom", plot.title = element_text(hjust = 0.5)) 
  
cuperdec_heatmap_rl_over50bp

```


```{r}
#Join environmental and plaque samples with the ancient samples
curves_full_rl_over50bp <- curves_rl_over50bp %>%
  full_join(., curves_ex_plaque)

#Add the plaque and environmental data
sample_metadata_w_sources_rl_over50bp <- bind_rows(sample_metadata_rl_over50bp, samples_df)

#Arrange in alphabetical order of library types followed by plaque and environment
group_order <- c("BAN001_10M_over50bp", "CDM019_10M_over50bp", "CDM020_10M_over50bp", "FUM001_10M_over50bp", "FUM002_10M_over50bp", "FUM003_10M_over50bp", "GDN001_10M_over50bp", "GOY005_10M_over50bp", "GOY006_10M_over50bp", "GUE001_10M_over50bp", "PES001_10M_over50bp", "SPM001_10M_over50bp", "VLD001_10M_over50bp", "ZAF001_10M_over50bp", "Plaque", "Environment")

#Calculate curved with adaptive burnin (50%)
burnin_result_rl_over50bp <- adaptive_burnin_filter(curves_full_rl_over50bp, percent_threshold = 50)

#Set order of panes
sample_metadata_w_sources_rl_over50bp$Sample_Source <- factor(sample_metadata_w_sources_rl_over50bp$Sample_Source, levels = group_order)

#Identify the column in the metadata_full table that contains the sample names
sample_column <- "Sequencing_ID"

#Plot 
cuperdec_w_sources_rl_over50bp <- plot_grouped_burnin(curves_full_rl_over50bp, sample_metadata_w_sources_rl_over50bp, burnin_result_rl_over50bp, metadata_rl_over50bp, restrict_x = 250)

cuperdec_w_sources_rl_over50bp

```


## Under 50 bp
```{r}
#Format metadata using load_map()  
sample_metadata_rl_under50bp <- load_map(metadata_rl_under50bp,
                     sample_col = "Sequencing_ID",
                     source_col = "Subsampled_Individual_ID")

#Format Kraken2 Species table using load_taxa_table() - converts to long form with proper column headings
kraken2_species_cuperdec_rl_under50bp <- load_taxa_table(kraken2_total_species_rl_under50bp)


#Calculate the curves
curves_rl_under50bp <- calculate_curve(kraken2_species_cuperdec_rl_under50bp, database = oral_db)

#Apply an adaptive burnin (only starting to considering whether a curve exceeds minimum percentage target source threshold (50%), from the point at which the difference of the percentage target source of one rank to the next, does not exceed the standard-deviation of all differences)
burnin_result_rl_under50bp <- adaptive_burnin_filter(curves_rl_under50bp, percent_threshold = 50) 

#Add additional metadata to the burnin result for a pass/fail heatmap
burnin_result_rl_under50bp <- burnin_result_rl_under50bp %>%
  rename(Sequencing_ID = 1) %>%
  full_join(., metadata_rl_under50bp)

#Heat map showing whether or not samples passed the 50% threshold with an adaptive burn-in 
#Reverse factor levels so samples are alphabetical
burnin_result_rl_under50bp$Subsampled_Individual_ID <- factor(burnin_result_rl_under50bp$Subsampled_Individual_ID, levels = unique(burnin_result_rl_under50bp$Subsampled_Individual_ID))

reversed_levels_rl_under50bp <- rev(levels(burnin_result_rl_under50bp$Subsampled_Individual_ID))

burnin_result_rl_under50bp$Subsampled_Individual_ID <- factor(burnin_result_rl_under50bp$Subsampled_Individual_ID, levels = reversed_levels_rl_under50bp)

#Plot the heat map 
cuperdec_heatmap_rl_under50bp <- burnin_result_rl_under50bp %>%
  rename(Preservation = Passed) %>%
  mutate(Preservation = str_replace(Preservation, "TRUE", "Good"),
         Preservation = str_replace(Preservation, "FALSE", "Bad"),
         Preservation = factor(Preservation, levels = c("Good", "Bad"))) %>%
  ggplot(., aes(x = Strandedness, y = Subsampled_Individual_ID, fill = Preservation)) +
  geom_tile() +
  theme_minimal() +
  geom_tile(color = "white",
            lwd = 1.5,
            linetype = 1) +
  labs(title = "", x = "Strandedness", y = "Individual ID") +
  scale_fill_discrete(name = "Preservation", type = c("#666666", "#CCCCCC")) +
  theme(legend.position="bottom", plot.title = element_text(hjust = 0.5)) 
  
cuperdec_heatmap_rl_under50bp

```


```{r}
#Join environmental and plaque samples with the ancient samples
curves_full_rl_under50bp <- curves_rl_under50bp %>%
  full_join(., curves_ex_plaque)

#Add the plaque and environmental data
sample_metadata_w_sources_rl_under50bp <- bind_rows(sample_metadata_rl_under50bp, samples_df)

#Arrange in alphabetical order of library types followed by plaque and environment
group_order <- c("BAN001_10M_under50bp", "CDM019_10M_under50bp", "CDM020_10M_under50bp", "FUM001_10M_under50bp", "FUM002_10M_under50bp", "FUM003_10M_under50bp", "GDN001_10M_under50bp", "GOY005_10M_under50bp", "GOY006_10M_under50bp", "GUE001_10M_under50bp", "PES001_10M_under50bp", "SPM001_10M_under50bp", "VLD001_10M_under50bp", "ZAF001_10M_under50bp", "Plaque", "Environment")

#Calculate curved with adaptive burnin (50%)
burnin_result_rl_under50bp <- adaptive_burnin_filter(curves_full_rl_under50bp, percent_threshold = 50)

#Set order of panes
sample_metadata_w_sources_rl_under50bp$Sample_Source <- factor(sample_metadata_w_sources_rl_under50bp$Sample_Source, levels = group_order)

#Identify the column in the metadata_full table that contains the sample names
sample_column <- "Sequencing_ID"

#Plot 
cuperdec_w_sources_rl_under50bp <- plot_grouped_burnin(curves_full_rl_under50bp, sample_metadata_w_sources_rl_under50bp, burnin_result_rl_under50bp, metadata_rl_under50bp, restrict_x = 250)

cuperdec_w_sources_rl_under50bp

```

## Combine read length subsampling
```{r}
#Combine plots
cuperdec_combined_rl <- ggarrange(
  cuperdec_w_sources_no_rl, cuperdec_heatmap_no_rl, 
  cuperdec_w_sources_rl_over50bp, cuperdec_heatmap_rl_over50bp,
  cuperdec_w_sources_rl_under50bp, cuperdec_heatmap_rl_under50bp,
  labels = c("A", "B", "C", "D", "E", "F"), 
  ncol = 2, nrow = 3, 
  legend = "bottom"
)

cuperdec_combined_rl

#ggsave(filename = file.path("./06-publication", paste0(folder), paste0("SI_Fig4_cuperdec_combined_plots_", file_modifier, "_read_length_subsampling.pdf")), plot = cuperdec_combined_rl, device = "pdf", scale = 1, width = 16, height = 18, units = c("in"), dpi = 300)
```


#Alpha Diversity - No read length subsampling
## Total Species
```{r}
#Kraken2 Total Species Table

#Difference in total number of species
kraken2_species_cuperdec_no_rl <- load_taxa_table(kraken2_total_species_no_rl)

#Calculate number of species reported
kraken2_total_species_table_no_rl <- kraken2_species_cuperdec_no_rl %>%
  group_by(Sample) %>%
  summarize(count_distinct_kraken2 = n_distinct(Taxon)) %>% #count how many unique taxa are reported in each library
  rename(Sequencing_ID = 1) %>%
  full_join(., metadata_no_rl) %>%
  drop_na()

```


## Oral Species
```{r}
#Oral species table
oral_species <- oral_db %>%
  filter(str_detect(Isolation_Source, "TRUE")) %>%
  rename(Species = 1) %>%
  dplyr::select(-Isolation_Source)
  
oral_species_table_no_rl <- kraken2_total_species_no_rl %>%
  inner_join(., oral_species, by = "Species") 

#fwrite(oral_species_table_no_rl, "./05-results/normalize_eager/dsvss_npre_no_rl_kraken2_gtdb207_oral_species_decontam_filtered.tsv", sep = "\t", col.names = TRUE)
```


```{r}
#Calculate the number of classified reads to add to the metadata table

classified_reads_no_rl <- kraken2_total_species_no_rl %>%
  adorn_totals() %>%
  filter(Species == "Total") %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "classified_reads") %>%
  dplyr::select(-Species)

classified_reads_table_no_rl <- classified_reads_no_rl %>%
  full_join(metadata_no_rl %>%
              dplyr::select(Sequencing_ID, Strandedness, read_count, kraken2_preservation)) %>%
  rename(total_reads = read_count,
         preservation = kraken2_preservation)

oral_reads_no_rl <- oral_species_table_no_rl %>%
  adorn_totals() %>%
  filter(Species == "Total") %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "oral_reads") %>%
  dplyr::select(-Species)

classified_reads_table_no_rl <- classified_reads_table_no_rl %>%
  full_join(., oral_reads_no_rl) %>%
  dplyr::select(Sequencing_ID, Strandedness, preservation, total_reads, classified_reads, oral_reads)

```


```{r}
norm_eager_percentages <- classified_reads_table_no_rl %>%
  mutate(percent_oral_of_total = (oral_reads/total_reads)*100) %>%
  mutate(percent_oral_of_class = (oral_reads/classified_reads)*100)
```


Count number of oral species per sample
```{r}
kraken2_oral_species_table_no_rl <- load_taxa_table(oral_species_table_no_rl)

kraken2_species_summary_table_no_rl <- kraken2_oral_species_table_no_rl %>%
  group_by(Sample) %>%
  summarize(count_distinct_kraken2_filtered = n_distinct(Taxon)) %>%
  dplyr::rename(Sequencing_ID = 1) %>%
  full_join(., metadata_no_rl) %>%
  drop_na() %>%
  full_join(., kraken2_total_species_table_no_rl) %>%
  dplyr::rename(Total = count_distinct_kraken2,
         Oral = count_distinct_kraken2_filtered) %>%
  relocate(Total, .after = Oral) %>%
  pivot_longer(Oral:Total, names_to = "pre_v_post", values_to = "count") %>%
  mutate(count = replace_na(count, 0))

```


## Raincloud plot
Raincloud plot for total observed species
```{r}
kraken2_species_summary_table_no_rl <- kraken2_species_summary_table_no_rl %>%
  mutate(pre_v_post = str_replace(pre_v_post, "Total", "Total Species"),
         pre_v_post = str_replace(pre_v_post, "Oral", "Oral Species")) 

#Statistics - Total v. Oral (ds % ss)
stats_kraken_alpha_diversity_wilcox_strandedness_no_rl <- kraken2_species_summary_table_no_rl %>%
  group_by(Strandedness) %>%
  wilcox_test(count ~ (pre_v_post), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

stats_kraken_alpha_diversity_effectsize_strandedness_no_rl <- kraken2_species_summary_table_no_rl %>%
  group_by(Strandedness) %>%
  wilcox_effsize(formula = count ~ pre_v_post, paired = TRUE) 

#Stats for plot
wilcox_results <- kraken2_species_summary_table_no_rl %>%
  group_by(forcats::fct_rev(pre_v_post)) %>%
  wilcox_test(count ~ Strandedness, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- kraken2_species_summary_table_no_rl %>%
  group_by(forcats::fct_rev(pre_v_post)) %>%
  wilcox_effsize(formula = count ~ Strandedness, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat)) %>%
  dplyr::rename(pre_v_post = 1)

annotations <- wilcox_results %>%
  group_by(pre_v_post) %>%
  summarize(annotation = dplyr::first(annotation),.groups = 'drop') 



#PLOT
color <- c(double_color, single_color)

kraken2_raincloud_plot_no_rl <- kraken2_species_summary_table_no_rl %>%
  ggplot(., aes(x=Strandedness, y=count)) +
  ggdist::stat_halfeye(
     aes(color = Strandedness,
        fill = after_scale(lighten(color, .2))),
    adjust = 0.5, 
    width = 0.6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
    ) + 
  ggdist::stat_dots(
    side = "left", 
    dotsize = 1.5, 
    justification = 1.175, 
    binwidth =10,
    alpha = .3,
    range_scale = .4,
    aes(color = Strandedness, color = after_scale(darken(color, .1, space = "HLS")))
  ) +
  geom_boxplot(
    aes(color = Strandedness,
        color = after_scale(darken(color, .1, space = "HLS")),
        fill = after_scale(desaturate(lighten(color, .8), .4))),
    width = .15, 
    outlier.shape = NA
    ) +
  facet_grid(~forcats::fct_rev(pre_v_post)) +
  theme_minimal() +
  theme(
    legend.position = "none", 
    plot.title = element_text(hjust = 0.5)
    ) +
  labs(
    title = "", 
    y = "Observed Species", 
    x = ""
    ) +
  coord_cartesian(xlim = c(1.2, NA), clip = "off") +
  scale_color_manual(values = color, guide = "none") +
  scale_fill_manual(values = color, guide = "none") +
  theme(axis.text.x = element_text(size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
            strip.text = element_text(size = 14)) +
  expand_limits(y = c(0, 500)) +
  geom_text(data = annotations, aes(x = 1.5, y = 460, label = annotation), vjust = -0.5)

kraken2_raincloud_plot_no_rl

```


## Shannon Diversity
Shannon Diversity plot
```{r}
#Calculate the Shannon diversity index
#Total Species
kraken2_total_species_formatted_no_rl <- kraken2_total_species_no_rl %>%
  pivot_longer(2:ncol(.), names_to = "Sample", values_to = "Counts") %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  column_to_rownames("Sample")

shannon_diversity_total_no_rl <- diversity(kraken2_total_species_formatted_no_rl, index="shannon") %>%
  as.data.frame() %>%
  dplyr::rename(shannon_index_total = 1) %>%
  rownames_to_column("Sequencing_ID")

#Oral Species
#Need to reformat species table so species are the columns and samples are the rows
kraken2_oral_species_formatted_no_rl <- oral_species_table_no_rl %>%
  pivot_longer(2:ncol(.), names_to = "Sample", values_to = "Counts") %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  column_to_rownames("Sample")

shannon_diversity_oral_no_rl <- diversity(kraken2_oral_species_formatted_no_rl, index="shannon") %>%
  as.data.frame() %>%
  dplyr::rename(shannon_index_oral = 1) %>%
  rownames_to_column("Sequencing_ID")

#Combine tables
metadata_w_shannon_no_rl <- metadata_no_rl %>%
  full_join(., shannon_diversity_total_no_rl, by = "Sequencing_ID") %>%
  full_join(., shannon_diversity_oral_no_rl, by = "Sequencing_ID") %>%
  pivot_longer(shannon_index_total:shannon_index_oral, names_to = "total_v_oral", values_to = "shannon_index") 

#Statistics - Total v. Oral (ds + ss)
stats_kraken_alpha_diversity_wilcox_strandedness_no_rl <- metadata_w_shannon_no_rl %>%
  group_by(Strandedness) %>%
  wilcox_test(shannon_index ~ (total_v_oral), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

stats_kraken_alpha_diversity_effectsize_strandedness_no_rl <- metadata_w_shannon_no_rl %>%
  group_by(Strandedness) %>%
  wilcox_effsize(formula = shannon_index ~ total_v_oral, paired = TRUE) 

#Stats for plot
wilcox_results <- metadata_w_shannon_no_rl %>%
  group_by(total_v_oral) %>%
  wilcox_test(shannon_index ~ Strandedness, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- metadata_w_shannon_no_rl %>%
  group_by(total_v_oral) %>%
  wilcox_effsize(formula = shannon_index ~ Strandedness, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat))

annotations <- wilcox_results %>%
  group_by(total_v_oral) %>%
  summarize(annotation = dplyr::first(annotation),
            y_position = 5, .groups = 'drop') %>%
  mutate(total_v_oral = str_replace(total_v_oral, "shannon_index_oral", "Oral Species"),
         total_v_oral = str_replace(total_v_oral, "shannon_index_total", "Total Species"))

#Plot
kraken2_raincloud_plot_shannon_no_rl <- metadata_w_shannon_no_rl %>%
  mutate(total_v_oral = str_replace_all(total_v_oral, "shannon_index_oral", "Oral Species"),
         total_v_oral = str_replace_all(total_v_oral, "shannon_index_total", "Total Species")) %>%
  ggplot(., aes(x=Strandedness, y=shannon_index)) +
  ggdist::stat_halfeye(
     aes(color = Strandedness,
        fill = after_scale(lighten(color, .2))),
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
    ) + 
  ggdist::stat_dots(
    side = "left", 
    dotsize = 0.8, 
    justification = 1.175, 
    binwidth = 0.2,
    alpha = .3,
    range_scale = .4,
    aes(color = Strandedness, color = after_scale(darken(color, .1, space = "HLS")))
  ) +
  geom_boxplot(
    aes(color = Strandedness,
        color = after_scale(darken(color, .1, space = "HLS")),
        fill = after_scale(desaturate(lighten(color, .8), .4))),
    width = .15, 
    outlier.shape = NA
    ) +
  facet_grid(~forcats::fct_rev(total_v_oral)) +
  theme_minimal() +
  theme(
    legend.position = "none", 
    plot.title = element_text(hjust = 0.5)
    ) +
  labs(
    title = "", 
    y = "Shannon Index", 
    x = ""
    ) +
  coord_cartesian(xlim = c(1.2, NA), clip = "off") +
  scale_color_manual(values = color, guide = "none") +
  scale_fill_manual(values = color, guide = "none") +
  theme(axis.text.x = element_text(size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
            strip.text = element_text(size = 14)) +
  #stat_compare_means(method = "wilcox.test", paired = TRUE, label = "p.signif", label.x = 1.5, label.y = 4.6, size = 4, hide.ns = TRUE) +
  expand_limits(y = c(0, 5)) +
  geom_text(data = annotations, aes(x = 1.5, y = 4.7, label = annotation), vjust = -0.5)

kraken2_raincloud_plot_shannon_no_rl

```


Combine the Observed species and Shannon diversity plots
```{r}
alpha_diversity_combined_no_rl <- ggarrange(
  kraken2_raincloud_plot_no_rl, kraken2_raincloud_plot_shannon_no_rl, 
  labels = c("A", "B"), 
  ncol = 1, nrow = 2, 
  common.legend = TRUE, 
  legend = "bottom"
)

alpha_diversity_combined_no_rl

#ggsave(file.path("./06-publication", paste0(folder), paste0("SI_Fig_18CD_", file_modifier, "_no_rl_alpha_diversity_combined_plot_decontam.pdf")), plot = alpha_diversity_combined_no_rl, device = "pdf", scale = 1, width = 8, height = 8, units = c("in"), dpi = 300)
```


#Alpha Diversity - Read length subsampling
## Total Species
```{r}
#Kraken2 Total Species Table

#Difference in total number of species
kraken2_species_cuperdec_rl <- load_taxa_table(kraken2_total_species_rl)

#Calculate number of species reported
kraken2_total_species_table_rl <- kraken2_species_cuperdec_rl %>%
  group_by(Sample) %>%
  summarize(count_distinct_kraken2 = n_distinct(Taxon)) %>% #count how many unique taxa are reported in each library
  rename(Sequencing_ID = 1) %>%
  full_join(., metadata_rl) %>%
  drop_na()

```


## Oral Species
```{r}
#Oral species table
oral_species <- oral_db %>%
  filter(str_detect(Isolation_Source, "TRUE")) %>%
  rename(Species = 1) %>%
  dplyr::select(-Isolation_Source)
  
oral_species_table_rl <- kraken2_total_species_rl %>%
  inner_join(., oral_species, by = "Species") 

#fwrite(oral_species_table_rl, "./05-results/normalize_eager/dsvss_npre_read_length_kraken2_gtdb207_oral_species.tsv", sep = "\t", col.names = TRUE)
```


```{r}
#Calculate the number of classified reads to add to the metadata table

classified_reads_rl <- kraken2_total_species_rl %>%
  adorn_totals() %>%
  filter(Species == "Total") %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "classified_reads") %>%
  dplyr::select(-Species)

classified_reads_table_rl <- classified_reads_rl %>%
  full_join(metadata_rl %>%
              dplyr::select(Sequencing_ID, Strandedness, Strandedness_rl, read_count, kraken2_preservation)) %>%
  rename(total_reads = read_count,
         preservation = kraken2_preservation)

oral_reads_rl <- oral_species_table_rl %>%
  adorn_totals() %>%
  filter(Species == "Total") %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "oral_reads") %>%
  dplyr::select(-Species)

classified_reads_table_rl <- classified_reads_table_rl %>%
  full_join(., oral_reads_rl) %>%
  dplyr::select(Sequencing_ID, Strandedness, Strandedness_rl, preservation, total_reads, classified_reads, oral_reads)

```


Does read length subsampling cause a different proportion of reads to be classified?
```{r}
classified_reads_table_rl_summary <- classified_reads_table_rl %>%
  mutate(percent_classified = (classified_reads/total_reads)*100,
         percent_oral = (oral_reads/total_reads)*100) %>%
  group_by(Strandedness_rl) %>%
  summarize(median_percent_classified = median(percent_classified),
            sd_classified = sd(percent_classified),
            median_percent_oral = median(percent_oral),
            sd_oral = sd(percent_oral),
            median_total_reads = median(total_reads)) 
```


Count number of oral species per sample
```{r}
kraken2_oral_species_table_rl <- load_taxa_table(oral_species_table_rl)

kraken2_species_summary_table_rl <- kraken2_oral_species_table_rl %>%
  group_by(Sample) %>%
  summarize(count_distinct_kraken2_filtered = n_distinct(Taxon)) %>%
  dplyr::rename(Sequencing_ID = 1) %>%
  full_join(., metadata_rl) %>%
  drop_na() %>%
  full_join(., kraken2_total_species_table_rl) %>%
  dplyr::rename(Total = count_distinct_kraken2,
         Oral = count_distinct_kraken2_filtered) %>%
  relocate(Total, .after = Oral) %>%
  pivot_longer(Oral:Total, names_to = "pre_v_post", values_to = "count")

```

## Raincloud plot
Raincloud plot for total observed species
```{r}
kraken2_species_summary_table_rl <- kraken2_species_summary_table_rl %>%
  #filter(Individual_ID != "BAN001") %>%
  mutate(count = replace_na(count, 0)) %>%
  mutate(pre_v_post = str_replace(pre_v_post, "Total", "Total Species"),
         pre_v_post = str_replace(pre_v_post, "Oral", "Oral Species")) %>%
  mutate(Strandedness_rl = str_replace_all(Strandedness_rl, "Double_over50bp", "Double Over 50bp"),
         Strandedness_rl = str_replace_all(Strandedness_rl, "Single_over50bp", "Single Over 50bp"),
         Strandedness_rl = str_replace_all(Strandedness_rl, "Double_under50bp", "Double Under 50bp"),
         Strandedness_rl = str_replace_all(Strandedness_rl, "Single_under50bp", "Single Under 50bp"))

#Statistics - Total v. Oral (ds % ss)
stats_kraken_alpha_diversity_wilcox_strandedness_rl <- kraken2_species_summary_table_rl %>%
  group_by(Strandedness_rl) %>%
  wilcox_test(count ~ (pre_v_post), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

stats_kraken_alpha_diversity_effectsize_strandedness_rl <- kraken2_species_summary_table_rl %>%
  group_by(Strandedness_rl) %>%
  wilcox_effsize(formula = count ~ pre_v_post, paired = TRUE) 

#Stats for plot
wilcox_results <- kraken2_species_summary_table_rl %>%
  group_by(forcats::fct_rev(pre_v_post)) %>%
  wilcox_test(count ~ Strandedness_rl, paired = TRUE,
              comparisons = list(
                c("Double Over 50bp", "Double Under 50bp"),
                c("Double Over 50bp", "Single Over 50bp"),
                c("Double Over 50bp", "Single Under 50bp"),
                c("Single Over 50bp", "Double Under 50bp"),
                c("Single Over 50bp", "Single Under 50bp"),
                c("Double Under 50bp", "Single Under 50bp")
              )) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- kraken2_species_summary_table_rl %>%
  group_by(forcats::fct_rev(pre_v_post)) %>%
  wilcox_effsize(formula = count ~ Strandedness_rl, paired = TRUE,
                 comparisons = list(
                  c("Double Over 50bp", "Double Under 50bp"),
                  c("Double Over 50bp", "Single Over 50bp"),
                  c("Double Over 50bp", "Single Under 50bp"),
                  c("Single Over 50bp", "Double Under 50bp"),
                  c("Single Over 50bp", "Single Under 50bp"),
                  c("Double Under 50bp", "Single Under 50bp")
                )) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat)) %>%
  dplyr::rename(pre_v_post = 1) %>%
  mutate(y.position = c(525, 625, 575, 475, 425, 675,  # Total Species
                        300, 375, 450, 525, 600, 675))        # Oral Species

#PLOT
color <- c(double_color, single_color)

kraken2_raincloud_plot_rl <- kraken2_species_summary_table_rl %>%
  ggplot(., aes(x=Strandedness_rl, y=count)) +
  ggdist::stat_halfeye(
     aes(color = Strandedness,
        fill = after_scale(lighten(color, .2))),
    adjust = 0.5, 
    width = 0.6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
    ) + 
  ggdist::stat_dots(
    side = "left", 
    dotsize = 1.5, 
    justification = 1.175, 
    binwidth = 8,
    alpha = .3,
    range_scale = .4,
    aes(color = Strandedness, color = after_scale(darken(color, .1, space = "HLS")))
  ) +
  geom_boxplot(
    aes(color = Strandedness,
        color = after_scale(darken(color, .1, space = "HLS")),
        fill = after_scale(desaturate(lighten(color, .8), .4))),
    width = .15, 
    outlier.shape = NA
    ) +
  facet_grid(~forcats::fct_rev(pre_v_post)) +
  theme_minimal() +
  theme(
    legend.position = "none", 
    plot.title = element_text(hjust = 0.5)
    ) +
  labs(
    title = "", 
    y = "Observed Species", 
    x = ""
    ) +
  coord_cartesian(xlim = c(1.2, NA), clip = "off") +
  scale_color_manual(values = color, guide = "none") +
  scale_fill_manual(values = color, guide = "none") +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
            strip.text = element_text(size = 14)) +
  expand_limits(y = c(0, 700))+
  stat_pvalue_manual(
    wilcox_results,
    label = "annotation",
    y.position = "y.position",
    xmin = "group1",
    xmax = "group2",
    tip.length = 0.01,
    bracket.size = 0.5,
    size = 3.5
  )

kraken2_raincloud_plot_rl

```


## Shannon Diversity
Shannon Diversity plot
```{r}
#Calculate the Shannon diversity index
#Total Species
kraken2_total_species_formatted_rl <- kraken2_total_species_rl %>%
  pivot_longer(2:ncol(.), names_to = "Sample", values_to = "Counts") %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  column_to_rownames("Sample")

shannon_diversity_total_rl <- diversity(kraken2_total_species_formatted_rl, index="shannon") %>%
  as.data.frame() %>%
  dplyr::rename(shannon_index_total = 1) %>%
  rownames_to_column("Sequencing_ID")

#Oral Species
#Need to reformat species table so species are the columns and samples are the rows
kraken2_oral_species_formatted_rl <- oral_species_table_rl %>%
  pivot_longer(2:ncol(.), names_to = "Sample", values_to = "Counts") %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  column_to_rownames("Sample")

shannon_diversity_oral_rl <- diversity(kraken2_oral_species_formatted_rl, index="shannon") %>%
  as.data.frame() %>%
  dplyr::rename(shannon_index_oral = 1) %>%
  rownames_to_column("Sequencing_ID")

#Combine tables
metadata_w_shannon_rl <- metadata_rl %>%
  full_join(., shannon_diversity_total_rl, by = "Sequencing_ID") %>%
  full_join(., shannon_diversity_oral_rl, by = "Sequencing_ID") %>%
  pivot_longer(shannon_index_total:shannon_index_oral, names_to = "total_v_oral", values_to = "shannon_index") %>%
  mutate(Strandedness_rl = str_replace_all(Strandedness_rl, "Double_over50bp", "Double Over 50bp"),
         Strandedness_rl = str_replace_all(Strandedness_rl, "Single_over50bp", "Single Over 50bp"),
         Strandedness_rl = str_replace_all(Strandedness_rl, "Double_under50bp", "Double Under 50bp"),
         Strandedness_rl = str_replace_all(Strandedness_rl, "Single_under50bp", "Single Under 50bp")) 
  

#Statistics - Total v. Oral (ds + ss)
stats_kraken_alpha_diversity_wilcox_strandedness_rl <- metadata_w_shannon_rl %>%
  group_by(Strandedness_rl) %>%
  wilcox_test(shannon_index ~ (total_v_oral), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

stats_kraken_alpha_diversity_effectsize_strandedness_rl <- metadata_w_shannon_rl %>%
  group_by(Strandedness_rl) %>%
  wilcox_effsize(formula = shannon_index ~ total_v_oral, paired = TRUE) 

#Stats for plot
wilcox_results <- metadata_w_shannon_rl %>%
  group_by(total_v_oral) %>%
  wilcox_test(shannon_index ~ Strandedness_rl, paired = TRUE,
               comparisons = list(
                c("Double Over 50bp", "Double Under 50bp"),
                c("Double Over 50bp", "Single Over 50bp"),
                c("Double Over 50bp", "Single Under 50bp"),
                c("Single Over 50bp", "Double Under 50bp"),
                c("Single Over 50bp", "Single Under 50bp"),
                c("Double Under 50bp", "Single Under 50bp")
              )) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- metadata_w_shannon_rl %>%
  group_by(total_v_oral) %>%
  wilcox_effsize(formula = shannon_index ~ Strandedness_rl, paired = TRUE,
                   comparisons = list(
                c("Double Over 50bp", "Double Under 50bp"),
                c("Double Over 50bp", "Single Over 50bp"),
                c("Double Over 50bp", "Single Under 50bp"),
                c("Single Over 50bp", "Double Under 50bp"),
                c("Single Over 50bp", "Single Under 50bp"),
                c("Double Under 50bp", "Single Under 50bp")
              )) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat)) %>%
  mutate(total_v_oral = str_replace(total_v_oral, "shannon_index_oral", "Oral Species"),
         total_v_oral = str_replace(total_v_oral, "shannon_index_total", "Total Species")) %>%
  mutate(y.position = c(5, 5.4, 5.8, 6.2, 6.6, 7,  # Oral Species
                        5, 5.4, 5.8, 6.2, 6.6, 7))  # Total Species

#Plot
kraken2_raincloud_plot_shannon_rl <- metadata_w_shannon_rl %>%
  mutate(total_v_oral = str_replace_all(total_v_oral, "shannon_index_oral", "Oral Species"),
         total_v_oral = str_replace_all(total_v_oral, "shannon_index_total", "Total Species")) %>%
  ggplot(., aes(x=Strandedness_rl, y=shannon_index)) +
  ggdist::stat_halfeye(
     aes(color = Strandedness,
        fill = after_scale(lighten(color, .2))),
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
    ) + 
  ggdist::stat_dots(
    side = "left", 
    dotsize = 0.2, 
    justification = 1.175, 
    binwidth = .6,
    alpha = .3,
    range_scale = .4,
    aes(color = Strandedness, color = after_scale(darken(color, .1, space = "HLS")))
  ) +
  geom_boxplot(
    aes(color = Strandedness,
        color = after_scale(darken(color, .1, space = "HLS")),
        fill = after_scale(desaturate(lighten(color, .8), .4))),
    width = .15, 
    outlier.shape = NA
    ) +
  facet_grid(~forcats::fct_rev(total_v_oral)) +
  theme_minimal() +
  theme(
    legend.position = "none", 
    plot.title = element_text(hjust = 0.5)
    ) +
  labs(
    title = "", 
    y = "Shannon Index", 
    x = ""
    ) +
  coord_cartesian(xlim = c(1.2, NA), clip = "off") +
  scale_color_manual(values = color, guide = "none") +
  scale_fill_manual(values = color, guide = "none") +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
            strip.text = element_text(size = 14)) +
  expand_limits(y = c(0, 7)) +
  stat_pvalue_manual(
    wilcox_results,
    label = "annotation",
    y.position = "y.position",
    xmin = "group1",
    xmax = "group2",
    tip.length = 0.01,
    bracket.size = 0.5,
    size = 3
  )
  

kraken2_raincloud_plot_shannon_rl

```


Combine the Observed species and Shannon diversity plots
```{r}
alpha_diversity_combined_rl <- ggarrange(
  kraken2_raincloud_plot_rl, kraken2_raincloud_plot_shannon_rl, 
  labels = c("A", "B"), 
  ncol = 1, nrow = 2, 
  common.legend = TRUE, 
  legend = "bottom"
)

alpha_diversity_combined_rl

#ggsave(file.path("./06-publication", paste0(folder), paste0("SI_Fig_26AB_", file_modifier, "_rl_subsampling_alpha_diversity_combined_plot.pdf")), plot = alpha_diversity_combined_rl, device = "pdf", scale = 1, width = 12, height = 10, units = c("in"), dpi = 300)
```



# Beta Diversity - No read length subsampling

## Total Species
Create a PCA of total number of species
```{r}
kraken2_total_species_no_rl_edited <- kraken2_total_species_no_rl %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "Counts") %>%
  mutate(Counts = Counts + 1) %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  filter(!str_detect(Sequencing_ID, "Isolation_Source")) %>%
  column_to_rownames("Sequencing_ID")
  
result.pca.kraken2_no_rl <- pca(kraken2_total_species_no_rl_edited, logratio = "CLR") 

exp_var <- paste0(round(result.pca.kraken2_no_rl$prop_expl_var$X * 100, 2), "%") # explained variance for x- and y-labels

plotIndiv(result.pca.kraken2_no_rl)

pca_kraken2_no_rl <- result.pca.kraken2_no_rl$variates %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  dplyr::rename(Sequencing_ID = 1,
         PC1 = 2,
         PC2 = 3) %>%
  inner_join(., metadata_no_rl, by = "Sequencing_ID")

permanova_dist_matrix <- dist(pca_kraken2_no_rl[, c("PC1", "PC2")], method = "euclidean")
adonis_result <- adonis2(permanova_dist_matrix ~ Strandedness, data = pca_kraken2_no_rl) #Strandedness, Individual_ID, kraken2_preservation
r_squared_value <- adonis_result$R2[1] 
p_value <- adonis_result$`Pr(>F)`[1]
p_value_formatted <- ifelse(p_value < 0.001, "< 0.001",
                            as.character(signif(p_value, digits = 2)))
r_squared_value_formatted <- ifelse(r_squared_value < 0.001, "< 0.001",
                            as.character(signif(r_squared_value, digits = 2)))


#keep color for strandedness, and change shape for preservation 
color <- c(double_color, single_color)


kraken2_pca_strandedness_no_rl <- pca_kraken2_no_rl %>%
  mutate(kraken2_preservation = factor(kraken2_preservation, levels = c("Good", "Bad"))) %>%
  ggplot(., aes(x = PC1, y = PC2, color = Strandedness)) +
    geom_point(aes(shape = kraken2_preservation), size = 4) +
    theme_minimal() +
    labs(title = "", y = paste("PC2", " - ", exp_var[2]), x = paste("PC1", " - ", exp_var[1])) +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_color_discrete(name = "Strandedness", labels = c("Double", "Single")) +
    scale_shape_discrete(name = "Preservation", labels = c("Good", "Bad")) +
    geom_line(aes(group = Individual_ID), color = "#555555ff", linetype = "dashed") +
    #scale_colour_manual(name = "Strandedness", values = c("Double" = after_scale(lighten("#4059AD", .2)), "Single" = after_scale(lighten("#5fc3a4", .2)))) +
  scale_colour_manual(name = "Strandedness", values = color) +
  theme(axis.text.x = element_text(size = 12),  
        axis.text.y = element_text(size = 12),
        strip.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        legend.position = "right") +
        geom_text_repel(data = subset(pca_kraken2_no_rl, Strandedness == "Double"), 
                        aes(label = Individual_ID), 
                        size = 3, color = "black", segment.color = NA,
                        nudge_y = -0.8) +
        coord_equal() +
    annotate("text", x = 30, y = -40,
           label = paste("R2 =", r_squared_value_formatted,
                         "\np-value =", p_value_formatted))
 
kraken2_pca_strandedness_no_rl


```


```{r}
#Define colors and shapes for plotting
Strandedness_colors <- c("Double" = double_color, "Single" = single_color)
kraken2_preservation_shapes <- c("Good" = 16, "Bad" = 17)

#Biplot Function
plot_pca_bi <- function(df, pc1, pc2, metadata_group1, metadata_group2, columntitle) {
    metadata_group_colors <- get(paste(metadata_group1, "_colors", sep = ""))
    metadata_group_shapes <- get(paste(metadata_group2, "_shapes", sep = ""))
   
    arrow_pc <- enquo(columntitle)
    
    exp_var <- paste0(round(df$prop_expl_var$X * 100, 2), "%") # explained variance for x- and y-labels
    
    # select only the PCs from the PCA and add metadata
    df_X <- df$variates$X %>%
              as.data.frame() %>%
              rownames_to_column("Sequencing_ID") %>%
              inner_join(metadata, by = "Sequencing_ID")
    
    metadata_group1 = df_X[[metadata_group1]]
    metadata_group2 = df_X[[metadata_group2]]
    
    corr_lam <- df$sdev[c("PC1", "PC2")] * sqrt(nrow(df_X))

    df_X <- df_X %>%
      mutate(PC1 = PC1 / corr_lam[1],
             PC2 = PC2 / corr_lam[2])

    # select the correct PC column and explained variance for PC1
    if (pc1 == 'PC1') {
        Pc1 <- df_X$PC1
        exp_var_pc1 <- exp_var[1]
        xaxis <- c("PC1")
    } else if (pc1 == 'PC2') {
        Pc1 <- df_X$PC2
        exp_var_pc1 <- exp_var[2]
        xaxis <- c("PC2")
    } 
    
    # select the correct PC column and explained variance for PC2
    if (pc2 == 'PC1') {
        Pc2 <- df_X$PC1
        exp_var_pc2 <- exp_var[1]
        yaxis <- c("PC1")
 } else if (pc2 == 'PC2') {
       Pc2 <- df_X$PC2
       exp_var_pc2 <- exp_var[2]
       yaxis <- c("PC2")
    }
    
    # Identify the 10 pathways that have highest positive and negative loadings in the selected PC
    pws_10 <- df$loadings$X %>%
      as.data.frame(.) %>%
      rownames_to_column(var = "Taxon") %>%
      top_n(10, !!arrow_pc) %>%
      arrange(desc(!!arrow_pc)) %>%   #sort by size of loading to get strongest first
      mutate(label = row_number())

    neg_10 <- df$loadings$X %>%
      as.data.frame(.) %>%
      rownames_to_column(var = "Taxon") %>%
      top_n(-10, !!arrow_pc) %>%
      arrange(!!arrow_pc) %>%  #sort to get the smallest (most negative/strongest) first
      mutate(label = row_number())

    pca_plot_bi <- df_X %>%
      mutate(metadata_group2 = factor(metadata_group2, levels = c("Good", "Bad"))) %>%
      ggplot(., aes(x = Pc1, y = Pc2, color = metadata_group1)) +
      geom_point(size = 4, aes(shape = metadata_group2) )+
      geom_segment(data = pws_10,
                   aes(xend = get(paste(pc1)), yend = get(paste(pc2))),
                   x = 0, y = 0, colour = "black",
                   size = 0.3,
                   arrow = arrow(length = unit(0.03, "npc"))) +
      #geom_label_repel(data = pws_10,
      #             aes(x = get(paste(pc1)), y = get(paste(pc2)), label = label), #using label instead of taxon
      #             size = 2, 
      #             colour = "black",
      #             label.padding = 0.1, 
      #             box.padding = 0.1, #
      #             point.padding = 0.1, 
      #             force = 10,
      #             max.overlaps = Inf) + 
      geom_segment(data = neg_10,
                   aes(xend = get(paste(pc1)), yend = get(paste(pc2))),
                   x = 0, y = 0, colour = "grey50",
                   size = 0.3,
                   arrow = arrow(length = unit(0.03, "npc"))) +
      #geom_label_repel(data = neg_10,
      #             aes(x = get(paste(pc1)), y = get(paste(pc2)), label = label), #using label instead of taxon
      #             size = 2, 
      #             colour = "red",
      #             label.padding = 0.1, 
      #             box.padding = 0.1, 
      #             point.padding = 0.1, 
      #             force = 10, 
      #             max.overlaps = Inf) + 
      labs(x = paste(xaxis, " - ", exp_var_pc1),
           y = paste(yaxis, " - ", exp_var_pc2),
           color = "Strandedness",
           shape = "Preservation") +
      scale_color_manual(values = metadata_group_colors) +
      scale_shape_manual(values = metadata_group_shapes) +
      theme_minimal() + theme(text = element_text(size = 16)) +
      theme(text = element_text(size=16)) +
      theme(legend.position = "bottom") +
      geom_line(aes(group = Individual_ID), color = "#555555ff", linetype = "dashed") +
      geom_text_repel(aes(label = Individual_ID), color = "black", segment.color = NA) +
      coord_equal()
      

    return(pca_plot_bi) #pca_plot_bi, pws_10, or neg_10
}

pca_kraken2_biplot_PC1_no_rl <- plot_pca_bi(result.pca.kraken2_no_rl, "PC1", "PC2", "Strandedness", "kraken2_preservation", PC1)
pca_kraken2_biplot_PC2_no_rl <- plot_pca_bi(result.pca.kraken2_no_rl, "PC1", "PC2", "Strandedness", "kraken2_preservation", PC2)

pca_kraken2_biplot_PC1_no_rl
pca_kraken2_biplot_PC2_no_rl

pca_biplots_combined_no_rl <- ggarrange(
  pca_kraken2_biplot_PC1_no_rl, pca_kraken2_biplot_PC2_no_rl, 
  labels = c("A", "B"), 
  ncol = 1, nrow = 2, 
  common.legend = TRUE, 
  legend = "bottom")

pca_biplots_combined_no_rl

#ggsave(file.path("./06-publication", paste0(folder), paste0("SI_Fig19AB_", file_modifier, "_no_rl_kraken2_pca_biplot_combined_PCs_total_species_decontam.pdf")) , plot = pca_biplots_combined_no_rl, device = "pdf", scale = 1, width = 8, height = 12, units = c("in"), dpi = 300)
```


## Oral Species
```{r}
#Fill in data for total oral species v. filtered oral species w/ no contam
kraken2_oral_species_no_rl <- fread("./05-results/normalize_eager/dsvss_npre_no_rl_kraken2_gtdb207_oral_species_decontam_filtered.tsv", header = TRUE) 

data <- kraken2_oral_species_no_rl #select the oral set to use

metadata <- fread("./05-results/normalize_eager/dsvss_npre_w_rl_subsampling_metadata.tsv") 

metadata_no_rl <- metadata %>%
  filter(Read_length_subsampling == "none")
```


```{r}
kraken2_oral_species_no_rl <- data %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "Counts") %>%
  mutate(Counts = Counts + 1) %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  filter(!str_detect(Sequencing_ID, "Isolation_Source")) %>%
  column_to_rownames("Sequencing_ID")
  
result.pca.kraken2_no_rl <- pca(kraken2_oral_species_no_rl, logratio = "CLR") 

exp_var <- paste0(round(result.pca.kraken2_no_rl$prop_expl_var$X * 100, 2), "%") # explained variance for x- and y-labels

pca_kraken2_no_rl <- result.pca.kraken2_no_rl$variates %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  dplyr::rename(Sequencing_ID = 1,
         PC1 = 2,
         PC2 = 3) %>%
  inner_join(., metadata_no_rl, by = "Sequencing_ID")
```

Perform a PERMANOVA (Permutational Multivariate Analysis of Variance) test using the adonis2 function from the vegan package
```{r}
#Create a metadata file with sample IDs, grouping info, and PCA scores
permanova_data_no_rl <- pca_kraken2_no_rl %>%
  dplyr::select(Sequencing_ID, Individual_ID, PC1, PC2, Strandedness, kraken2_preservation) %>%
  column_to_rownames("Sequencing_ID")

#Create a distance matrix of just PCA scores
permanova_dist_matrix_no_rl <- dist(permanova_data_no_rl[, c("PC1", "PC2")], method = "euclidean")

#Run adonis2
adonis_result <- adonis2(permanova_dist_matrix_no_rl ~ Strandedness, data = permanova_data_no_rl) #Strandedness, Individual_ID, kraken2_preservation

r_squared_value <- adonis_result$R2[1] 
  
p_value <- adonis_result$`Pr(>F)`[1]

p_value_formatted <- ifelse(p_value < 0.001, "< 0.001",
                            as.character(signif(p_value, digits = 2)))

r_squared_value_formatted <- ifelse(r_squared_value < 0.001, "< 0.001",
                            as.character(signif(r_squared_value, digits = 2)))
```


```{r}
#keep color for strandedness, and change shape for preservation 
color <- c(double_color, single_color)

kraken2_pca_strandedness_no_rl <- pca_kraken2_no_rl %>%
  mutate(kraken2_preservation = factor(kraken2_preservation, levels = c("Good", "Bad"))) %>%
  ggplot(., aes(x = PC1, y = PC2, color = Strandedness)) +
    geom_point(aes(shape = kraken2_preservation), size = 4) +
    theme_minimal() +
    labs(title = "", y = paste("PC2", " - ", exp_var[2]), x = paste("PC1", " - ", exp_var[1])) +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_color_discrete(name = "Strandedness", labels = c("Double", "Single")) +
    scale_shape_discrete(name = "Preservation") +
    geom_line(aes(group = Individual_ID), color = "#555555ff", linetype = "dashed") +
    #scale_colour_manual(name = "Strandedness", values = c("Double" = after_scale(lighten("#4059AD", .2)), "Single" = after_scale(lighten("#5fc3a4", .2)))) +
    #scale_color_manual(name = "Strandedness", values = c("Double" = "#098BD9", "Single" = "#7DCCFF")) +
    scale_color_manual(name = "Strandedness", values = color) +
    theme(axis.text.x = element_text(size = 12),  
          axis.text.y = element_text(size = 12),
          strip.text = element_text(size = 12),
          legend.text = element_text(size = 12),
          legend.title = element_text(size = 12),
          axis.title.x = element_text(size = 12),
          axis.title.y = element_text(size = 12),
          legend.position = "bottom") +
    geom_text_repel(data = subset(pca_kraken2_no_rl, Strandedness == "Double"), 
                        aes(label = Individual_ID), 
                        size = 3, color = "black", segment.color = NA,
                        nudge_y = -0.8) +
    coord_equal() +
  annotate("text", x = -14, y = 15,
           label = paste("R2 =", r_squared_value_formatted,
                         "\np-value =", p_value_formatted)) 
 
kraken2_pca_strandedness_no_rl

```

### Biplot
```{r}
kraken2_oral_species_no_rl <- data %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "Counts") %>%
  mutate(Counts = Counts + 1) %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  filter(!str_detect(Sequencing_ID, "Isolation_Source")) %>%
  column_to_rownames("Sequencing_ID")

#Create PCA using mixomics package
result.pca.kraken2_no_rl <- pca(kraken2_oral_species_no_rl, ncomp = 2, logratio = "CLR") 

#Metadata table (will be added w/in the function)
full_metadata <- fread("./05-results/normalize_eager/dsvss_npre_w_rl_subsampling_metadata.tsv") %>%
  filter(Read_length_subsampling == "none") %>%
  dplyr::select(Sequencing_ID, Strandedness, kraken2_preservation, Individual_ID)

#Define colors and shapes for plotting
Strandedness_colors <- c("Double" = double_color, "Single" = single_color)
kraken2_preservation_shapes <- c("Good" = 16, "Bad" = 17)

#Biplot Function
plot_pca_bi <- function(df, pc1, pc2, metadata_group1, metadata_group2, columntitle) {
    metadata_group_colors <- get(paste(metadata_group1, "_colors", sep = ""))
    metadata_group_shapes <- get(paste(metadata_group2, "_shapes", sep = ""))
   
    arrow_pc <- enquo(columntitle)
    
    exp_var <- paste0(round(df$prop_expl_var$X * 100, 2), "%") # explained variance for x- and y-labels
    
    # select only the PCs from the PCA and add metadata
    df_X <- df$variates$X %>%
              as.data.frame() %>%
              rownames_to_column("Sequencing_ID") %>%
              inner_join(full_metadata, by = "Sequencing_ID")
    
    metadata_group1 = df_X[[metadata_group1]]
    metadata_group2 = df_X[[metadata_group2]]
    
    corr_lam <- df$sdev[c("PC1", "PC2")] * sqrt(nrow(df_X))

    df_X <- df_X %>%
      mutate(PC1 = PC1 / corr_lam[1],
             PC2 = PC2 / corr_lam[2])

    # select the correct PC column and explained variance for PC1
    if (pc1 == 'PC1') {
        Pc1 <- df_X$PC1
        exp_var_pc1 <- exp_var[1]
        xaxis <- c("PC1")
    } else if (pc1 == 'PC2') {
        Pc1 <- df_X$PC2
        exp_var_pc1 <- exp_var[2]
        xaxis <- c("PC2")
    } 
    
    # select the correct PC column and explained variance for PC2
    if (pc2 == 'PC1') {
        Pc2 <- df_X$PC1
        exp_var_pc2 <- exp_var[1]
        yaxis <- c("PC1")
 } else if (pc2 == 'PC2') {
       Pc2 <- df_X$PC2
       exp_var_pc2 <- exp_var[2]
       yaxis <- c("PC2")
    }
    
    # Identify the 10 pathways that have highest positive and negative loadings in the selected PC
    pws_10 <- df$loadings$X %>%
      as.data.frame(.) %>%
      rownames_to_column(var = "Taxon") %>%
      top_n(10, !!arrow_pc) %>%
      arrange(desc(!!arrow_pc)) %>%   #sort by size of loading to get strongest first
      mutate(label = row_number())

    neg_10 <- df$loadings$X %>%
      as.data.frame(.) %>%
      rownames_to_column(var = "Taxon") %>%
      top_n(-10, !!arrow_pc) %>%
      arrange(!!arrow_pc) %>%  #sort to get the smallest (most negative/strongest) first
      mutate(label = row_number())

    pca_plot_bi <- df_X %>%
      mutate(metadata_group2 = factor(metadata_group2, levels = c("Good", "Bad"))) %>%
      ggplot(., aes(x = Pc1, y = Pc2, color = metadata_group1)) +
      geom_point(size = 4, aes(shape = metadata_group2) )+
      geom_segment(data = pws_10,
                   aes(xend = get(paste(pc1)), yend = get(paste(pc2))),
                   x = 0, y = 0, colour = "black",
                   size = 0.3,
                   arrow = arrow(length = unit(0.03, "npc"))) +
      #geom_label_repel(data = pws_10,
      #             aes(x = get(paste(pc1)), y = get(paste(pc2)), label = label), #using label instead of taxon
      #             size = 2, 
      #             colour = "black",
      #             label.padding = 0.1, 
      #             box.padding = 0.1, #
      #             point.padding = 0.1, 
      #             force = 10,
      #             max.overlaps = Inf) + 
      geom_segment(data = neg_10,
                   aes(xend = get(paste(pc1)), yend = get(paste(pc2))),
                   x = 0, y = 0, colour = "grey",
                   size = 0.3,
                   arrow = arrow(length = unit(0.03, "npc"))) +
      #geom_label_repel(data = neg_10,
      #             aes(x = get(paste(pc1)), y = get(paste(pc2)), label = label), #using label instead of taxon
      #             size = 2, 
      #             colour = "red",
      #             label.padding = 0.1, 
      #             box.padding = 0.1, 
      #             point.padding = 0.1, 
      #             force = 10, 
      #             max.overlaps = Inf) + 
      labs(x = paste(xaxis, " - ", exp_var_pc1),
           y = paste(yaxis, " - ", exp_var_pc2),
           color = "Strandedness",
           shape = "Preservation") +
      scale_color_manual(values = metadata_group_colors) +
      scale_shape_manual(values = metadata_group_shapes) +
      theme_minimal() + theme(text = element_text(size = 16)) +
      theme(text = element_text(size=16)) +
      theme(legend.position = "bottom") +
      geom_line(aes(group = Individual_ID), color = "#CCCCCC", linetype = "dashed") +
      geom_text_repel(aes(label = Individual_ID), color = "black", segment.color = NA) +
      coord_equal()
      
    return(pca_plot_bi) #pca_plot_bi, pws_10, neg_10
}

pca_kraken2_biplot_PC1 <- plot_pca_bi(result.pca.kraken2_no_rl, "PC1", "PC2", "Strandedness", "kraken2_preservation", PC1)
pca_kraken2_biplot_PC2 <- plot_pca_bi(result.pca.kraken2_no_rl, "PC1", "PC2", "Strandedness", "kraken2_preservation", PC2)

pca_kraken2_biplot_PC1
pca_kraken2_biplot_PC2

oral_pca_biplots_combined_no_rl <- ggarrange(
  pca_kraken2_biplot_PC1, pca_kraken2_biplot_PC2, 
  labels = c("A", "B"), 
  ncol = 1, nrow = 2, 
  common.legend = TRUE, 
  legend = "bottom")

oral_pca_biplots_combined_no_rl

#ggsave(file.path("./06-publication", paste0(folder), paste0("SI_Fig19CD_", file_modifier, "_no_rl_kraken2_pca_biplot_combined_PCs_oral_species_decontam.pdf")), plot = oral_pca_biplots_combined_no_rl, device = "pdf", scale = 1, width = 8, height = 12, units = c("in"), dpi = 300)
```


# Sources
```{r}
#Read in sources kraken2 species table
kraken2_sources <- fread("./05-results/dsvss_sources_controls_kraken2_gtdb207_total_species.tsv") 

#Remove controls samples so it is just the source samples
control_sample_list <- fread("./05-results/dsvss_controls_sourcetracker_metadata.tsv") %>%
  dplyr::select(Library_ID) %>%
  pull()

kraken2_sources <- kraken2_sources %>%
  dplyr::select(-all_of(control_sample_list))

#Filter with a 0.005% relative abundance cut-off and clean up the species names

kraken2_sources_totals <-  kraken2_sources %>%
  column_to_rownames("Species") %>%
   mutate(Total = rowSums(across(where(is.numeric)))) %>%
  rownames_to_column("Species") %>%
  relocate("Total", .after = Species) 

total_sum <- sum(kraken2_sources_totals$Total)

kraken2_sources_abundance <- kraken2_sources_totals %>%
  mutate(Relative_Abundance = (Total / total_sum)*100) %>% 
  relocate(Relative_Abundance, .after = Species)  

kraken2_sources_filtered <- kraken2_sources_abundance %>%
  filter(Relative_Abundance >= 0.005) %>% #using a 0.005% cut-off
  dplyr::select(-Total, -Relative_Abundance) %>%
  mutate(Species = str_replace(Species, " (?:[A-Z]+)$", "")) %>%
  group_by(Species) %>%
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) %>%
  mutate(across(-Species, ~ if_else(.x < 10, 0, .x)))
```


```{r}
#read in eager 10M no read length subsampling total species table
kraken2_total_species <- fread("./05-results/normalize_eager/dsvss_npre_w_read_length_kraken2_gtdb207_total_species_decontam_filtered.tsv") 
  
no_rl_samples <- metadata_no_rl %>%
  dplyr::select(Sequencing_ID) %>%
  pull()

kraken2_total_species_no_rl <- kraken2_total_species %>%
  column_to_rownames("Species") %>%
  dplyr::select(all_of(no_rl_samples)) %>%
  rownames_to_column("Species")

#Combine with total species kraken2 table
combined_kraken2_species <- full_join(kraken2_total_species_no_rl, kraken2_sources_filtered, by = "Species") %>%
  mutate(across(everything(), ~ replace_na(.x, 0)))

#Read in source metadata
mapfile <- fread("./05-results/normalize_eager/dsvss_npre_source_tracker_mappingfile_complete.tsv") %>%
  dplyr::rename(Sequencing_ID = 1) %>%
  dplyr::select(Sequencing_ID, Study, Env, SourceSink) %>%
  filter(!str_detect(Sequencing_ID, "bp"))
  
metadata_full <- metadata_no_rl %>%
  full_join(., mapfile) %>%
  mutate(Control_or_Sample = case_when(grepl("source", SourceSink) ~ "Control", TRUE ~ "Sample")) 
```



## Rarefied PCoA
To resolve the issue with samples clustering by the number of reads and low read length, try making a rarefied PCoA instead
```{r}
#Convert to a phyloseq object to rarefy the table before creating the PCoA
library(phyloseq)

#Make the species column rownames
combined_kraken2_species_formatted <- combined_kraken2_species %>%
  select(-ERR1883424, -ERR1883437) %>% # filter out samples with a species sum of 0 after filtering
  column_to_rownames("Species")

#Convert the species table to a matrix
species_matrix <- as.matrix(combined_kraken2_species_formatted)

#Convert the matrix to a phyloseq object
ps <- phyloseq(otu_table(species_matrix, taxa_are_rows=TRUE))

#Format metadata correctly
metadata_formatted <- metadata_full %>%
  dplyr::rename(SampleID = Sequencing_ID) %>%
  column_to_rownames("SampleID")

#Convert metadata to a phyloseq object
sample_data_obj <- sample_data(metadata_formatted)

#Add the metadata to the phyloseq object
ps <- merge_phyloseq(ps, sample_data_obj)

#Rarefy the species table
#Extract the sample data as a data frame
species_table <- otu_table(ps)
sample_sums <- sample_sums(species_table)

#Find the minimum sample sum
min_sample_sum <- min(sample_sums)

#Rarefy
physeq_rarefied <- rarefy_even_depth(species_table, sample.size = min_sample_sum, rngseed = 1)

#Convert to distance matrix (Jaccard dissimilarity)
physeq_rarefied_formatted <- as.data.frame(physeq_rarefied)

physeq_rarefied_formatted <- physeq_rarefied_formatted %>%
  rownames_to_column("Species") %>%
  pivot_longer(2:ncol(.), names_to = "SampleID", values_to = "counts") %>%
  pivot_wider(names_from = "Species", values_from = "counts") %>%
  arrange(SampleID) %>%
  column_to_rownames("SampleID") 

dist_matrix <- vegdist(physeq_rarefied_formatted, method = "jaccard")

#PCoA Calculation
pcoa_result <- ape::pcoa(dist_matrix)
pcoa_coordinates <- pcoa_result$vectors[, 1:2]
colnames(pcoa_coordinates) <- c("PCoA1", "PCoA2")

#Combine PCoA coordinates with metadata
pcoa_df <- as.data.frame(pcoa_coordinates)
pcoa_df$Sequencing_ID <- rownames(pcoa_df)
pcoa_df <- merge(pcoa_df, metadata_full, by = "Sequencing_ID")

#Variation explained by each axis
eig_values <- pcoa_result$values$Relative_eig
var_explained <- round(eig_values[1:2] * 100, digits = 2)
x_lab <- paste0("PCoA1 - ", round(var_explained[1], 2), "%")
y_lab <- paste0("PCoA2 - ", round(var_explained[1], 2), "%")

#Permanova
adonis_result <- adonis2(dist_matrix ~ Env, data = pcoa_df)

r_squared_value <- adonis_result$R2[1] 
p_value <- adonis_result$`Pr(>F)`[1]
p_value_formatted <- ifelse(p_value < 0.001, "< 0.001",
                            as.character(signif(p_value, digits = 2)))
r_squared_value_formatted <- ifelse(r_squared_value < 0.001, "< 0.001",
                            as.character(signif(r_squared_value, digits = 2)))

#Define colors, shapes, and plot

#colors <- c("Ancient Calculus - ssLib" = single_color, "Ancient Calculus - dsLib" = double_color, "Modern Calculus" = microshades_palette("micro_purple")[2], "Plaque" = microshades_palette("micro_purple")[5], "Gut" = "#B78560", "Skin" = "#fdae6b", "Environment" = "#74c476")

colors <- c("Ancient Calculus - ssLib" = single_color, "Ancient Calculus - dsLib" = double_color,
  "Modern Calculus" = "#bfaed4",  
  "Plaque"          = "#3f007d",  
  "Gut"             = "#c05b16",  
  "Skin"            = "#fdc187",  
  "Environment"     = "#7fc97f",  
  "Unknown"         = "#d9d9d9"   
)

shapes <- c("Double" = 16, "Source" = 15, "Single" = 17)

sources_pcoa <- pcoa_df %>%
  mutate(Env = str_replace_all(Env, "calculus", "Ancient Calculus"),
         Env = str_replace_all(Env, "environment", "Environment"),
         Env = str_replace_all(Env, "modern_calc", "Modern Calculus"),
         Env = str_replace_all(Env, "skin", "Skin"),
         Env = str_replace_all(Env, "gut", "Gut"),
         Env = str_replace_all(Env, "plaque", "Plaque")) %>%
  mutate(Env = case_when(Strandedness == "Single" ~ "Ancient Calculus - ssLib", TRUE ~ Env)) %>%
  mutate(Env = case_when(Strandedness == "Double" ~ "Ancient Calculus - dsLib", TRUE ~ Env)) %>%
  mutate(SourceSink = str_replace_all(SourceSink, "source", "Source"),
         SourceSink = str_replace_all(SourceSink, "sink", "Sink")) %>%
  mutate(Individual_ID = case_when(is.na(Individual_ID) ~ Library_ID, TRUE ~ Individual_ID)) %>%
  mutate(Strandedness = case_when(is.na(Strandedness) ~ "Source", TRUE ~ Strandedness)) %>%
  mutate(Env = factor(Env, levels = c("Ancient Calculus - dsLib", "Ancient Calculus - ssLib", "Modern Calculus", "Plaque", "Gut", "Skin", "Environment"))) %>%
  ggplot(., aes(x = PCoA1, y = PCoA2, color = Env, shape = Strandedness)) +
    geom_point(aes(alpha = Control_or_Sample), size = 2.5) +
    theme_minimal() +
    labs(title = "") +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_color_manual(name = "Environment", values = colors) +
    scale_shape_manual(name = "Sample", values = shapes) +
    #geom_line(aes(group = Individual_ID), color = "#CCCCCC", linetype = "dashed") +
    theme(axis.text.x = element_text(size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14)) +
  scale_alpha_manual(values = c("Control" = 0.7, "Sample" = 1)) +
  xlab(x_lab) +
  ylab(y_lab) +
  guides(alpha = "none") +
  geom_line(data = pcoa_df %>% filter(!is.na(Individual_ID)), 
              aes(group = Individual_ID), 
              color = "#555555ff", 
              linetype = "dashed") +
  geom_text(data = pcoa_df %>% filter(Strandedness %in% c("Single")), 
              aes(label = Individual_ID), 
              hjust = -0.3, 
              vjust = -0.3, 
              size = 3, color = "black") +
  coord_equal() +
  annotate("text", x = 0.2, y =  0.25,
           label = paste("R2 =", r_squared_value_formatted,
                         "\np-value =", p_value_formatted))

sources_pcoa

#ggsave(file.path("./06-publication", paste0(folder), paste0("SI_Fig7B_", file_modifier, "_no_rl_kraken2_sources_total_species_pcoa_rarefied.pdf")), plot = sources_pcoa, device = "pdf", scale = 1, width = 8, height = 5, units = c("in"), dpi = 300)
```

## Sourcetracker
### No read length subsampling
```{r}
source_colors <- c(
  "Modern Calculus" = "#bfaed4",  
  "Plaque"          = "#3f007d",  
  "Gut"             = "#c05b16",  
  "Skin"            = "#fdc187",  
  "Environment"     = "#7fc97f",  
  "Unknown"         = "#d9d9d9"   
)

source_patterns <- c(
  "Unknown"         = "none",
  "Environment"     = "stripe",
  "Skin"            = "stripe",
  "Gut"             = "stripe",
  "Plaque"          = "none",
  #"Modern Calculus" = "crosshatch"
  "Modern Calculus" = "none"
)

source_angles <- c(
  "Unknown"         = 0,
  "Environment"     = 45,
  "Skin"            = 135,
  "Gut"             = 0,
  "Plaque"          = 0,
  #"Modern Calculus" = 45
  "Modern Calculus" = 0
)


#Prepare SourceTracker Results
source_results <- fread("./05-results/normalize_eager/dsvss_npre_sourcetracker_sink_predictions.txt") %>%
  rename(Sequencing_ID = 1) %>%
  full_join(., metadata_no_rl)

library(ggpattern)

sources_plot_no_subsampling_individual_facet <- source_results %>%
  rename(Environment = environment, Gut = gut, "Modern Calculus" = modern_calc, Plaque = plaque, Skin = skin) %>%
  gather("Sources", "Percent",2:7) %>%
  mutate(Percent = Percent * 100) %>%
  mutate(Sources = fct_relevel(Sources, "Unknown", "Environment", "Skin","Gut","Plaque","Modern Calculus")) %>%
  ggplot(., aes(x = Strandedness, y = Percent, fill = Sources, pattern = Sources)) +
          geom_bar_pattern(
            aes(pattern_angle = Sources),
            linewidth = .02,
            stat = "identity",
            color = "black",
            #pattern_alpha = 0.45,
            pattern_fill = NA,
            pattern_density = 0.03,
            pattern_spacing = 0.07,
            pattern_size = 0.15,
            pattern_key_scale_factor = 0.2) +
          theme_minimal(base_size = 14) +
          scale_fill_manual(values = source_colors) +
          scale_pattern_manual(values = source_patterns) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1),
                axis.text = element_text(size = 8)) +
          geom_hline(yintercept=75, linetype="dotted") +
          geom_hline(yintercept=90, linetype="dotted", color = "gray50") +
          ylab("% attributed to source") +
          xlab("Individual ID and Strandedness") +
          facet_wrap(~Individual_ID, ncol = 7) +
          scale_pattern_angle_manual(values = source_angles)

sources_plot_no_subsampling_individual_facet

```

### Read length subsampling
```{r}
sourcetracker_10M <- fread("./05-results/normalize_eager/dsvss_npre_sourcetracker_sink_predictions.txt")

sourcetracker_10M_over50bp <- fread("./05-results/normalize_eager/dsvss_npre_over50bp_sourcetracker_sink_predictions.txt")

sourcetracker_10M_under50bp <- fread("./05-results/normalize_eager/dsvss_npre_under50bp_sourcetracker_sink_predictions.txt")

sourcetracker_10M_all <- sourcetracker_10M %>%
  full_join(., sourcetracker_10M_over50bp) %>%
  full_join(., sourcetracker_10M_under50bp) %>%
  rename(Sequencing_ID = 1) %>%
  full_join(., metadata) %>%
  mutate(Read_length_subsampling = str_replace_all(Read_length_subsampling, "none", "None"),
         Read_length_subsampling = str_replace_all(Read_length_subsampling, "over50bp", "Over 50 bp"),
         Read_length_subsampling = str_replace_all(Read_length_subsampling, "under50bp", "Under 50 bp"))

sources_plot_rl_subsampling <- sourcetracker_10M_all %>%
  rename(Environment = environment, Gut = gut, "Modern Calculus" = modern_calc, Plaque = plaque, Skin = skin) %>%
  gather("Sources", "Percent",2:7) %>%
  mutate(Percent = Percent * 100) %>%
  mutate(Sources = fct_relevel(Sources, "Unknown", "Environment", "Skin","Gut","Plaque","Modern Calculus")) %>%
  ggplot(., aes(x = Read_length_subsampling, y = Percent, fill = Sources, pattern = Sources)) +
          geom_bar_pattern(
            aes(pattern_angle = Sources),
            linewidth = .01,
            stat = "identity",
            color = "black",
            #pattern_alpha = 0.45,
            pattern_fill = NA,
            pattern_density = 0.03,
            pattern_spacing = 0.07,
            pattern_size = 0.15,
            pattern_key_scale_factor = 0.2) +
          theme_minimal(base_size = 14) +
          scale_fill_manual(values = source_colors) +
          scale_pattern_manual(values = source_patterns) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1),
                axis.text = element_text(size = 8)) +
          geom_hline(yintercept=75, linetype="dotted") +
          geom_hline(yintercept=90, linetype="dotted", color = "gray50") +
          #scale_y_continuous(breaks=scales::pretty_breaks(n=6)) +
          ylab("% attributed to source") +
          xlab("Individual ID and Strandedness") +
          facet_grid(Strandedness~Individual_ID) +
          scale_pattern_angle_manual(values = source_angles)


sources_plot_rl_subsampling

#ggsave(file.path("./06-publication", paste0(folder), paste0("SI_Fig6B_", file_modifier, "_rl_subsampling_kraken2_sourcetracker_IndividualID_facet_edited.pdf")), plot = sources_plot_rl_subsampling, device = "pdf", scale = 1, width = 10, height = 6, units = c("in"), dpi = 300)

```


#Beta diversity - Read length subsampling
## Total species PCA
```{r}
kraken2_total_species_table_rl <- kraken2_total_species_rl %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "Counts") %>%
  mutate(Counts = Counts + 1) %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  filter(!str_detect(Sequencing_ID, "Isolation_Source")) %>%
  column_to_rownames("Sequencing_ID")
  
result.pca.kraken2_rl <- pca(kraken2_total_species_table_rl, logratio = "CLR") 

exp_var <- paste0(round(result.pca.kraken2_rl$prop_expl_var$X * 100, 2), "%") # explained variance for x- and y-labels

plotIndiv(result.pca.kraken2_rl)

pca_kraken2_rl <- result.pca.kraken2_rl$variates %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  dplyr::rename(Sequencing_ID = 1,
         PC1 = 2,
         PC2 = 3) %>%
  full_join(., metadata_rl, by = "Sequencing_ID")

permanova_dist_matrix <- dist(pca_kraken2_rl[, c("PC1", "PC2")], method = "euclidean")
adonis_result <- adonis2(permanova_dist_matrix ~ Strandedness, data = pca_kraken2_rl) #Strandedness, Individual_ID, subsampling
r_squared_value <- adonis_result$R2[1] 
p_value <- adonis_result$`Pr(>F)`[1]
p_value_formatted <- ifelse(p_value < 0.001, "< 0.001",
                            as.character(signif(p_value, digits = 2)))
r_squared_value_formatted <- ifelse(r_squared_value < 0.001, "< 0.001",
                            as.character(signif(r_squared_value, digits = 2)))


#keep color for strandedness, and change shape for preservation 
color <- c(double_color, single_color)


kraken2_pca_strandedness_rl <- pca_kraken2_rl %>%
  mutate(subsampling = str_replace(subsampling, "10M_over50bp", "Over 50 bp"),
         subsampling = str_replace(subsampling, "10M_under50bp", "Under 50 bp")) %>%
  mutate(subsampling = factor(subsampling, levels = c("Under 50 bp", "Over 50 bp"))) %>%
  ggplot(., aes(x = PC1, y = PC2, color = Strandedness)) +
    geom_point(aes(shape = subsampling), size = 3) +
    theme_minimal() +
    labs(title = "", y = paste("PC2", " - ", exp_var[2]), x = paste("PC1", " - ", exp_var[1])) +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_color_discrete(name = "Strandedness", labels = c("Double", "Single")) +
    scale_shape_discrete(name = "Read length subsampling", labels = c("Under 50 bp", "Over 50 bp")) +
    geom_line(aes(group = Individual_ID), color = "#555555ff", linetype = "dashed") +
    scale_colour_manual(name = "Strandedness", values = color) +
    scale_shape_manual(name = "Read length subsampling", values = c(16, 17)) +
    theme(axis.text.x = element_text(size = 12),  
          axis.text.y = element_text(size = 12),
          strip.text = element_text(size = 12),
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 12),
          axis.title.x = element_text(size = 12),
          axis.title.y = element_text(size = 12)) +
    geom_text_repel(data = subset(pca_kraken2_rl, Strandedness == "Double"), 
                          aes(label = Individual_ID), 
                          size = 3, color = "black", segment.color = NA,
                          nudge_y = -0.8) +
      coord_equal() +
    annotate("text", x = 50, y = -35,
           label = paste("R2 =", r_squared_value_formatted,
                         "\np-value =", p_value_formatted))
 
kraken2_pca_strandedness_rl


```


## Total species PCA biplot
```{r}
#Define colors and shapes for plotting
Strandedness_colors <- c("Double" = double_color, "Single" = single_color)
subsampling_shapes <- c("Under 50 bp" = 16, "Over 50 bp" = 17)

#Biplot Function
plot_pca_bi_rl <- function(df, pc1, pc2, metadata_group1, metadata_group2, columntitle) {
    metadata_group_colors <- get(paste(metadata_group1, "_colors", sep = ""))
    metadata_group_shapes <- get(paste(metadata_group2, "_shapes", sep = ""))
   
    arrow_pc <- enquo(columntitle)
    
    exp_var <- paste0(round(df$prop_expl_var$X * 100, 2), "%") # explained variance for x- and y-labels
    
    # select only the PCs from the PCA and add metadata
    df_X <- df$variates$X %>%
              as.data.frame() %>%
              rownames_to_column("Sequencing_ID") %>%
              inner_join(metadata_rl, by = "Sequencing_ID") %>%
              mutate(subsampling = str_replace_all(subsampling, "10M_over50bp", "Over 50 bp"),
                     subsampling = str_replace_all(subsampling, "10M_under50bp", "Under 50 bp"))
    
    metadata_group1 = df_X[[metadata_group1]]
    metadata_group2 = df_X[[metadata_group2]]
    
    corr_lam <- df$sdev[c("PC1", "PC2")] * sqrt(nrow(df_X))

    df_X <- df_X %>%
      mutate(PC1 = PC1 / corr_lam[1],
             PC2 = PC2 / corr_lam[2])

    # select the correct PC column and explained variance for PC1
    if (pc1 == 'PC1') {
        Pc1 <- df_X$PC1
        exp_var_pc1 <- exp_var[1]
        xaxis <- c("PC1")
    } else if (pc1 == 'PC2') {
        Pc1 <- df_X$PC2
        exp_var_pc1 <- exp_var[2]
        xaxis <- c("PC2")
    } 
    
    # select the correct PC column and explained variance for PC2
    if (pc2 == 'PC1') {
        Pc2 <- df_X$PC1
        exp_var_pc2 <- exp_var[1]
        yaxis <- c("PC1")
 } else if (pc2 == 'PC2') {
       Pc2 <- df_X$PC2
       exp_var_pc2 <- exp_var[2]
       yaxis <- c("PC2")
    }
    
    # Identify the 10 pathways that have highest positive and negative loadings in the selected PC
    pws_10 <- df$loadings$X %>%
      as.data.frame(.) %>%
      rownames_to_column(var = "Taxon") %>%
      top_n(10, !!arrow_pc) %>%
      arrange(desc(!!arrow_pc)) %>%   #sort by size of loading to get strongest first
      mutate(label = row_number())

    neg_10 <- df$loadings$X %>%
      as.data.frame(.) %>%
      rownames_to_column(var = "Taxon") %>%
      top_n(-10, !!arrow_pc) %>%
      arrange(!!arrow_pc) %>%  #sort to get the smallest (most negative/strongest) first
      mutate(label = row_number())

    pca_plot_bi <- df_X %>%
      mutate(metadata_group2 = factor(metadata_group2, levels = c("Under 50 bp", "Over 50 bp"))) %>%
      ggplot(., aes(x = Pc1, y = Pc2, color = metadata_group1)) +
      geom_point(size = 2.5, aes(shape = metadata_group2) )+
      geom_segment(data = pws_10,
                   aes(xend = get(paste(pc1)), yend = get(paste(pc2))),
                   x = 0, y = 0, colour = "black",
                   size = 0.3,
                   arrow = arrow(length = unit(0.03, "npc"))) +
      #geom_label_repel(data = pws_10,
      #             aes(x = get(paste(pc1)), y = get(paste(pc2)), label = label), #using label instead of taxon
      #             size = 2, 
      #             colour = "black",
      #             label.padding = 0.1, 
      #             box.padding = 0.1, #
      #             point.padding = 0.1, 
      #             force = 10,
      #             max.overlaps = Inf) + 
      geom_segment(data = neg_10,
                   aes(xend = get(paste(pc1)), yend = get(paste(pc2))),
                   x = 0, y = 0, colour = "grey50",
                   size = 0.3,
                   arrow = arrow(length = unit(0.03, "npc"))) +
      #geom_label_repel(data = neg_10,
      #             aes(x = get(paste(pc1)), y = get(paste(pc2)), label = label), #using label instead of taxon
      #             size = 2, 
      #             colour = "red",
      #             label.padding = 0.1, 
      #             box.padding = 0.1, 
      #             point.padding = 0.1, 
      #             force = 10, 
      #             max.overlaps = Inf) + 
      labs(x = paste(xaxis, " - ", exp_var_pc1),
           y = paste(yaxis, " - ", exp_var_pc2),
           color = "Strandedness",
           shape = "Read length subsampling") +
      scale_color_manual(values = metadata_group_colors) +
      scale_shape_manual(values = metadata_group_shapes) +
      theme_minimal() + theme(text = element_text(size = 16)) +
      theme(text = element_text(size=16)) +
      theme(legend.position = "bottom") +
      geom_line(aes(group = Individual_ID), color = "#555555ff", linetype = "dashed") +
      geom_text_repel(aes(label = Individual_ID), color = "black", segment.color = NA) +
      coord_equal() +
      coord_equal()
      

    return(pca_plot_bi) #pca_plot_bi, pws_10, or neg_10
}

pca_kraken2_biplot_PC1_rl <- plot_pca_bi_rl(result.pca.kraken2_rl, "PC1", "PC2", "Strandedness", "subsampling", PC1)
pca_kraken2_biplot_PC2_rl <- plot_pca_bi_rl(result.pca.kraken2_rl, "PC1", "PC2", "Strandedness", "subsampling", PC2)

pca_kraken2_biplot_PC1_rl
pca_kraken2_biplot_PC2_rl

pca_biplots_combined_rl <- ggarrange(
  pca_kraken2_biplot_PC1_rl, pca_kraken2_biplot_PC2_rl, 
  labels = c("A", "B"), 
  ncol = 1, nrow = 2, 
  common.legend = TRUE, 
  legend = "bottom")

pca_biplots_combined_rl

#ggsave(file.path("./06-publication", paste0(folder), paste0("SI_Fig27AB_", file_modifier, "_rl_subsampling_kraken2_pca_biplot_combined_PCs_total_species.pdf")) , plot = pca_biplots_combined_rl, device = "pdf", scale = 1, width = 8, height = 12, units = c("in"), dpi = 300)

```


## Oral species PCA

```{r}
kraken2_oral_species_table_rl <- fread("./05-results/normalize_eager/dsvss_npre_read_length_kraken2_gtdb207_oral_species.tsv", header = TRUE) 

kraken2_oral_species_rl <- kraken2_oral_species_table_rl %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "Counts") %>%
  mutate(Counts = Counts + 1) %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  filter(!str_detect(Sequencing_ID, "Isolation_Source")) %>%
  column_to_rownames("Sequencing_ID")
  
result.pca.kraken2_rl <- pca(kraken2_oral_species_rl, logratio = "CLR") 

exp_var <- paste0(round(result.pca.kraken2_rl$prop_expl_var$X * 100, 2), "%") # explained variance for x- and y-labels

pca_kraken2_rl <- result.pca.kraken2_rl$variates %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  dplyr::rename(Sequencing_ID = 1,
         PC1 = 2,
         PC2 = 3) %>%
  full_join(., metadata_rl, by = "Sequencing_ID")

#Create a metadata file with sample IDs, grouping info, and PCA scores
permanova_data_rl <- pca_kraken2_rl %>%
  dplyr::select(Sequencing_ID, Individual_ID, PC1, PC2, Strandedness, subsampling) %>%
  column_to_rownames("Sequencing_ID")


#Create a distance matrix of just PCA scores
permanova_dist_matrix_rl <- dist(permanova_data_rl[, c("PC1", "PC2")], method = "euclidean")

#Run adonis2
adonis_result <- adonis2(permanova_dist_matrix_rl ~ subsampling, data = permanova_data_rl) #Strandedness, Individual_ID, subsampling

r_squared_value <- adonis_result$R2[1] 
  
p_value <- adonis_result$`Pr(>F)`[1]

p_value_formatted <- ifelse(p_value < 0.001, "< 0.001",
                            as.character(signif(p_value, digits = 2)))

r_squared_value_formatted <- ifelse(r_squared_value < 0.001, "< 0.001",
                            as.character(signif(r_squared_value, digits = 2)))

#keep color for strandedness, and change shape for preservation 
color <- c(double_color, single_color)

kraken2_pca_oral_rl <- pca_kraken2_rl %>%
   mutate(subsampling = str_replace(subsampling, "10M_over50bp", "Over 50 bp"),
         subsampling = str_replace(subsampling, "10M_under50bp", "Under 50 bp")) %>%
  mutate(subsampling = factor(subsampling, levels = c("Under 50 bp", "Over 50 bp"))) %>%
  ggplot(., aes(x = PC1, y = PC2, color = Strandedness)) +
    geom_point(aes(shape = subsampling), size = 3) +
    theme_minimal() +
    labs(title = "", y = paste("PC2", " - ", exp_var[2]), x = paste("PC1", " - ", exp_var[1])) +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_color_discrete(name = "Strandedness", labels = c("Double", "Single")) +
    scale_shape_discrete(name = "Read length subsampling", labels = c("Under 50 bp", "Over 50 bp")) +
    geom_line(aes(group = Individual_ID), color = "#555555ff", linetype = "dashed") +
    scale_colour_manual(name = "Strandedness", values = color) +
    scale_shape_manual(name = "Read length subsampling", values = c(16, 17)) +
    theme(axis.text.x = element_text(size = 12),  
          axis.text.y = element_text(size = 12),
          strip.text = element_text(size = 12),
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 12),
          axis.title.x = element_text(size = 12),
          axis.title.y = element_text(size = 12)) +
    geom_text_repel(data = subset(pca_kraken2_rl, Strandedness == "Double"), 
                        aes(label = Individual_ID), 
                        size = 3, color = "black", segment.color = NA,
                        nudge_y = -0.8) +
    coord_equal() +
  annotate("text", x = -25, y = 15,
           label = paste("R2 =", r_squared_value_formatted,
                         "\np-value =", p_value_formatted))
 
kraken2_pca_oral_rl


```


## Oral species PCA biplot
```{r}
#Define colors and shapes for plotting
Strandedness_colors <- c("Double" = double_color, "Single" = single_color)
subsampling_shapes <- c("Under 50 bp" = 16, "Over 50 bp" = 17)

#Biplot Function
plot_pca_bi_rl <- function(df, pc1, pc2, metadata_group1, metadata_group2, columntitle) {
    metadata_group_colors <- get(paste(metadata_group1, "_colors", sep = ""))
    metadata_group_shapes <- get(paste(metadata_group2, "_shapes", sep = ""))
   
    arrow_pc <- enquo(columntitle)
    
    exp_var <- paste0(round(df$prop_expl_var$X * 100, 2), "%") # explained variance for x- and y-labels
    
    # select only the PCs from the PCA and add metadata
    df_X <- df$variates$X %>%
              as.data.frame() %>%
              rownames_to_column("Sequencing_ID") %>%
              inner_join(metadata_rl, by = "Sequencing_ID") %>%
              mutate(subsampling = str_replace_all(subsampling, "10M_over50bp", "Over 50 bp"),
                     subsampling = str_replace_all(subsampling, "10M_under50bp", "Under 50 bp"))
    
    metadata_group1 = df_X[[metadata_group1]]
    metadata_group2 = df_X[[metadata_group2]]
    
    corr_lam <- df$sdev[c("PC1", "PC2")] * sqrt(nrow(df_X))

    df_X <- df_X %>%
      mutate(PC1 = PC1 / corr_lam[1],
             PC2 = PC2 / corr_lam[2])

    # select the correct PC column and explained variance for PC1
    if (pc1 == 'PC1') {
        Pc1 <- df_X$PC1
        exp_var_pc1 <- exp_var[1]
        xaxis <- c("PC1")
    } else if (pc1 == 'PC2') {
        Pc1 <- df_X$PC2
        exp_var_pc1 <- exp_var[2]
        xaxis <- c("PC2")
    } 
    
    # select the correct PC column and explained variance for PC2
    if (pc2 == 'PC1') {
        Pc2 <- df_X$PC1
        exp_var_pc2 <- exp_var[1]
        yaxis <- c("PC1")
 } else if (pc2 == 'PC2') {
       Pc2 <- df_X$PC2
       exp_var_pc2 <- exp_var[2]
       yaxis <- c("PC2")
    }
    
    # Identify the 10 pathways that have highest positive and negative loadings in the selected PC
    pws_10 <- df$loadings$X %>%
      as.data.frame(.) %>%
      rownames_to_column(var = "Taxon") %>%
      top_n(10, !!arrow_pc) %>%
      arrange(desc(!!arrow_pc)) %>%   #sort by size of loading to get strongest first
      mutate(label = row_number())

    neg_10 <- df$loadings$X %>%
      as.data.frame(.) %>%
      rownames_to_column(var = "Taxon") %>%
      top_n(-10, !!arrow_pc) %>%
      arrange(!!arrow_pc) %>%  #sort to get the smallest (most negative/strongest) first
      mutate(label = row_number())

    pca_plot_bi <- df_X %>%
      mutate(metadata_group2 = factor(metadata_group2, levels = c("Under 50 bp", "Over 50 bp"))) %>%
      ggplot(., aes(x = Pc1, y = Pc2, color = metadata_group1)) +
      geom_point(size = 2.5, aes(shape = metadata_group2) )+
      geom_segment(data = pws_10,
                   aes(xend = get(paste(pc1)), yend = get(paste(pc2))),
                   x = 0, y = 0, colour = "black",
                   size = 0.3,
                   arrow = arrow(length = unit(0.03, "npc"))) +
      #geom_label_repel(data = pws_10,
      #             aes(x = get(paste(pc1)), y = get(paste(pc2)), label = label), #using label instead of taxon
      #             size = 2, 
      #             colour = "black",
      #             label.padding = 0.1, 
      #             box.padding = 0.1, #
      #             point.padding = 0.1, 
      #             force = 10,
      #             max.overlaps = Inf) + 
      geom_segment(data = neg_10,
                   aes(xend = get(paste(pc1)), yend = get(paste(pc2))),
                   x = 0, y = 0, colour = "grey50",
                   size = 0.3,
                   arrow = arrow(length = unit(0.03, "npc"))) +
      #geom_label_repel(data = neg_10,
      #             aes(x = get(paste(pc1)), y = get(paste(pc2)), label = label), #using label instead of taxon
      #             size = 2, 
      #             colour = "red",
      #             label.padding = 0.1, 
      #             box.padding = 0.1, 
      #             point.padding = 0.1, 
      #             force = 10, 
      #             max.overlaps = Inf) + 
      labs(x = paste(xaxis, " - ", exp_var_pc1),
           y = paste(yaxis, " - ", exp_var_pc2),
           color = "Strandedness",
           shape = "Read length subsampling") +
      scale_color_manual(values = metadata_group_colors) +
      scale_shape_manual(values = metadata_group_shapes) +
      theme_minimal() + theme(text = element_text(size = 16)) +
      theme(text = element_text(size=16)) +
      theme(legend.position = "bottom") +
      geom_line(aes(group = Individual_ID), color = "#555555ff", linetype = "dashed") +
      #geom_text_repel(aes(label = Individual_ID), color = "black", segment.color = NA, max.overlaps = 20) +
      coord_equal()
      

    return(pca_plot_bi) #pca_plot_bi, pws_10, or neg_10
}

pca_kraken2_biplot_PC1_rl <- plot_pca_bi_rl(result.pca.kraken2_rl, "PC1", "PC2", "Strandedness", "subsampling", PC1)
pca_kraken2_biplot_PC2_rl <- plot_pca_bi_rl(result.pca.kraken2_rl, "PC1", "PC2", "Strandedness", "subsampling", PC2)

pca_kraken2_biplot_PC1_rl
pca_kraken2_biplot_PC2_rl

pca_biplots_combined_rl_oral <- ggarrange(
  pca_kraken2_biplot_PC1_rl, pca_kraken2_biplot_PC2_rl, 
  labels = c("A", "B"), 
  ncol = 1, nrow = 2, 
  common.legend = TRUE, 
  legend = "bottom")

pca_biplots_combined_rl_oral

#ggsave(file.path("./06-publication", paste0(folder), paste0("SI_Fig27CD_", file_modifier, "_rl_subsampling_kraken2_pca_biplot_combined_PCs_oral_species.pdf")) , plot = pca_biplots_combined_rl_oral, device = "pdf", scale = 1, width = 8, height = 12, units = c("in"), dpi = 300)

```


## Oral species PCA biplot - well-preserved only
```{r}
#Filter out BAN001 and FUM003
kraken2_oral_species_rl <- kraken2_oral_species_table_rl %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "Counts") %>%
  filter(!str_detect(Sequencing_ID, "BAN001")) %>%
  filter(!str_detect(Sequencing_ID, "FUM003")) %>%
  filter(!str_detect(Sequencing_ID, "GUE001")) %>%
  filter(!str_detect(Sequencing_ID, "SPM002")) %>%
  filter(!str_detect(Sequencing_ID, "ZAF001")) %>%
  mutate(Counts = Counts + 1) %>%
  pivot_wider(names_from = "Species", values_from = "Counts") %>%
  filter(!str_detect(Sequencing_ID, "Isolation_Source")) %>%
  column_to_rownames("Sequencing_ID")
  
result.pca.kraken2_rl <- pca(kraken2_oral_species_rl, logratio = "CLR") 

exp_var <- paste0(round(result.pca.kraken2_rl$prop_expl_var$X * 100, 2), "%") # explained variance for x- and y-labels

pca_kraken2_rl <- result.pca.kraken2_rl$variates %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  dplyr::rename(Sequencing_ID = 1,
         PC1 = 2,
         PC2 = 3) %>%
  inner_join(., metadata_rl, by = "Sequencing_ID")

#Create a metadata file with sample IDs, grouping info, and PCA scores
permanova_data_rl <- pca_kraken2_rl %>%
  dplyr::select(Sequencing_ID, Individual_ID, PC1, PC2, Strandedness, subsampling) %>%
  column_to_rownames("Sequencing_ID")

#Create a distance matrix of just PCA scores
permanova_dist_matrix_rl <- dist(permanova_data_rl[, c("PC1", "PC2")], method = "euclidean")

#Run adonis2
adonis_result <- adonis2(permanova_dist_matrix_rl ~ Strandedness, data = permanova_data_rl) #Strandedness, Individual_ID, subsampling

#Define colors and shapes for plotting
Strandedness_colors <- c("Double" = double_color, "Single" = single_color)
subsampling_shapes <- c("Under 50 bp" = 16, "Over 50 bp" = 17)

#Biplot Function
plot_pca_bi_rl <- function(df, pc1, pc2, metadata_group1, metadata_group2, columntitle) {
    metadata_group_colors <- get(paste(metadata_group1, "_colors", sep = ""))
    metadata_group_shapes <- get(paste(metadata_group2, "_shapes", sep = ""))
   
    arrow_pc <- enquo(columntitle)
    
    exp_var <- paste0(round(df$prop_expl_var$X * 100, 2), "%") # explained variance for x- and y-labels
    
    # select only the PCs from the PCA and add metadata
    df_X <- df$variates$X %>%
              as.data.frame() %>%
              rownames_to_column("Sequencing_ID") %>%
              inner_join(metadata_rl, by = "Sequencing_ID") %>%
              mutate(subsampling = str_replace_all(subsampling, "10M_over50bp", "Over 50 bp"),
                     subsampling = str_replace_all(subsampling, "10M_under50bp", "Under 50 bp"))
    
    metadata_group1 = df_X[[metadata_group1]]
    metadata_group2 = df_X[[metadata_group2]]
    
    corr_lam <- df$sdev[c("PC1", "PC2")] * sqrt(nrow(df_X))

    df_X <- df_X %>%
      mutate(PC1 = PC1 / corr_lam[1],
             PC2 = PC2 / corr_lam[2])

    # select the correct PC column and explained variance for PC1
    if (pc1 == 'PC1') {
        Pc1 <- df_X$PC1
        exp_var_pc1 <- exp_var[1]
        xaxis <- c("PC1")
    } else if (pc1 == 'PC2') {
        Pc1 <- df_X$PC2
        exp_var_pc1 <- exp_var[2]
        xaxis <- c("PC2")
    } 
    
    # select the correct PC column and explained variance for PC2
    if (pc2 == 'PC1') {
        Pc2 <- df_X$PC1
        exp_var_pc2 <- exp_var[1]
        yaxis <- c("PC1")
 } else if (pc2 == 'PC2') {
       Pc2 <- df_X$PC2
       exp_var_pc2 <- exp_var[2]
       yaxis <- c("PC2")
    }
    
    # Identify the 10 pathways that have highest positive and negative loadings in the selected PC
    pws_10 <- df$loadings$X %>%
      as.data.frame(.) %>%
      rownames_to_column(var = "Taxon") %>%
      top_n(10, !!arrow_pc) %>%
      arrange(desc(!!arrow_pc)) %>%   #sort by size of loading to get strongest first
      mutate(label = row_number())

    neg_10 <- df$loadings$X %>%
      as.data.frame(.) %>%
      rownames_to_column(var = "Taxon") %>%
      top_n(-10, !!arrow_pc) %>%
      arrange(!!arrow_pc) %>%  #sort to get the smallest (most negative/strongest) first
      mutate(label = row_number())

    pca_plot_bi <- df_X %>%
      mutate(metadata_group2 = factor(metadata_group2, levels = c("Under 50 bp", "Over 50 bp"))) %>%
      ggplot(., aes(x = Pc1, y = Pc2, color = metadata_group1)) +
      geom_point(size = 2.5, aes(shape = metadata_group2) )+
      geom_segment(data = pws_10,
                   aes(xend = get(paste(pc1)), yend = get(paste(pc2))),
                   x = 0, y = 0, colour = "black",
                   size = 0.3,
                   arrow = arrow(length = unit(0.03, "npc"))) +
      #geom_label_repel(data = pws_10,
      #             aes(x = get(paste(pc1)), y = get(paste(pc2)), label = label), #using label instead of taxon
      #             size = 2, 
      #             colour = "black",
      #             label.padding = 0.1, 
      #             box.padding = 0.1, #
      #             point.padding = 0.1, 
      #             force = 10,
      #             max.overlaps = Inf) + 
      geom_segment(data = neg_10,
                   aes(xend = get(paste(pc1)), yend = get(paste(pc2))),
                   x = 0, y = 0, colour = "#555555ff",
                   size = 0.3,
                   arrow = arrow(length = unit(0.03, "npc"))) +
      #geom_label_repel(data = neg_10,
      #             aes(x = get(paste(pc1)), y = get(paste(pc2)), label = label), #using label instead of taxon
      #             size = 2, 
      #             colour = "red",
      #             label.padding = 0.1, 
      #             box.padding = 0.1, 
      #             point.padding = 0.1, 
      #             force = 10, 
      #             max.overlaps = Inf) + 
      labs(x = paste(xaxis, " - ", exp_var_pc1),
           y = paste(yaxis, " - ", exp_var_pc2),
           color = "Strandedness",
           shape = "Read length subsampling") +
      scale_color_manual(values = metadata_group_colors) +
      scale_shape_manual(values = metadata_group_shapes) +
      theme_minimal() + theme(text = element_text(size = 16)) +
      theme(text = element_text(size=16)) +
      theme(legend.position = "bottom") +
      geom_line(aes(group = Individual_ID), color = "#CCCCCC", linetype = "dashed") +
      geom_text_repel(aes(label = Individual_ID), color = "black", segment.color = NA, max.overlaps = 20) +
      coord_equal()
      

    return(pca_plot_bi) #pca_plot_bi, pws_10, or neg_10
}

pca_kraken2_biplot_PC1_rl <- plot_pca_bi_rl(result.pca.kraken2_rl, "PC1", "PC2", "Strandedness", "subsampling", PC1)
pca_kraken2_biplot_PC2_rl <- plot_pca_bi_rl(result.pca.kraken2_rl, "PC1", "PC2", "Strandedness", "subsampling", PC2)

pca_kraken2_biplot_PC1_rl
pca_kraken2_biplot_PC2_rl

pca_biplots_combined_rl_oral <- ggarrange(
  pca_kraken2_biplot_PC1_rl, pca_kraken2_biplot_PC2_rl, 
  labels = c("A", "B"), 
  ncol = 1, nrow = 2, 
  common.legend = TRUE, 
  legend = "bottom")

pca_biplots_combined_rl_oral

#ggsave(file.path("./06-publication", paste0(folder), paste0("SI_Fig_27EF_", file_modifier, "_rl_subsampling_kraken2_pca_biplot_combined_PCs_oral_species_well_preserved.pdf")) , plot = pca_biplots_combined_rl_oral, device = "pdf", scale = 1, width = 8, height = 12, units = c("in"), dpi = 300)

```


## Sources PCoA
```{r}
#Convert to a phyloseq object to rarefy the table before creating the PCoA

combined_kraken2_species_rl <- full_join(kraken2_total_species_rl, kraken2_sources_filtered, by = "Species") %>%
  mutate(across(everything(), ~ replace_na(.x, 0)))

#Make the species column rownames
combined_kraken2_species_formatted <- combined_kraken2_species_rl %>%
  select(-ERR1883424, -ERR1883437) %>% # filter out samples with a species sum of 0 after filtering
  column_to_rownames("Species")

#Convert the species table to a matrix
species_matrix <- as.matrix(combined_kraken2_species_formatted)

#Convert the matrix to a phyloseq object
ps <- phyloseq(otu_table(species_matrix, taxa_are_rows=TRUE))

#Format metadata correctly
mapfile <- fread("./05-results/normalize_eager/dsvss_npre_source_tracker_mappingfile_complete.tsv") %>%
  dplyr::rename(Sequencing_ID = 1) %>%
  dplyr::select(Sequencing_ID, Study, Env, SourceSink) %>%
  filter(!str_ends(Sequencing_ID, "10M"))

metadata_full <- metadata_rl %>%
  full_join(., mapfile) %>%
  mutate(Control_or_Sample = case_when(grepl("source", SourceSink) ~ "Control", TRUE ~ "Sample")) 

metadata_formatted <- metadata_full %>%
  dplyr::rename(SampleID = Sequencing_ID) %>%
  column_to_rownames("SampleID")

#Convert metadata to a phyloseq object
sample_data_obj <- sample_data(metadata_formatted)

#Add the metadata to the phyloseq object
ps <- merge_phyloseq(ps, sample_data_obj)

#Rarefy the species table
#Extract the sample data as a data frame
species_table <- otu_table(ps)
sample_sums <- sample_sums(species_table)

#Find the minimum sample sum
min_sample_sum <- min(sample_sums)

#Rarefy
physeq_rarefied <- rarefy_even_depth(species_table, sample.size = min_sample_sum, rngseed = 1)

#Convert to distance matrix (Jaccard dissimilarity)
physeq_rarefied_formatted <- as.data.frame(physeq_rarefied)

physeq_rarefied_formatted <- physeq_rarefied_formatted %>%
  rownames_to_column("Species") %>%
  pivot_longer(2:ncol(.), names_to = "SampleID", values_to = "counts") %>%
  pivot_wider(names_from = "Species", values_from = "counts") %>%
  arrange(SampleID) %>%
  column_to_rownames("SampleID") 

dist_matrix <- vegdist(physeq_rarefied_formatted, method = "jaccard")

#PCoA Calculation
pcoa_result <- ape::pcoa(dist_matrix)
pcoa_coordinates <- pcoa_result$vectors[, 1:2]
colnames(pcoa_coordinates) <- c("PCoA1", "PCoA2")

#Combine PCoA coordinates with metadata
pcoa_df <- as.data.frame(pcoa_coordinates)
pcoa_df$Sequencing_ID <- rownames(pcoa_df)
pcoa_df <- merge(pcoa_df, metadata_full, by = "Sequencing_ID")

#Variation explained by each axis
eig_values <- pcoa_result$values$Relative_eig
var_explained <- round(eig_values[1:2] * 100, digits = 2)
x_lab <- paste0("PCoA1 - ", round(var_explained[1], 2), "%")
y_lab <- paste0("PCoA2 - ", round(var_explained[1], 2), "%")

#Permanova
adonis_result <- adonis2(dist_matrix ~ Env, data = pcoa_df)

r_squared_value <- adonis_result$R2[1] 
p_value <- adonis_result$`Pr(>F)`[1]
p_value_formatted <- ifelse(p_value < 0.001, "< 0.001",
                            as.character(signif(p_value, digits = 2)))
r_squared_value_formatted <- ifelse(r_squared_value < 0.001, "< 0.001",
                            as.character(signif(r_squared_value, digits = 2)))

#Define colors, shapes, and plot

colors <- c("Ancient Calculus - ssLib" = single_color, "Ancient Calculus - dsLib" = double_color,
  "Modern Calculus" = "#bfaed4",  
  "Plaque"          = "#3f007d",  
  "Gut"             = "#c05b16",  
  "Skin"            = "#fdc187",  
  "Environment"     = "#7fc97f",  
  "Unknown"         = "#d9d9d9"   
)

shapes <- c("Source" = 16, "Over 50 bp" = 15, "Under 50 bp" = 17)

sources_pcoa_rl <- pcoa_df %>%
  mutate(Env = str_replace_all(Env, "calculus", "Ancient Calculus"),
         Env = str_replace_all(Env, "environment", "Environment"),
         Env = str_replace_all(Env, "modern_calc", "Modern Calculus"),
         Env = str_replace_all(Env, "skin", "Skin"),
         Env = str_replace_all(Env, "gut", "Gut"),
         Env = str_replace_all(Env, "plaque", "Plaque")) %>%
  mutate(Env = case_when(Strandedness == "Single" ~ "Ancient Calculus - ssLib", TRUE ~ Env)) %>%
  mutate(Env = case_when(Strandedness == "Double" ~ "Ancient Calculus - dsLib", TRUE ~ Env)) %>%
  mutate(SourceSink = str_replace_all(SourceSink, "source", "Source")) %>%
  mutate(subsampling = replace_na(subsampling, "Source")) %>%
  mutate(SourceSink = ifelse(str_detect(subsampling, "10M_under50bp"), "Under 50 bp", SourceSink),
         SourceSink = ifelse(str_detect(subsampling, "10M_over50bp"), "Over 50 bp", SourceSink)) %>%
  mutate(Individual_ID = case_when(is.na(Individual_ID) ~ Library_ID, TRUE ~ Individual_ID)) %>%
  mutate(Strandedness = case_when(is.na(Strandedness) ~ "Source", TRUE ~ Strandedness)) %>%
  mutate(Env = factor(Env, levels = c("Ancient Calculus - dsLib", "Ancient Calculus - ssLib", "Modern Calculus", "Plaque", "Gut", "Skin", "Environment"))) %>%
  mutate(SourceSink = factor(SourceSink, levels = c("Under 50 bp", "Over 50 bp", "Source"))) %>%
  ggplot(., aes(x = PCoA1, y = PCoA2, color = Env, shape = SourceSink)) +
    geom_point(aes(alpha = Control_or_Sample), size = 2.5) +
    theme_minimal() +
    labs(title = "") +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_color_manual(name = "Environment", values = colors) +
    scale_shape_manual(name = "Sample", values = shapes) +
    #geom_line(aes(group = Individual_ID), color = "#CCCCCC", linetype = "dashed") +
    theme(axis.text.x = element_text(size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14)) +
  scale_alpha_manual(values = c("Control" = 0.7, "Sample" = 1)) +
  xlab(x_lab) +
  ylab(y_lab) +
  guides(alpha = "none") +
  geom_line(data = pcoa_df %>% filter(!is.na(Individual_ID)), 
              aes(group = Individual_ID), 
              color = "#555555ff", 
              linetype = "dashed") +
  #geom_text(data = pcoa_df %>% filter(Strandedness %in% c("Single")), 
  #            aes(label = Individual_ID), 
  #            hjust = -0.3, 
  #            vjust = -0.3, 
  #            size = 3, color = "black") +
  coord_equal() +
  annotate("text", x = 0.1, y =  -0.25,
           label = paste("R2 =", r_squared_value_formatted,
                         "\np-value =", p_value_formatted))

sources_pcoa_rl

```


#Classified Reads
##Median read length of classified reads
```{r}
#color by the percentage of reads classified by kraken2 -> out of initial 10M going into eager, how many are classified?

#read in file with each the number of reads classified by kraken2 by read length (per library)
rl_classified <- fread("./05-results/normalize_eager/dsvss_npre_classified_read_length.tsv") %>%
  mutate(Study = str_replace(Study, ".classified_gc_rl.tsv.gz", "")) %>%
  separate(Study, into = "Sequencing_ID", sep = ".tract", extra = "drop") %>%
  full_join(., metadata_no_rl) %>%
  dplyr::rename(counts_classified = counts_length)

rl_levels <- c(30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 73, 74, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99, 100, 101, 102, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 113, 114, 115, 116, 117, 118, 119, 120, 121, 122, 123, 124, 125, 126, 127, 128, 129, 130, 131, 132, 133, 134, 135, 136, 137, 138, 139, 140, 141, 142, 143, 144, 145, 146, 147, 148, 149, 150)

rl_summary_classified <- rl_classified %>%
    mutate(Length = factor(Length, level=rl_levels)) 

#function to calculate the median read length of oral reads per libraryID
median_rl_summary_table <- function(metadata, sample_list, oral_reads_length_table) {
  
  #Create a blank dataframe
  read_length_summary_table <- data.frame(Individual_ID = sample_list,
                                          Double = numeric(length(sample_list)),
                                          Single = numeric(length(sample_list)),
                                          stringsAsFactors = FALSE)
  
  #Function to calculate median read length based on dataset, sample, and strandedness
  calculate_median_rl <- function(data, sample, strandedness) {
    rl_summary <- data %>%
      filter(!str_detect(Library_ID, "bp")) %>%
      dplyr::select(Library_ID, Individual_ID, Strandedness, Length, counts_classified) %>%
      mutate(Length = as.integer(as.character(Length))) %>%
      filter(Individual_ID == sample, Strandedness == strandedness) %>%
      uncount(counts_classified) %>%
      mutate(Length = as.integer(Length)) %>%
      summarize(median_rl = median(Length))
    
    return(rl_summary$median_rl)
  }
  
  
  for (sample in sample_list) {
    #adapter removal
    rl_value_of_median_ds <- calculate_median_rl(oral_reads_length_table, sample, "Double")
    rl_value_of_median_ss <- calculate_median_rl(oral_reads_length_table, sample, "Single")
    read_length_summary_table[read_length_summary_table$Individual_ID == sample, "Double"] <- rl_value_of_median_ds
    read_length_summary_table[read_length_summary_table$Individual_ID == sample, "Single"] <- rl_value_of_median_ss
    
   }

  return(read_length_summary_table)
  
}

samples <- metadata_no_rl %>%
  dplyr::select(Individual_ID) %>%
  unique() %>%
  pull()

median_rl <- median_rl_summary_table(metadata_no_rl, samples, rl_summary_classified) %>%
  pivot_longer(-Individual_ID, values_to = "median_rl", names_to = "Strandedness")

#calculate the percent of reads classified (out of 10M reads going into nf-core/eager)
kraken10M_classified_species_table <- kraken2_total_species_no_rl %>%
  column_to_rownames("Species") %>%
  dplyr::select(all_of(no_rl_samples)) %>%
  rownames_to_column("Species") %>%
  adorn_totals() %>%
  filter(Species == "Total") %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "classified_reads") %>%
  dplyr::select(-Species)

metadata_no_rl <- metadata_no_rl %>%
  full_join(., kraken10M_classified_species_table) %>%
  full_join(., median_rl) %>%
  mutate(percent_classified = (classified_reads/10000000)*100) %>%
  mutate(Library = Strandedness) %>%
  mutate(Library = str_replace_all(Library, "Double", "dsLib"),
         Library = str_replace_all(Library, "Single", "ssLib"))

```


```{r}
wilcox_results <- metadata_no_rl %>%
  wilcox_test(median_rl ~ Library, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- metadata_no_rl %>%
  wilcox_effsize(formula = median_rl ~ Library, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat))

annotations <- wilcox_results %>%
  summarize(annotation = first(annotation),
            y_position = 77, .groups = 'drop')

read_length_median_percent_plot <- metadata_no_rl %>%
  ggplot(., aes(x = Library, y = median_rl)) + 
  geom_boxplot() +
  geom_jitter(width = 0.075, size = 3, aes(color = percent_classified)) +
  scale_color_viridis(option = "inferno", begin = 0.1) +
  theme_classic() +
  scale_shape_manual(values = c(16, 18)) +
  labs(x = "Library Type",
       y = "Median Read Length of Classified Reads",
       color = "Percent of Reads Classified by kraken2") +
  expand_limits(y = 80) +
  guides(alpha = "none") +
  geom_text(data = annotations, aes(x = 1.5, y = y_position, label = annotation), vjust = -0.5, size = 5) +
  theme(axis.text.x = element_text( size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
  geom_line(aes(group = Individual_ID), linetype = "dashed", color = "#555555", linewidth = 0.2) +
  geom_text_repel(aes(label = Individual_ID), size = 2)

read_length_median_percent_plot
```


## Median %GC of classified reads
```{r}
gc_levels <- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 73, 74, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99, 100, 101, 102, 103, 104)

#read in file with each the number of reads classified by kraken2 by %GC (per library)
gc_classified <- fread("./05-results/normalize_eager/dsvss_npre_classified_gc.tsv") %>%
  mutate(Study = str_replace(Study, ".classified_gc_rl.tsv.gz", "")) %>%
  separate(Study, into = "Sequencing_ID", sep = ".tract", extra = "drop") %>%
  full_join(., metadata_no_rl) %>%
  dplyr::rename(counts_classified = counts_gc) %>%
  separate(gc_bin, into = c("drop", "gc"), sep = ",", remove = FALSE) %>%
  mutate(gc = str_remove(gc, "\\]")) %>%
  dplyr::select(-gc_bin, -drop) %>%
  mutate(gc = as.factor(gc))

gc_summary_classified <- gc_classified %>%
    mutate(gc = factor(gc, level=gc_levels)) 

#function to calculate the median read length of oral reads per libraryID
median_gc_summary_table <- function(metadata, sample_list, oral_reads_gc_table) {
  
  #Create a blank dataframe
  gc_summary_table <- data.frame(Individual_ID = sample_list,
                                          Double = numeric(length(sample_list)),
                                          Single = numeric(length(sample_list)),
                                          stringsAsFactors = FALSE)
  
  #Function to calculate median read length based on dataset, sample, and strandedness
  calculate_median_gc <- function(data, sample, strandedness) {
    gc_summary <- data %>%
      filter(!str_detect(Library_ID, "bp")) %>%
      dplyr::select(Library_ID, Individual_ID, Strandedness, gc, counts_classified) %>%
      filter(Individual_ID == sample, Strandedness == strandedness) %>%
      mutate(gc = as.numeric(as.character(gc))) %>%
      uncount(counts_classified) %>%
      summarize(median = median(gc))
    
    return(gc_summary$median)
  }
  
  
  for (sample in sample_list) {
    #adapter removal
    gc_value_of_median_ds <- calculate_median_gc(oral_reads_gc_table, sample, "Double")
    gc_value_of_median_ss <- calculate_median_gc(oral_reads_gc_table, sample, "Single")
    gc_summary_table[gc_summary_table$Individual_ID == sample, "Double"] <- gc_value_of_median_ds
    gc_summary_table[gc_summary_table$Individual_ID == sample, "Single"] <- gc_value_of_median_ss
    
   }

  return(gc_summary_table)
  
}

samples <- metadata_no_rl %>%
  dplyr::select(Individual_ID) %>%
  unique() %>%
  pull()

median_gc <- median_gc_summary_table(metadata_no_rl, samples, gc_summary_classified) %>%
  pivot_longer(-Individual_ID, values_to = "median_gc", names_to = "Strandedness")

metadata_no_rl <- metadata_no_rl %>%
  full_join(., median_gc)

```

```{r}
wilcox_results <- metadata_no_rl %>%
  wilcox_test(median_gc ~ Library, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- metadata_no_rl %>%
  wilcox_effsize(formula = median_gc ~ Library, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat))

annotations <- wilcox_results %>%
  summarize(annotation = first(annotation),
            y_position = 65, .groups = 'drop')

gc_median_percent_plot <- metadata_no_rl %>%
  ggplot(., aes(x = Library, y = median_gc)) + 
  geom_boxplot() +
  geom_jitter(width = 0.075, size = 3, aes(color = percent_classified)) +
  scale_color_viridis(option = "inferno", begin = 0.1) +
  theme_classic() +
  scale_shape_manual(values = c(16, 18)) +
  labs(x = "Library Type",
       y = "Median %GC of Classified Reads",
       color = "Percent of Reads Classified by kraken2") +
  expand_limits(y = 70) +
  guides(alpha = "none") +
  geom_text(data = annotations, aes(x = 1.5, y = y_position, label = annotation), vjust = -0.5, size = 5) +
  theme(axis.text.x = element_text( size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
  geom_line(aes(group = Individual_ID), linetype = "dashed", color = "#555555", linewidth = 0.2) +
  geom_text_repel(aes(label = Individual_ID), size = 2)

gc_median_percent_plot
```


```{r}
summary_plots <- ggarrange(
  read_length_median_percent_plot, gc_median_percent_plot,
  labels = c("A", "B"), 
  ncol = 2, nrow = 1,  common.legend = TRUE, legend = "bottom"
)

summary_plots

#ggsave("./06-publication/supplemental_figures/SI_Fig17CD_NPre_classified_reads_gc_rl_boxplot.pdf", plot = summary_plots, device = "pdf", scale = 1, width = 10, height = 5, units = c("in"), dpi = 300)
```


## Classified read length distributions of all samples
```{r}
#function that creates a plot per individual ID and then arranges them all together using ggpubr (ggarange)
make_rl_plot <- function(individual_id, rl_classified, metadata_no_rl) {

  rl_df <- rl_classified %>%
    filter(Individual_ID == individual_id) %>%
    mutate(total_counts = counts_classified / 10000) %>%
    mutate(Length = as.numeric(Length))

  median_rl_info <- metadata_no_rl %>%
    filter(Individual_ID == individual_id) %>%
    select(Strandedness, median_rl)

  rl_plot <- ggplot(rl_df, aes(x = Length, y = total_counts,
                         fill = Strandedness, linetype = Strandedness)) +
    geom_bar(
      data = rl_df %>% filter(Strandedness == "Double"),
      aes(x = Length, y = total_counts),
      fill = "#04387a", alpha = 1, stat = "identity"
    ) +
    geom_bar(
      data = rl_df %>% filter(Strandedness == "Single"),
      aes(x = Length, y = total_counts),
      fill = "#6fbfdc", alpha = 0.6, stat = "identity"
    ) +
    theme_minimal() +
    labs(
      title = individual_id,
      x = "Read Length (bp)",
      y = expression(`Classified Reads` ~ (x10^4))
    ) +
    guides(
      fill = guide_legend(title = "Strandedness"),
      linetype = guide_legend(title = "Strandedness")
    ) +
    scale_x_continuous(
      limits = c(30, 150),
      breaks = seq(30, 140, by = 15),
      expand = c(0, 0)) +
    theme(
      legend.position = "bottom",
      axis.text.x = element_text(angle = 90, hjust = 1, size = 12),
      axis.text.y = element_text(size = 12),
      axis.title = element_text(size = 14),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 14),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")
    ) +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl, linetype = Strandedness), linewidth = 0.5, color = "red")

  rl_plot
}

#Run the function for all individual IDs and then arrange
all_samples <- rl_classified %>%
  distinct(Individual_ID) %>%
  pull(Individual_ID)

plot_list <- purrr::map(
  all_samples,
  ~ make_rl_plot(.x, rl_classified = rl_classified, metadata_no_rl = metadata_no_rl)
)

classified_rl_plots <- ggarrange(plotlist = plot_list, nrow = 2, ncol = 7, common.legend = TRUE, legend = "bottom")

classified_rl_plots

#ggsave("./06-publication/supplemental_figures/SI_Fig15B_NPre_classified_reads_rl_plots_all.pdf", plot = classified_rl_plots, device = "pdf", scale = 1, width = 16, height = 6, units = c("in"), dpi = 300)


```


## GC content distribution of classified reads across all samples
```{r}
#function that creates a plot per individual ID and then arranges them all together using ggpubr (ggarange)
make_gc_plot <- function(individual_id, gc_classified, metadata_no_rl) {

  gc_df <- gc_classified %>%
    filter(Individual_ID == individual_id) %>%
    mutate(total_counts = counts_classified / 10000) %>%
    mutate(gc = as.numeric(as.character(gc))) 

  median_gc_info <- metadata_no_rl %>%
    filter(Individual_ID == individual_id) %>%
    select(Strandedness, median_gc)

  gc_plot <- ggplot(gc_df, aes(x = gc, y = total_counts,
                         fill = Strandedness, linetype = Strandedness)) +
    geom_bar(
      data = gc_df %>% filter(Strandedness == "Double"),
      aes(x = gc, y = total_counts),
      fill = "#04387a", alpha = 1, stat = "identity"
    ) +
    geom_bar(
      data = gc_df %>% filter(Strandedness == "Single"),
      aes(x = gc, y = total_counts),
      fill = "#6fbfdc", alpha = 0.6, stat = "identity"
    ) +
    theme_minimal() +
    labs(
      title = individual_id,
      x = "%GC",
      y = expression(`Classified Reads` ~ (x10^4))
    ) +
    guides(
      fill = guide_legend(title = "Strandedness"),
      linetype = guide_legend(title = "Strandedness")
    ) +
    scale_x_continuous(
      limits = c(0, 101),
      breaks = seq(10, 90, by = 10),
      expand = c(0, 0)) +
    theme(
      legend.position = "bottom",
      axis.text.x = element_text(angle = 90, hjust = 1, size = 12),
      axis.text.y = element_text(size = 12),
      axis.title = element_text(size = 14),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 14),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")
    ) +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) +
    geom_vline(data = median_gc_info, aes(xintercept = median_gc, linetype = Strandedness), linewidth = 0.5, color = "red")

  gc_plot
}

#Run the function for all individual IDs and then arrange
all_samples <- gc_classified %>%
  distinct(Individual_ID) %>%
  pull(Individual_ID)

plot_list <- purrr::map(
  all_samples,
  ~ make_gc_plot(.x, gc_classified = gc_classified, metadata_no_rl = metadata_no_rl)
)

classified_gc_plots <- ggarrange(plotlist = plot_list, nrow = 2, ncol = 7, common.legend = TRUE, legend = "bottom")

classified_gc_plots

#ggsave("./06-publication/supplemental_figures/SI_Fig16B_NPre_classified_reads_gc_plots_all.pdf", plot = classified_gc_plots, device = "pdf", scale = 1, width = 17, height = 6, units = c("in"), dpi = 300)


```


#Oral reads
## Median read length of oral reads
```{r}
#color by the percentage of reads classified as oral

#read in file with each the number of reads classified by kraken2 by read length (per library)
rl_oral <- fread("./05-results/normalize_eager/dsvss_npre_oral_read_length.tsv") %>%
  mutate(Study = str_replace(Study, ".oral_gc_rl.tsv.gz", "")) %>%
  separate(Study, into = "Sequencing_ID", sep = ".tract", extra = "drop") %>%
  full_join(., metadata_no_rl) %>%
  dplyr::rename(counts_oral = counts_length)

rl_levels <- c(30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 73, 74, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99, 100, 101, 102, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 113, 114, 115, 116, 117, 118, 119, 120, 121, 122, 123, 124, 125, 126, 127, 128, 129, 130, 131, 132, 133, 134, 135, 136, 137, 138, 139, 140, 141, 142, 143, 144, 145, 146, 147, 148, 149, 150)

rl_summary_oral <- rl_oral %>%
    mutate(Length = factor(Length, level=rl_levels)) 

#function to calculate the median read length of oral reads per libraryID
median_rl_summary_table <- function(metadata, sample_list, oral_reads_length_table) {
  
  #Create a blank dataframe
  read_length_summary_table <- data.frame(Individual_ID = sample_list,
                                          Double = numeric(length(sample_list)),
                                          Single = numeric(length(sample_list)),
                                          stringsAsFactors = FALSE)
  
  #Function to calculate median read length based on dataset, sample, and strandedness
  calculate_median_rl <- function(data, sample, strandedness) {
    rl_summary <- data %>%
      filter(!str_detect(Library_ID, "bp")) %>%
      dplyr::select(Library_ID, Individual_ID, Strandedness, Length, counts_oral) %>%
      mutate(Length = as.integer(as.character(Length))) %>%
      filter(Individual_ID == sample, Strandedness == strandedness) %>%
      uncount(counts_oral) %>%
      mutate(Length = as.integer(Length)) %>%
      summarize(median_rl = median(Length))
    
    return(rl_summary$median_rl)
  }
  
  
  for (sample in sample_list) {
    #adapter removal
    rl_value_of_median_ds <- calculate_median_rl(oral_reads_length_table, sample, "Double")
    rl_value_of_median_ss <- calculate_median_rl(oral_reads_length_table, sample, "Single")
    read_length_summary_table[read_length_summary_table$Individual_ID == sample, "Double"] <- rl_value_of_median_ds
    read_length_summary_table[read_length_summary_table$Individual_ID == sample, "Single"] <- rl_value_of_median_ss
    
   }

  return(read_length_summary_table)
  
}

samples <- metadata_no_rl %>%
  dplyr::select(Individual_ID) %>%
  unique() %>%
  pull()

median_rl <- median_rl_summary_table(metadata_no_rl, samples, rl_summary_oral) %>%
  pivot_longer(-Individual_ID, values_to = "median_rl_oral", names_to = "Strandedness")

#calculate the percent of reads classified
kraken_10M_oral_species_table <- oral_species_table_no_rl %>%
  adorn_totals() %>%
  filter(Species == "Total") %>%
  pivot_longer(!Species, names_to = "Sequencing_ID", values_to = "oral_reads") %>%
  dplyr::select(-Species)

metadata_no_rl <- metadata_no_rl %>%
  full_join(., kraken_10M_oral_species_table) %>%
  full_join(., median_rl) %>%
  mutate(percent_oral = (oral_reads/10000000)*100) %>%
  mutate(Library = Strandedness) %>%
  mutate(Library = str_replace_all(Library, "Double", "dsLib"),
         Library = str_replace_all(Library, "Single", "ssLib"))

```


```{r}
wilcox_results <- metadata_no_rl %>%
  wilcox_test(median_rl_oral ~ Library, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- metadata_no_rl %>%
  wilcox_effsize(formula = median_rl_oral ~ Library, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat))

annotations <- wilcox_results %>%
  summarize(annotation = first(annotation),
            y_position = 77, .groups = 'drop')

read_length_median_percent_plot <- metadata_no_rl %>%
  ggplot(., aes(x = Library, y = median_rl_oral)) + 
  geom_boxplot() +
  geom_jitter(width = 0.075, size = 3, aes(color = percent_oral)) +
  scale_color_viridis(option = "viridis", begin = 0.1) +
  theme_classic() +
  scale_shape_manual(values = c(16, 18)) +
  labs(x = "Library Type",
       y = "Median Read Length of Oral Reads",
       color = "Percent of Reads Classified as Oral") +
  expand_limits(y = 80) +
  guides(alpha = "none") +
  geom_text(data = annotations, aes(x = 1.5, y = y_position, label = annotation), vjust = -0.5, size = 5) +
  theme(axis.text.x = element_text( size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
  geom_line(aes(group = Individual_ID), linetype = "dashed", color = "#555555", linewidth = 0.2) +
  geom_text_repel(aes(label = Individual_ID), size = 2)

read_length_median_percent_plot

```


## Median %GC of classified reads
```{r}
gc_levels <- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 73, 74, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99, 100, 101, 102, 103, 104)

#read in file with each the number of reads classified by kraken2 by %GC (per library)
gc_oral <- fread("./05-results/normalize_eager/dsvss_npre_oral_gc.tsv") %>%
  mutate(Study = str_replace(Study, ".oral_gc_rl.tsv.gz", "")) %>%
  separate(Study, into = "Sequencing_ID", sep = ".tract", extra = "drop") %>%
  full_join(., metadata_no_rl) %>%
  dplyr::rename(counts_oral = counts_gc) %>%
  separate(gc_bin, into = c("drop", "gc"), sep = ",", remove = FALSE) %>%
  mutate(gc = str_remove(gc, "\\]")) %>%
  dplyr::select(-gc_bin, -drop) %>%
  mutate(gc = as.factor(gc))

gc_summary_oral <- gc_oral %>%
    mutate(gc = factor(gc, level=gc_levels)) 

#function to calculate the median read length of oral reads per libraryID
median_gc_summary_table <- function(metadata, sample_list, oral_reads_gc_table) {
  
  #Create a blank dataframe
  gc_summary_table <- data.frame(Individual_ID = sample_list,
                                          Double = numeric(length(sample_list)),
                                          Single = numeric(length(sample_list)),
                                          stringsAsFactors = FALSE)
  
  #Function to calculate median read length based on dataset, sample, and strandedness
  calculate_median_gc <- function(data, sample, strandedness) {
    gc_summary <- data %>%
      filter(!str_detect(Library_ID, "bp")) %>%
      dplyr::select(Library_ID, Individual_ID, Strandedness, gc, counts_oral) %>%
      filter(Individual_ID == sample, Strandedness == strandedness) %>%
      mutate(gc = as.numeric(as.character(gc))) %>%
      uncount(counts_oral) %>%
      summarize(median = median(gc))
    
    return(gc_summary$median)
  }
  
  
  for (sample in sample_list) {
    #adapter removal
    gc_value_of_median_ds <- calculate_median_gc(oral_reads_gc_table, sample, "Double")
    gc_value_of_median_ss <- calculate_median_gc(oral_reads_gc_table, sample, "Single")
    gc_summary_table[gc_summary_table$Individual_ID == sample, "Double"] <- gc_value_of_median_ds
    gc_summary_table[gc_summary_table$Individual_ID == sample, "Single"] <- gc_value_of_median_ss
    
   }

  return(gc_summary_table)
  
}

samples <- metadata_no_rl %>%
  dplyr::select(Individual_ID) %>%
  unique() %>%
  pull()

median_gc <- median_gc_summary_table(metadata_no_rl, samples, gc_summary_oral) %>%
  pivot_longer(-Individual_ID, values_to = "median_gc_oral", names_to = "Strandedness")

metadata_no_rl <- metadata_no_rl %>%
  full_join(., median_gc)

```


```{r}
wilcox_results <- metadata_no_rl %>%
  wilcox_test(median_gc_oral ~ Library, paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")

effect_size_labels <- metadata_no_rl %>%
  wilcox_effsize(formula = median_gc_oral ~ Library, paired = TRUE) %>%
  mutate(effectsize_cat = case_when(
    magnitude == "small" ~ "†",     
    magnitude == "moderate" ~ "††",
    magnitude == "large" ~ "†††"))

wilcox_results <- wilcox_results %>%
  left_join(effect_size_labels) %>%
  mutate(annotation = paste(p.adj.signif, effectsize_cat))

annotations <- wilcox_results %>%
  summarize(annotation = first(annotation),
            y_position = 65, .groups = 'drop')

gc_median_percent_plot <- metadata_no_rl %>%
  ggplot(., aes(x = Library, y = median_gc_oral)) + 
  geom_boxplot() +
  geom_jitter(width = 0.075, size = 3, aes(color = percent_oral)) +
  scale_color_viridis(option = "viridis", begin = 0.1) +
  theme_classic() +
  scale_shape_manual(values = c(16, 18)) +
  labs(x = "Library Type",
       y = "Median %GC of Oral Reads",
       color = "Percent of Reads Classified as Oral") +
  expand_limits(y = 70) +
  guides(alpha = "none") +
  geom_text(data = annotations, aes(x = 1.5, y = y_position, label = annotation), vjust = -0.5, size = 5) +
  theme(axis.text.x = element_text( size = 12), 
          axis.text.y = element_text(size = 12), 
          axis.title = element_text(size = 14),  
          legend.text = element_text(size = 12),  
          legend.title = element_text(size = 14),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
  geom_line(aes(group = Individual_ID), linetype = "dashed", color = "#555555", linewidth = 0.2) +
  geom_text_repel(aes(label = Individual_ID), size = 2)

gc_median_percent_plot
```


```{r}
summary_plots <- ggarrange(
  read_length_median_percent_plot, gc_median_percent_plot,
  labels = c("A", "B"), 
  ncol = 2, nrow = 1,  common.legend = TRUE, legend = "bottom"
)

summary_plots

#ggsave("./06-publication/supplemental_figures/SI_Fig25CD_NPre_oral_reads_combined_plot.pdf", plot = summary_plots, device = "pdf", scale = 1, width = 10, height = 5, units = c("in"), dpi = 300)
```


## Read length distribution of classified reads across all samples
```{r}
#function that creates a plot per individual ID and then arranges them all together using ggpubr (ggarange)
make_rl_plot <- function(individual_id, rl_oral, metadata_no_rl) {

  rl_df <- rl_oral %>%
    filter(Individual_ID == individual_id) %>%
    mutate(total_counts = counts_oral / 1000) %>%
    mutate(Length = as.numeric(as.character(Length))) 

  median_rl_info <- metadata_no_rl %>%
    filter(Individual_ID == individual_id) %>%
    select(Strandedness, median_rl_oral)

  rl_plot <- ggplot(rl_df, aes(x = Length, y = total_counts,
                         fill = Strandedness, linetype = Strandedness)) +
    geom_bar(
      data = rl_df %>% filter(Strandedness == "Double"),
      aes(x = Length, y = total_counts),
      fill = "#04387a", alpha = 1, stat = "identity"
    ) +
    geom_bar(
      data = rl_df %>% filter(Strandedness == "Single"),
      aes(x = Length, y = total_counts),
      fill = "#6fbfdc", alpha = 0.6, stat = "identity"
    ) +
    theme_minimal() +
    labs(
      title = individual_id,
      x = "Read Length (bp)",
      y = expression(`Oral Reads` ~ (x10^3))
    ) +
    guides(
      fill = guide_legend(title = "Strandedness"),
      linetype = guide_legend(title = "Strandedness")
    ) +
    scale_x_continuous(
      limits = c(30, 150),
      breaks = seq(30, 140, by = 15),
      expand = c(0, 0)) +
    theme(
      legend.position = "bottom",
      axis.text.x = element_text(angle = 90, hjust = 1, size = 12),
      axis.text.y = element_text(size = 12),
      axis.title = element_text(size = 14),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 14),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")
    ) +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) +
    geom_vline(data = median_rl_info, aes(xintercept = median_rl_oral, linetype = Strandedness), linewidth = 0.5, color = "red")

  rl_plot
}

#Run the function for all individual IDs and then arrange
all_samples <- rl_oral %>%
  distinct(Individual_ID) %>%
  pull(Individual_ID)

plot_list <- purrr::map(
  all_samples,
  ~ make_rl_plot(.x, rl_oral = rl_oral, metadata_no_rl = metadata_no_rl)
)

oral_rl_plots <- ggarrange(plotlist = plot_list, nrow = 2, ncol = 7, common.legend = TRUE, legend = "bottom")

oral_rl_plots

#ggsave("./06-publication/supplemental_figures/SI_Fig23B_NPre_oral_reads_rl_plots_all.pdf", plot = oral_rl_plots, device = "pdf", scale = 1, width = 16, height = 6, units = c("in"), dpi = 300)


```


## GC content distribution of classified reads across all samples
```{r}
#function that creates a plot per individual ID and then arranges them all together using ggpubr (ggarange)
make_gc_plot <- function(individual_id, gc_oral, metadata_no_rl) {

  gc_df <- gc_oral %>%
    filter(Individual_ID == individual_id) %>%
    mutate(total_counts = counts_oral / 1000) %>%
    mutate(gc = as.numeric(as.character(gc)))

  median_gc_info <- metadata_no_rl %>%
    filter(Individual_ID == individual_id) %>%
    select(Strandedness, median_gc_oral)

  gc_plot <- ggplot(gc_df, aes(x = gc, y = total_counts,
                         fill = Strandedness, linetype = Strandedness)) +
    geom_bar(
      data = gc_df %>% filter(Strandedness == "Double"),
      aes(x = gc, y = total_counts),
      fill = "#04387a", alpha = 1, stat = "identity"
    ) +
    geom_bar(
      data = gc_df %>% filter(Strandedness == "Single"),
      aes(x = gc, y = total_counts),
      fill = "#6fbfdc", alpha = 0.6, stat = "identity"
    ) +
    theme_minimal() +
    labs(
      title = individual_id,
      x = "%GC",
      y = expression(`Oral Reads` ~ (x10^3))
    ) +
    guides(
      fill = guide_legend(title = "Strandedness"),
      linetype = guide_legend(title = "Strandedness")
    ) +
    scale_x_continuous(
      limits = c(0, 100),
      breaks = seq(10, 90, by = 10),
      expand = c(0, 0)) +
    theme(
      legend.position = "bottom",
      axis.text.x = element_text(angle = 90, hjust = 1, size = 12),
      axis.text.y = element_text(size = 12),
      axis.title = element_text(size = 14),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 14),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.margin = margin(t = 5, r = 20, b = 5, l = 5, unit = "pt")
    ) +
    scale_linetype_manual(values = c("Double" = "solid", "Single" = "dashed")) +
    #geom_vline(data = median_gc_info, aes(xintercept = factor(median_gc), linetype = Strandedness), linewidth = 0.5, color = "red") 
    geom_vline(data = median_gc_info, aes(xintercept = median_gc_oral, linetype = Strandedness), linewidth = 0.5, color = "red")

  gc_plot
}

#Run the function for all individual IDs and then arrange
all_samples <- gc_oral %>%
  distinct(Individual_ID) %>%
  pull(Individual_ID)

plot_list <- purrr::map(
  all_samples,
  ~ make_gc_plot(.x, gc_oral = gc_oral, metadata_no_rl = metadata_no_rl)
)

oral_gc_plots <- ggarrange(plotlist = plot_list, nrow = 2, ncol = 7, common.legend = TRUE, legend = "bottom")

oral_gc_plots

#ggsave("./06-publication/supplemental_figures/SI_Fig24B_NPre_oral_reads_gc_plots_all.pdf", plot = oral_gc_plots, device = "pdf", scale = 1, width = 17, height = 6, units = c("in"), dpi = 300)


```

